### 2023-04-13 forwards-and-backwards.tex
added MMT

\documentclass[12pt]{article}
\usepackage{fullpage}
\usepackage{amssymb}
\usepackage{ntheorem}
\usepackage{graphics} 
\usepackage{amsmath}
\usepackage[color=yellow]{todonotes}
\usepackage{url}
\usepackage{xcolor}
\usepackage{longtable}
\usepackage[hidelinks]{hyperref}
\usepackage{bbm}

\usepackage{natbib}

\newenvironment {proof}{{\noindent\bf Proof }}{\hfill $\Box$ \medskip}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{hypothesis}[theorem]{Hypothesis}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{example}[theorem]{Example}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{assumptions}[theorem]{Assumptions}
\newtheorem{terminology}[theorem]{Terminology}
\newtheorem*{terminology-non}[theorem]{Terminology}

\renewcommand{\theequation}{\arabic{section}.\arabic{equation}}
\def \non{{\nonumber}}
\def \hat{\widehat}
\def \tilde{\widetilde}
\def \bar{\overline}
\newcommand{\IP}{\mathbb P}
\newcommand{\IQ}{\mathbb Q}
\newcommand{\IE}{\mathbb E}
\newcommand{\IR}{\mathbb R}
\newcommand{\IZ}{\mathbb Z}
\newcommand{\IN}{\mathbb N}
\newcommand{\IT}{\mathbb T}
\newcommand{\IC}{\mathbb C}
\newcommand{\ind}{\mathbf{1}}
\newcommand{\bigO}{\mathcal{O}}
\newcommand{\grad}{\nabla}
\newcommand{\dif}{\mathrm{d}\,}

%%%%%%% notation
\newcommand{\DG}{\mathcal{B}}  % generator of the dispersal process
\newcommand{\DD}{\mathcal{D}}  % the second-order part of the generator of dispersal
\newcommand{\meanq}{\vec b}    % mean of dispersal, times theta
\newcommand{\covq}{\mathbf{C}}     % covariance matrix of dispersal, times theta
\newcommand{\kernel}{\rho}  % interaction kernels
\newcommand{\smooth}[1]{\kernel_{#1} \! * \!}  % convolution by the interaction kernel
\newcommand{\wavespeed}{\mathfrak{c}}    % speed (vector) of a wave
\newcommand{\Lgen}{\mathcal{L}}    % generator of a lineage
\newcommand{\Pgen}{\mathcal{P}}    % generator of the population process
\newcommand{\lp}{\xi}              % process with levels
\newcommand{\labelspace}{\mathcal{I}} % space of labels
\newcommand{\concat}{\oplus}   % concatenation of labels
\newcommand{\measures}{\mathcal{M}_F(\IR^d)} % finite measures on Rd
\newcommand{\cmeasures}{\mathcal{M}_F(\overline{\IR}^d)} % finite measures on compactified Rd
\newcommand{\lpmeasures}{\mathcal{M}(\overline{\IR^d} \times [0,\infty))} % locally finite measures on (compactified space) x level
\newcommand{\flpmeasures}{\mathcal{M}(\IR^d \times [0,\infty))} % locally finite measures on space x level

\newcommand{\plr}[1]{\todo[inline]{Peter: #1}}
\newcommand{\comment}[1]{{\color{blue} \it #1}}

\begin{document}

\title{\large{\bf
Looking forwards and backwards through locally regulated populations
}}

\author{ \begin{small}
\begin{tabular}{ll}                              
Alison M. Etheridge 
 & Thomas G. Kurtz \\   
Department of Statistics & Departments of Mathematics and Statistics\\       
Oxford University & University of Wisconsin - Madison \\                   
24-29 St Giles & 480 Lincoln Drive\\                                                         
Oxford OX1 3LB & Madison, WI  53706-1388\\
UK & USA \\                        
etheridg@stats.ox.ac.uk & kurtz@math.wisc.edu     \\
\url{http://www.stats.ox.ac.uk/~etheridg/} & 
\url{http://www.math.wisc.edu/~kurtz/}  \\       \\
\\
Ian Letter&  Peter L. Ralph 
\\   
Department of Statistics & Departments of Mathematics and Biology\\
Oxford University &University of Oregon\\                   
24-29 St Giles & Fenton Hall\\
Oxford OX1 3LB & Eugene, OR 97403-1222\\
UK & USA \\
restucci@stats.ox.ac.uk  & plr@oregon.edu \\
\url{https://www.stats.ox.ac.uk/~restucci/}&
\url{https://math.uoregon.edu/profile/plr} \\
\\
Terence Tsui 
 &  \\   
Department of Statistics & \\
Oxford University & \\                   
24-29 St Giles& \\
Oxford OX1 3LB & \\
UK & \\
terence.tsui@sjc.ox.ac.uk &      \\
\url{https://www.maths.ox.ac.uk/people/terence.tsui}&  \\
\end{tabular}
\end{small}}

\date{\today}
\maketitle

\comment{get equations to restart numbering in sections}

\begin{abstract}

We introduce a broad class of mechanistic spatial models to describe how spatially 
heterogeneous populations live, die, and reproduce.
Individuals are represented by points of a point measure,
whose birth and death rates can depend 
both on spatial position and local population density, defined at a location
to be the convolution of the point measure
with a suitable non-negative integrable kernel
centred on that location. We pass to three different scaling limits:
an interacting superprocess, a nonlocal partial differential equation (PDE), 
and a classical PDE. 
The classical PDE is obtained both by a two-step convergence argument, in which we 
first scale time and population size and
pass to the nonlocal PDE, 
and then scale the kernel that 
determines local population density; and 
in the important special case in which the limit is a 
reaction diffusion equation, directly 
by simultaneously scaling the kernel and the
timescale and population size in our individual based model.

A novelty of our model is that we explicitly
model a juvenile phase: the number of juveniles produced by an individual depends
on local population density at the location of the parent. 
Juvenile offspring are 
thrown off in a (possibly heterogeneous, anisotropic) Gaussian distribution 
around the location of the parent; they then reach (instant) maturity with a probability
that can depend on the local population density
% (possibly measured through convolution with a different kernel from that used in determining fecundity) 
at the location at which they land. 
Although we only record mature individuals, a trace of this two-step description 
remains in our population models, resulting in novel limits in which the spatial
motion is governed by a nonlinear diffusion.

Using a lookdown representation, we are able to retain information about 
genealogies relating individuals in our population and, in the case of 
deterministic limiting models, we use this to deduce 
the backwards in time motion of the ancestral lineage of an individual
sampled from the population. We observe that
knowing the history of the population density is not enough to determine the motion of
ancestral lineages in our model. We also investigate (and
contrast) the behaviour of lineages for 
three different deterministic models of a population expanding its range as a 
travelling wave: the Fisher-KPP equation, the Allen-Cahn equation, and a porous 
medium equation with logistic growth.

\vspace{.1in}

	\comment{Need to thank KAVLI, Paris Cit\'e, various funders}

\vspace{.1in}

\noindent {\bf Key words:}  population model, interacting superprocess, 
lookdown construction, porous medium equation, 
reaction-diffusion equation, travelling waves, genealogies,
Fisher-KPP equation

\vspace{.1in}

\noindent {\bf MSC 20}10 {\bf Subject Classification:}  Primary:  
%60J25, 92D10, 92D15, 92D25 92D40  
\\Secondary:   %60F05, 60G09, 60G55, 60G57, 60H15, 60J68

\end{abstract}
\tableofcontents
\newpage

%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\label{introduction}

As one takes a journey, long or short, the landscape changes:
forests thicken or thin or change their composition;
springtime grasslands host intergrading mosaics of different types of flowers.
The aim of this paper is to introduce and study
a broad class of mechanistic spatial models
that might describe how spatially heterogeneous populations live, die, and reproduce.
Questions that we (start to) address include:
How does population density change across space and time?
How might we learn about the underlying dynamics from genealogical or genetic data?
And, how does genetic ancestry spread across geography
when looking back through time in these populations?

Reproduction of individuals naturally leads to spatial branching process models,
including branching random walk, 
branching Brownian motion, and the Dawson-Watanabe superprocesses.  
However, as a result of the branching assumption (once born, individuals behave 
independently of one another), 
a population evolving according to any of these models will
either die out or grow without bound and, in so doing, can
develop clumps of arbitrarily large density and extent. 
Our starting point here is an individual-based model of a single species 
in continuous space in which
birth, death, and establishment may all depend on local population density
as well as on spatial location,
allowing for stable populations through density-dependent feedback.

Although it is often mathematically convenient to assume that individuals follow
Brownian motion during their lifetime, 
in our model, offspring are thrown off according to some 
spatial distribution centred on the location of the parent
and do not subsequently move.
This is particularly appropriate for modelling plant populations, in which
this dispersal of offspring around the parent is the only source of
spatial motion.

Most models do not distinguish between juveniles and adults, so,
for example, the number of adults produced by a single parent is determined
by the degree of crowding at the location of the parent.
Although we shall similarly only follow the adult population, in formulating the dynamics of the
models we shall distinguish
between production of juveniles, which will depend upon the location of 
the adult, and their successful establishment, which will depend on the
location in which a juvenile lands. The result is that not only the absolute 
number, but also the spatial distribution
around their parent, 
of those offspring that survive to adulthood
will depend upon the local population 
density.

We shall consider three different classes of scaling limits for our model.
The first yields a class of (generalised) superprocesses in which coefficients 
governing both the spatial motion and the branching components of the process can depend
on local population density; the second is a corresponding class of deterministic
non-local differential equations; and the third are classical PDEs.
We measure local population density around a point by convolving with
a smooth kernel $\rho(\cdot)$, which may differ
for the two stages of reproduction. 
When the limiting population process is deterministic,
it is a (weak) solution of an equation of the form
\begin{equation}
	\label{general deterministic limit}
        \partial_t \varphi_t(x)
        =
        r\left(x, \varphi_t \right)
        \DG^* \left[
            \varphi_t(\cdot)
            \gamma\big( x, \varphi_t \big)
        \right](x)
        +
        \varphi_t(x)
        F\left(x, \varphi_t \right)
        ,
\end{equation}
where
$\varphi_t(x)$ can be thought of as the population density at $x$
(although the limit may be a measure without a density), and
$\DG^*$ is (the adjoint of) a strictly uniformly elliptic second order differential operator, typically the Laplacian.
The dependence of each of the terms $r$, $\gamma$, and $F$ on $\varphi$ is only through the local density at $x$,
e.g., $F(x, \varphi) = F(x, \smooth{} \varphi(x))$.
We shall be more specific about the parameters below.

By replacing $\rho$ by
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon^d$, 
we can also scale the ``width'' of the 
region over which we measure local population density. 
When the population follows~(\ref{general deterministic limit}), 
we expect that if we take a second limit of $\epsilon \to 0$,
thus scaling the kernels appearing in $r$, $\gamma$, and $F$ and making interactions pointwise,
we should recover a nonlinear PDE.
We verify that this is indeed the case in two important examples:
a special case of the porous medium equation with a logistic growth term, in which we 
retain a signal of our two stage approach to modelling in the nonlinear diffusion term
and the limiting equation takes the form 
\begin{equation}
	\label{PME1}
	\partial_t \varphi = \Delta (\varphi^2)+\varphi(1-\varphi);
\end{equation}
and a wide class of semi-linear PDEs of the form 
\begin{equation}
\label{semilinear PDE}
\partial_t \varphi = \DG^*\varphi+ \varphi F(\varphi),
\end{equation}
which includes the Fisher-KPP equation and the Allen-Cahn equation.
Equations of this form have been studied extensively
in the context of spatial ecology;
see for instance \citet{lam2023introduction} and \citet{cantrell2004spatial}.

It is of interest to understand under what conditions we can replace the two-step limiting
process described above by one in which we simultaneously scale the 
kernels
and the other parameters in 
our population model to arrive at the PDE limit.
This is mathematically much more challenging, but we establish such 
one-step convergence in cases for which the limit is a classical 
reaction diffusion equation of the form~(\ref{semilinear PDE}) with 
$\DG=\Delta$,
and $\rho_F$ is a Gaussian density. We allow a wide class of reaction terms, $F$, so
that the Fisher-KPP equation 
(that is equation~(\ref{semilinear PDE}) with $\DG=\Delta$ and $F(\varphi)=1-\varphi$)
emerges as a special case.

We remark that such results on (one-step) convergence to reaction diffusion 
equation limits have been
achieved for a variety of interacting particle systems. The vast majority of these
results concern spin systems \comment{REFs}. However,~\cite{flandoli/huang:2021} 
obtain the Fisher-KPP equation
as a scaling limit of interacting diffusions in
$\IR^d$. In that setting a major challenge (also apparent in our models), not
faced in spin systems, is 
the control of concentration of particles. We also note that the scaling 
considered in~\cite{flandoli/huang:2021} is different from that considered
below. In their language we consider only `moderate interactions', meaning that
the number of individuals in the neighbourhood over which we measure local
population density tends to infinity, whereas they also consider the situation in which
that number remains finite;
furthermore we shall consider a scaling in which 
the reproduction mechanism is near-critical and look at large timescales,
while \citet{flandoli/huang:2021} only scale the number of
particles in the system and the size of the neighbourhood over which indivduals interact.

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity that have been laid down; from 
genetic data, one can try to infer the genealogical trees that relate 
individuals in a sample from the population, and these have been shaped 
by its history \citep{neigel1993application,kelleher2019inferring}.
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model.
To achieve this we exploit a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a huge range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

In general, even armed with a lookdown construction, calculation of relevant statistics 
of the genealogy remains a difficult question. 
However, in special
circumstances, some progress can be made. As an illustration, 
we shall consider a scenario that
has received a great deal of attention in recent years, in which 
a population is expanding into new territory as a travelling wave. 
In Section~\ref{ancestral lineages for travelling waves}
we shall consider three different (deterministic) wavefronts, moving 
across $\IR^1$, and we shall explore the position of a single ancestral 
lineage relative to that front.

Most work on this topic has focused on models that caricature the classical Fisher-KPP equation with a 
stochastic term, i.e.
$$dw=\big(\Delta w +sw(1-w)\big)dt +\sqrt{\frac{\alpha(w)}{N}}W(dt,dx),$$
where $W$ is space-time white noise, and $N$ is a measure of the 
local population density. The coefficient
$\alpha(w)$ is generally taken to be either
$w$, corresponding to a superprocess limit, or $w(1-w)$ giving a 
spatial analogue of a Wright-Fisher diffusion. 
For instance, \citep{ghosh2022emergent}
derives a stochastic version of~\eqref{semilinear PDE} to describe abundances of mutant bacteria strains
along the human gut.
For very large $N$, and on suitable timescales, 
starting with the pioneering work of~\cite{brunet/derrida/mueller/munier:2006},
a considerable body of evidence has been amassed to underpin the
conjecture that for this, and a wide class of 
related models, 
genealogies converge to a 
Bolthausen-Sznitman coalescent. %, e.g.~\cite{brunet/derrida/mueller/munier:2007}.
This reflects the fact that, for this equation, 
ancestral lineages become trapped in 
the wavefront, where the growth rate of the population is highest. 
Once there, they will experience 
rapid periods of coalescence 
corresponding to significant proportions of individuals in the front being 
descended from particularly reproductively successful ancestors.

If one replaces the logistic growth term of the classical
Fisher-KPP equation with a nonlinearity that reflects cooperative
behaviour in the population, such as
\begin{equation}
	\label{Birzu nonlinearity}
	wF(w)=w(1-w)(Cw-1),
\end{equation}
then, for sufficiently large $C$ (strong cooperation),
the nature of the deterministic
wave changes from ``pulled'' to ``pushed'', 
\citep{birzu/hallatschek/korolev:2017, birzu/hallatschek/korolev:2021}.
In that setting, the genealogies will be quite different
from the Fisher-KPP case. 
%\comment{Could refer to more recent Birzu et al paper here.}
For example, \citet{etheridge/penington:2022}
show that for a discrete space model corresponding to this 
nonlinearity with $C>2$, after suitable scaling, the genealogy of a
sample converges not to a Bolthausen-Sznitman coalescent, but to
a Kingman coalescent. 
The reason, roughly, is that ancestral lineages
settle to a stationary distribution relative to the position of the 
wavefront which puts very little weight close to the `tip' of the wave, so
that when ancestral lineages meet
it is typically at a location in which population density is
high, where no single ancestor produces a disproportionately large number of 
descendants in a short space of time.

The shape of the wave is not determined solely by the reaction term.
For example, as a result of the nonlinear diffusion, 
for suitable initial conditions, the solution to the one-dimensional 
porous medium equation with logistic growth~(\ref{PME1}) converges to a travelling
wave with a sharp cut-off; i.e., in contrast to the classical 
Fisher KPP equation, the solution at time $t$
vanishes beyond $x=x_0+ct$ for some constant wavespeed $c>0$, \citep{kamin/rosenau:2004}.
As a first step towards understanding what we should expect in models with
nonlinear diffusion, one can ask about the position of an ancestral lineage
relative to the wavefront in the deterministic models. In 
Section~\ref{ancestral lineages for travelling waves}
we shall see that in our framework, even with logistic growth, the nonlinear diffusion 
corresponding to the porous medium equation results
in a stationary distribution for the ancestral lineage
that is concentrated behind the wavefront, leading 
us to conjecture that in the stochastic equation
the cooperative behaviour captured by the nonlinear diffusion will also 
result in a qualitatively different pattern of coalescence to that seen under the 
stochastic Fisher-KPP equation. Indeed, we believe that it should be feasible to 
show that in an appropriate limit one recovers a Kingman coalescent.

% % % % % % % % % % % % % %
\paragraph{Structure of the paper}

\comment{This paragraph to be rewritten last}

In this paper we study scaling limits of spatial population models,
obtaining convergence of both the population process
(i.e., the population density as a function of time,
although strictly speaking it is a measure that may not have a density)
and of lineages traced back through such a population.
We retain information about lineages as we pass to the scaling limit
by means of a lookdown construction.
First, in Section~\ref{sec: Model and main results},
we describe the model and the main results
(Theorems~\ref{thm:nonlocal_convergence}, \ref{thm:local_convergence}, and~\ref{thm:lineages}).
Next, in Section~\ref{sec:applications}, we discuss a few striking consequences of these results,
namely, X Y and Z \comment{TODO}.
In Section~\ref{sec:heuristics}, we provide heuristic explanations
of why the theorems ought to be true,
and some key ideas behind them.
In Section \ref{sec:lookdown} we define and discuss the lookdown construction.
This is followed by the technical proofs in Section~\ref{sec:proofs}.

\comment{What's in the Appendix?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Model and main results}
    \label{sec: Model and main results}

Our model is one of individuals distributed across a continuous space 
which we shall take to be $\IR^d$. 
For applications, $d=1$ or $d=2$
(or even $d=3$ for cells within the body),
but our main results apply more generally.
At time zero, the population is distributed over a bounded region, with 
$\bigO(N)$ individuals per unit area in that region,
so the total number of individuals will also be $\bigO(N)$.
The population changes in continuous time, and
we encode the state of the population at time $t$ by a counting measure $X(t)$,
which assigns one unit of mass to the location of each individual.

Population dynamics are controlled by three quantities,
birth ($\gamma$), establishment ($r$), and death ($\mu$),
each of which can depend on spatial location and local population density
in a way specified below.
Each individual gives birth at rate $\gamma$ to a single (juvenile) offspring,
which is dispersed according to a kernel $q(x, \cdot)$ away from the location $x$ of the parent.
We assume that $q$ is the density of a multivariate Gaussian,
allowing a nonzero mean and anisotropic variance.
Both the mean and covariance can change across space,
but do not depend on population density.
The offspring does not necessarily survive to be counted in the population:
it ``establishes'' with probability $r$,
or else it dies immediately.
Independently, each individual dies with rate $\mu$.

We aim to capture universal behaviour by passing to a scaling limit. Specifically, 
we shall take the ``density'', $N$, to infinity,
and also scale time by a factor of $\theta = \theta(N)$,
in such a way that defining $\eta^N(t) = X(\theta t) / N$,
the process $\{\eta^N(t)\}_{t \ge 0}$
will converge to a suitable measure-valued process
as $N$ and $\theta$ tend to infinity,
with the nature of the limit depending on how they tend to infinity together.
Evidently, we shall also need to scale the dispersal kernel if we are to obtain a
nontrivial limit. We write $q_\theta(x,\cdot)$ for the density of the 
multivariate Gaussian obtained by
multiplying the mean and variance components corresponding to $q(x,\cdot)$ by $1/\theta$.

Birth, establishment, and death can depend on the location of the individual
and the local population density.
Since we would like the population density to scale with $N$,
these are functions of $X/N$, i.e.,
the counting measure with mass $1/N$ placed at the location of each individual.
First consider birth rates, defined by
a nonnegative function $\gamma(x, m) : \IR^d \times \IR_{\ge 0} \to \IR_{\ge 0}$
of location $x$ 
and local population density $m$.
Local population density is defined as the convolution of $X/N$ with 
a smooth (non-negative, integrable) kernel $\rho_\gamma(\cdot)$.
We write this convolution as $\smooth{\gamma} X/N$.
Then, when the state of the population is $X$, an individual at location $x$ 
gives birth to a single juvenile offspring at rate
$\gamma(x, \smooth{\gamma} X(x) / N)$.
Similarly, the establishment probability of an offspring at location $y$ is
is $r(y, \smooth{r} X(y) / N)$,
where $r(y, m) : \IR^d \times \IR_{\ge 0} \to [0, 1]$
and again $\smooth{r} X$ is the convolution of 
$X/N$ with the smooth kernel $\rho_r$.

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of mature individuals
in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\[
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\]
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation defines $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as discussed later,
differs because offsprings' establishment probability is measured at their new location
rather than their parents', and so differs from $r(x)$.
For the most part, we work with $F$ instead of $\mu$.

Note that each of the three demographic parameters $r$, $\gamma$, $F$,
depends on local density,
measured by convolution with a smooth kernel. Each of the three kernels can be different.
As a result, death rate depends (in principle) on population densities measured in 
three different ways, 
so that we could write 
$\mu_\theta(x) = \mu_\theta(x, \smooth{\gamma} X(x) / N, \smooth{r} X(x) / N, \smooth{F} X(x) / N)$.
This may seem unnecessarily complex.
However, not only is it natural from a 
biological perspective, it also turns out to be necessary if we are
to capture nontrivial examples in our scaling results.

\begin{remark}
Although this model allows fairly general birth and death mechanisms,
there are a number of limitations.
Perhaps most obviously, to simplify the notation
individuals give birth to only one offspring at a time,
although this restriction could be easily lifted
\citep[as in Section 3.4 of][]{etheridge/kurtz:2019}.
Furthermore, individuals do not move during their lifetime,
and the age of an individual does not affect its fecundity or death rate.
Finally, there is no notion of mating
(although limitations on reproduction due to availability of mates can be incorporated into the birth rate, $\gamma$),
so the lineages we follow will be uniparental.
For these reasons, the model is most obviously applicable to bacterial populations or selfing plants,
although we do not anticipate that incorporation of these complications
will change the general picture.
\end{remark}

For each $N$ and $\theta$, the rescaled process
$\big(\eta^N_t\big)_{t\geq 0} := \big(X(\theta t)/N\big)_{t\geq 0}$, 
takes values in the space of c\`adl\`ag paths in
$\measures$ (the space of finite measures on $\IR^d$
endowed with the weak topology). In fact $\eta_t^N$ will be a purely atomic
measure comprised of atoms of mass $1/N$.

\begin{notation}
Expressions like $\gamma(x, \smooth{\gamma} \eta(x))$ will appear repeatedly in what follows.
To make formulae more readable, we overload notation to define
$$
    \gamma(x, \eta) := \gamma(x, \smooth{\gamma} \eta(x)) ,
$$
and similarly write $r(x, \eta)$ for $r(x, \smooth{r} \eta(x))$,
$F(x, \eta)$ for $F(x, \smooth{F} \eta(x))$,
and $\mu_\theta(x, \eta)$ for the expression of equation~\eqref{eqn:mu_defn}.
When convenient, we may also suppress the arguments completely,
writing simply $\gamma$, $r$, $F$, and $\mu_\theta$ for these quantities.
\end{notation}

\begin{terminology*}
In our prelimiting model,
the population is represented by a point measure in which each
individual is assigned a mass $1/N$. We use the term ``population density'' for
this process, as it is supposed to measure population size relative to a 
nominal occupancy of $N$ individuals per unit area. There is no implication 
that the measure representing the population is absolutely continuous with
respect to Lebesgue measure; indeed in the prelimit it is certainly not.
\end{terminology*}

In summary, at each time $t$, $\eta^N_t$ is purely atomic, consisting of atoms of mass $1/N$.
At instantaneous rate $\theta \gamma(x, \eta^N_t) N \eta^N_t(dx)$ an offspring of mass $1/N$ is 
produced at location $x$; it disperses to 
a location $y$ offset from $x$
by an independent Gaussian random variable with mean $\meanq(x) / \theta$
and covariance matrix $\covq(x) / \theta$, and once there establishes instantaneously
with probability $r(y, \eta^N_t)$, or else dies. 
At instantaneous rate $\theta \mu_\theta(x, \eta^N_t) N \eta^N_t(dx)$ an individual at location 
$x$ dies.
Note that the process 
$\left(\eta^N_t\right)_{t\ge 0}$, which
records numbers and locations of adult individuals, is just a scaled spatial birth and 
death process. If, for example, we insist that $\gamma(x,m)$ is bounded, then 
existence (and in particular non-explosion) is guaranteed by comparison with
a pure birth process. We do not dwell on this, as we shall require more stringent conditions
if we are to pass to the limit as $\theta$ and $N$ tend to infinity.

It is convenient to characterise the process as a solution to a martingale problem.
We write $C_b^\infty(\IR^d)$ for the space of bounded smooth functions on $\IR^d$, and, where
convenient, we
write $\langle f, \eta \rangle = \int_{\IR^d} f(x) \eta(dx)$.

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:mgale_construction}
	For each value of $N$ and $\theta$, and each purely atomic 
$\eta_0^N\in\measures$ with atoms of mass $1/N$,
$(\eta^N_t)_{t \geq 0}$ is the (scaled) empirical measure of a birth-death 
process with c\`adl\`ag paths in $\measures$ for which,
for all $f \in C^{\infty}_{b}(\IR^d)$,
writing $q_\theta(x, dy)$ for the Gaussian kernel
with mean $x + \meanq(x)/\theta$ and covariance $\covq(x) / \theta$,
\begin{equation}
    \label{eqn:eta_martingale}
\begin{aligned}
M^N_t(f)
&:=  \langle f, \eta^N_t \rangle
        -\langle f, \eta^N_0 \rangle
 \\ &\qquad {}
 -  \int_{0}^{t}\bigg\{
        \bigg\langle \left( \int \theta
             \left(
                f(z)r(z, \eta^N_{s})
                - f(x)r(x, \eta^N_{s})
            \right)
        q_\theta(x,dz) \right)
%\\ & \qquad \qquad \qquad \qquad {}
         \gamma(x, \eta^N_{s}), \eta^N_{s}(dx) \bigg\rangle
\\ & \qquad \qquad \qquad \qquad \qquad {}
    + \bigg\langle f(x) F(x, \eta^N_{s}),
    \eta^N_{s}(dx) \bigg\rangle
    \bigg\} ds
\end{aligned}    
\end{equation}
is a martingale (with respect to the natural filtration), 
with quadratic variation
    \begin{equation} \label{eqn:prelimit_martingale_variation}
\begin{aligned} \relax
[ M^N(f)]_t =& 
\frac{\theta}{N} \int_{0}^{t}\bigg\{
    \Big\langle \gamma(x, \eta^N_{s})
        \int f^2(z)r(z, \eta^N_{s}) q_\theta(x,dz) 
    , \eta^N_{s}(dx) \Big\rangle \\
& \qquad \qquad {}
    + \Big\langle \mu_\theta(x, \eta^N_{s}) f^2(x) 
    , \eta^N_{s}(dx)\Big\rangle
    \bigg\}ds. 
\end{aligned}    
\end{equation}
\end{definition}
The form of~(\ref{eqn:eta_martingale}) and~(\ref{eqn:prelimit_martingale_variation})
is explained in Section~\ref{sec:heuristics}.
Note that since (juvenile) individuals are produced at rate $N \gamma \eta$, but each
has mass $1/N$,
these factors of $N$ cancel in~\eqref{eqn:eta_martingale}. 
Under our scaling, $N$ and $\theta=\theta(N)$ will tend to infinity in such a 
way that $\alpha:=\lim_{N\to\infty}\theta(N)/N$ exists and is finite.
From the expression~(\ref{eqn:prelimit_martingale_variation}) it is easy to guess that 
whether the limiting processes will be deterministic or stochastic
is determined by whether $\alpha$ is zero or nonzero.

It is convenient to record some notation for the 
generator of the diffusion 
limit of a random walk with jump distribution determined by $q_\theta(x,dy)$.

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    This implies that if we define, for $f : \IR^d \to \IR$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    then
    \begin{align}
        \theta \int \left(
            f(y) - f(x)
        \right) q_\theta(x, dy)
    \to \DG f(x) 
        \qquad \text{as } \theta \to \infty .
    \end{align}
    We will also denote the adjoint of $\DG$ by
    \begin{align}
	    \nonumber
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    \label{drift in adjoint}
        % &=
        % \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align}
\end{definition}

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \smooth{\gamma} \eta(x))
    \int r(y, \smooth{r} \eta(y)) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$$
    q_\theta^\mathfrak{m}(x,\eta,  dy)
    :=
    \frac{
        r(y, \smooth{r} \eta(y)) q_\theta(x, dy)
    }{
        \int r(z, \smooth{r} \eta(z)) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \smooth{r} \eta(\cdot)) \right](x)
    }{
        r(x, \smooth{r} \eta^{N}(x))
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}
In order to pass to a scaling limit, we will need to impose some 
conditions on the parameters of our model.

\begin{assumptions}%[Assumptions]
\label{def:model_setup}
We shall make the following assumptions on the parameters of our model.

{\bf Dispersal generator:}
We assume that
\begin{enumerate}
    \item $\meanq(x)$ and $\covq(x)$ are
        $\alpha$-H\"older continuous for some $\alpha \in (0, 1]$
        and uniformly bounded
        in each component;
    % \item $\meanq(x)$ and $\covq(x)$ are $\alpha$-H\"older continuous in each component, for some $\alpha \in (0,1]$;
    \item the operator $\DG$ is uniformly strictly elliptic.
\end{enumerate}

{\bf Reproduction parameters:}
We assume that 
\begin{enumerate}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
    derivatives in both arguments;
    \label{r_gamma_derivs_condition}
\item $\gamma(x,m)$ is uniformly bounded;
    \label{gamma_bounded_condition}
\item For each $f \in C_b^2(\IR^d)$, there is a $C_f$ such that
    $$ |\gamma(x, \smooth{\gamma} \eta(x)) \theta \int 
    (r(y, \smooth{r} \eta(y)) f(y) - r(x, \smooth{r} \eta(x)) f(x))
    q_\theta(x, dy)| \le C_f $$
    for all $x \in \IR^d$ and $\eta \in \measures$.
        \label{gamma_B_condition}
\end{enumerate}

To keep expressions manageable, we shall also assume that 
\[
\mu_\theta(x) = r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N),
\]
that is, this expression is non-negative so that there is no need to take the maximum
	with zero in~(\ref{eqn:mu_defn}). (This is anyway true for 
	sufficiently large $\theta$.)
\end{assumptions}

We give two concrete situations in which the fourth condition is satisfied.

\begin{lemma}
    \label{lem:conditions on r}
    Assume that Conditions~\ref{def:model_setup} are satisfied,
    except for condition~\ref{gamma_B_condition}.
    If either
\begin{enumerate}
\item
\label{control through r} 
    $|\partial_{x_i}r(x,\smooth{r}\eta(x))|$ and 
    $|\partial_{x_ix_j}r(x,\smooth{r}\eta(x))|$
    are uniformly bounded for $x\in\IR^d$, $\eta\in\measures$;
\item 
\label{control through gamma}
    or, $m^2\gamma(x,m)$ is uniformly bounded and there exists $C<\infty$
    such that for $\theta$ sufficiently large, and all $x\in\IR^d$, $\eta\in\measures$,
    \[
    \theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)q_\theta(x,dy)\leq C\smooth{\gamma}\eta(x),
    \]
    and
    \[\theta\int\big(\smooth{r}\eta(y)-\smooth{r}\eta(x)\big)^2q_\theta(x,dy)
    \leq C(\smooth{\gamma}\eta(x))^2 ,
    \]
\end{enumerate} 
    then condition~\ref{gamma_B_condition} is also satisfied.
\end{lemma}

The purpose of the conditions that we have placed on the reproduction parameters is to 
ensure that the net per capita reproduction rate (before time scaling)
is order $1/\theta$. As remarked above, 
because of the non-local reproduction mechanism, it no longer suffices to assume that 
$r(x,\smooth{r}\eta(x))\gamma(x, \smooth{\gamma}\eta(x))-\mu_\theta(x)$ is of order $1/\theta$.
An instructive example is the case where $\gamma \equiv 1$ and $F \equiv 0$, so that $\mu_\theta = r$,
and $\eta = \delta_x$
(i.e., the population has all individuals at a single location)
in which case the mean rate of change of the total population size
is $\int (r(y, N \smooth{r}(y) ) - r(x, N \smooth{r}(x))) q_\theta(x, dy)$;
the first condition of Lemma~\ref{lem:conditions on r} would ensure this is of order $1/\theta$.

If $r(x,m)$ is independent of $m$, then the conditions are easy to 
satisfy; they just require some regularity of $r$ as a function of $x$. 
Condition~\ref{control through r} is also satisfied if for example
$\|\nabla\rho_r\|\leq C\rho_r$ 
and $m\partial_mr(x,m)$, $m^2\partial_{mm}r(x,m)$ are bounded.
This is the case, for instance, if $\rho_r$ decays exponentially.
On the other hand, it might seem more natural to take $\rho_r$ to be a Gaussian 
density with parameter $\sigma_r$, say. Then, 
as we check in Lemma~\ref{lem:gamma_bound},
Condition~\ref{control through gamma} is satisfied if $\rho_\gamma$ is also
Gaussian with parameter $\sigma_\gamma$ and $\sigma_\gamma>\sigma_r$. For large enough
$\theta$, this condition guarantees that 
$\sigma_r+1/\theta <\sigma_\gamma$, so that the establishment probability of a juvenile
is controlled by individuals that are already `felt' by the fecundity kernel $\rho_\gamma$
at the location of their parent.

% % % % % % % % % % % % % % %
\subsection{Scaling limits of the population process}

Our main results depend on two dichotomies:
Is the limiting process deterministic or a (generalized) superprocess?
And, are interactions pointwise in the limit or nonlocal?
Below we have results for deterministic limits with pointwise and nonlocal interactions,
and for superprocess limits with nonlocal interactions.
See Figure \ref{fig:super_vs_det_2d} for snapshots of the population
from direct simulation of the process using SLiM \citep{haller_slim_2019}.

\begin{figure}
    \begin{center}
        \includegraphics{figures/ex1a/fkpp_123.density.3.0}
        \includegraphics{figures/ex1b/fkpp_123.density.3.0}
    \end{center}
    \caption{
        Snapshots of two simulations, with small $\alpha=\theta/N$ (left) and large 
	$\alpha =\theta/N$ (right).
        Simulations are run with a Fisher-KPP-like parameterization:
        birth and establishment are constant, while death increases linearly with density,
        at slope $1/\theta$.
        Left: $\alpha=0.1$. Right: $\alpha=10$.
        Other parameters were the same:
        dispersal and interactions distance were set to 1,
        and the equilibrium density is 10 individuals per unit area.
        Blue curves are contours from a kernel density estimate.
        \label{fig:super_vs_det_2d}
    }
\end{figure}

\paragraph{Scaling limits with nonlocal interactions:}
Recall that the process $(\eta_t^N)_{t\geq 0}$ takes its values in the space
${\cal D}([0,\infty),\measures)$ of c\`adl\`ag
paths on $\measures$. We endow $\measures$ with the topology of weak convergence
and 
${\cal D}([0,\infty),\measures)$ with the Skorohod topology.
A sequence of processes taking values in 
${\cal D}([0,\infty),\measures)$ is said to be tight if the corresponding
sequence of distributions
is tight, so that any infinite sequence has a weakly convergent subsequence.
Our first result establishes tightness of our rescaled population processes in 
the case in which interactions remain nonlocal under the scaling, and 
characterises limit points as solutions to a martingale problem.

\begin{theorem} \label{thm:nonlocal_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction}
	and assume that as $N \to \infty$, $\theta(N) \to \infty$
	in such a way that $\theta(N)/N \to \alpha$.
    (However, the kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
    remain fixed.) 
Suppose that Assumptions~\ref{def:model_setup} hold and, 
further, that $\{\eta_0^N\}_{N\geq 1}$ is a sequence of 
purely atomic measures, with $\eta_0^N$ comprised of atoms of mass $1/N$, which is
tight 
in $\measures$ and such that there exists $\lambda>0$ for 
which $\langle e^{\lambda\|x\|}, \eta_0^N(dx)\rangle<C<\infty$
with $C$ independent of $N$.
    Then the sequence of processes
    $(\eta^N_t)_{t \geq 0}$ is tight, and for any limit point
    $(\eta_t)_{t \geq 0}$, for every $f \in C^\infty_b(\IR^d)$,
    \begin{align} \label{eqn:limiting_mgale_problem}
        \begin{split}
        M_t(f)
            &:=
            \langle f(x), \eta_t(dx) \rangle
            -
            \langle f(x), \eta_0(dx) \rangle
            \\ & \qquad
            -
            \int_0^t \big\langle
                \gamma(x, \eta_s)
                \mathcal{B}\left(
                    f(\cdot) r(\cdot, \eta_s)
                \right)(x)
                %\\ &\qquad \qquad \qquad {}
                +
                f(x)
                F(x, \eta_s),
                \eta_{s}(dx)
            \big\rangle ds
        \end{split}
    \end{align}
    is a martingale (with respect to the natural filtration),
    with quadratic variation
    \begin{align} \label{eqn:limiting_mgale_variation}
        [ M(f) ]_t
        =
        \alpha
        \int_0^t
        \big\langle
            %\left(
	    2\gamma\left( x, \eta_{s} \right)
            r\left(x, \eta_{s} \right)% + \mu\left(x, \eta_{s} \right)\right)
            f^2(x),
            \eta_{s} (dx)
        \big\rangle ds. 
    \end{align}
    Furthermore, if $\alpha = 0$ the limit is deterministic.
\end{theorem}

Recall when interpreting~\eqref{eqn:limiting_mgale_problem}
that, for instance, $r(x, \eta_s) = r(x, \smooth{r} \eta_s(x))$,
and so $\DG(fr)(x) = \DG(f(\cdot) r(\cdot, \smooth{r} \eta_s(\cdot)))(x)$.

Theorem \ref{thm:nonlocal_convergence} provides tightness of the
rescaled processes. If the limit points are unique, then this
is enough to guarantee convergence.

\begin{corollary} \label{cor:superprocess_uniqueness}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    if the martingale problem
    defined by equations~\eqref{eqn:limiting_mgale_problem} 
and~\eqref{eqn:limiting_mgale_variation}
    has a unique solution,
    then $(\eta^N_t)_{t \ge 0}$ converges weakly
to that solution
    as $N \to \infty$.
\end{corollary}
When $\alpha>0$, the limit points can be thought of as interacting superprocesses. For
example, when $r$ and $\gamma$ are constant, and 
$F(x,\eta_s)=1-\smooth{F} \eta_s(x)$,
we recover a superprocess with nonlinear death rates corresponding to logistic growth
\citep{etheridge:2004}
that is a continuous limit of the Bolker-Pacala model \citep{bolker/pacala:1997,bolker/pacala:1999}.
We are not aware of a general result to determine when we will have uniqueness of 
solutions to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}
when $\alpha>0$.
However, the Dawson--Girsanov transform tells us that we have uniqueness in this special 
case of the superprocess with nonlinear death rates, and 
Perkins stochastic calculus (and its adaptation to a 
lookdown setting) provides uniqueness for cases with interactions in the 
dispersal mechanism of
the superprocess. We refer 
to~\cite{dawson:1993, perkins:1992}, and \cite{donnelly/kurtz:1999} 
for approaches to showing that the martingale problem is well-posed in the case $\alpha>0$.

Recall the following notion of solution to a PDE.
\begin{definition}[Weak solutions]
    \label{defn:weak_solutions}
    We say that $(\eta_t)_{t \ge 0}$, with $\eta_t \in \measures$,
    is a \emph{weak solution} to the PDE
    \begin{equation} \label{eqn:nonlocal_pde}
        \partial_t \varphi = r \DG^*(\gamma \varphi) + \varphi F
    \end{equation}
	(where $r$, $\gamma$ and $F$ can all be functions of $\varphi$)
    if, for all $f \in C_b^\infty(\IR^d)$,
    \[
        \frac{d}{dt} \langle f, \eta_t \rangle
        =
        \langle
	    \gamma \DG(rf) + f F,
            \eta_t
        \rangle .
    \]
\end{definition}
The notation $\varphi$ is meant to be suggestive of a density,
and recall that equation~\eqref{eqn:nonlocal_pde} has made dependencies on $x$ and $\varphi$ implicit;
written out more explicitly, \eqref{eqn:nonlocal_pde} is
\[
    \partial_t \varphi_t(x)
    =
    r\left(x, \smooth{r} \varphi_t(x) \right)
    \DG^* \left[
        \varphi_t(\cdot)
        \gamma\big( x, \smooth{\gamma} \varphi_t(\cdot) \big)
    \right](x)
    +
    \varphi_t(x)
    F\left(x, \smooth{F} \varphi_t(x) \right)
    .
\]
Because Theorem~\ref{thm:nonlocal_convergence} only tells us about 
weak convergence, in the case $\alpha=0$ we can only deduce that 
any limit point $\eta_t$ is a weak solution to this nonlocal PDE.

Specialising the results of 
\citet{KURTZ1999103} to the deterministic setting
provides general conditions under which we
have existence and uniqueness of solutions to~(\ref{eqn:nonlocal_pde}) which
have an $L^2$-density with respect to Lebesgue measure. 
Recall that the Wasserstein metric, defined by 
\[\rho (\nu_1,\nu_2)=\sup\Big\{\Big|\int fd\nu_1-\int fd\nu_2\Big|:\sup_x|f(x
)|\leq 1,|f(x)-f(y)|\leq \|x-y\|\Big\},\]
determines the topology of weak convergence on $\measures$.

We write 
$r(x,\eta) \gamma(x,\eta) C(x) = J(x,\eta) J(x,\eta)^T$, and 
$\beta(x,\eta) = r(x,\eta) \gamma(x,\eta) \big(\meanq(x)+2C(x)\nabla\log r(x,\eta)\big)$
(quantities that will appear in Proposition~\ref{prop:limiting_construction}).
If $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
that
\begin{equation}
\label{lipschitz for kurtz xiong}
|J (x_1, \nu_1)-J (x_2, \nu_2)|,|\beta (x_1,\nu_1)-\beta 
	(x_2,\nu_2)|, |F(x_1,\nu_1)-F(x_2,\nu_2)|
	%,|b(x_1,\nu_1)-b(x_2,\nu_2)|
	\leq C(\|x_1-x_2\|+\rho (\nu_ 1,\nu_2))
\end{equation}
for some $C>0$,   
the methods of \cite{KURTZ1999103} show that %under the conditions
%of Theorem~\ref{thm:kurtzxiong}, 
if the initial condition $\eta_0$ for our 
population process has
an $L^2$ density, then so does $\eta_t$ for $t>0$. 
Although the necessary estimates (for which we refer to the original paper)
are highly nontrivial, 
the idea of the proof is simple. Take a solution to the
equation and use it to calculate the coefficients $r$, $\gamma$ and $F$ that depend
on local population density. Then $\eta$ solves the 
{\em linear} equation obtained by regarding those values of $r$, $\gamma$ and $F$ as given.
It remains to prove that the solution to the linear equation has a density
which is achieved by obtaining
$L^2$ bounds on its convolution with the heat semigroup at time $\delta$ and letting 
$\delta\to 0$.
We also have the following uniqueness result.
\begin{theorem}[Special case of \cite{KURTZ1999103}, Theorem~3.5]
	\label{thm:kurtzxiong}
Suppose  $J$, $\beta$, and $F$ are bounded and Lipschitz in the sense 
of~(\ref{lipschitz for kurtz xiong}).
If $\eta_0$ has an $L^2({\mathbb R}^d)$-density, then there 
exists a unique $L^2({\mathbb R}^d)$-valued  
	solution of (\ref{eqn:nonlocal_pde}) in the sense of 
	Definition~\ref{defn:weak_solutions}.
\end{theorem}
\begin{remark}
\cite{KURTZ1999103} considers an infinite system of stochastic differential equations
for the locations and weights of a collection of particles that interact through their 
weighted empirical measure, which is shown to be the unique solution to a stochastic
	PDE. As we shall see through our lookdown representation in 
Section~\ref{sec:lookdown}, the
	solution to our
	deterministic equation can be seen as the empirical measure of a 
	countable number of particles (all with the same weight)
which, in the notation above, evolve according
	to 
	\[X(t)=X(0)+\int_0^t\beta\big(X(s), \eta_s\big)ds
+\int_0^tJ\big(X(s), \eta_s\big)dW(s)\]
	(with an independent Brownian motion $W$ for each particle).
\end{remark}

% PLR: perhaps we don't need this?
% To understand the form of equation~(\ref{eqn:nonlocal_pde}) a little better, 
% and, in particular, why we see $\DG^*$ rather than $\DG$ in the equation for the 
% limiting population density, we consider the two terms that contribute to the drift in our 
% expression~(\ref{drift in adjoint})
% for $\DG^*$. First consider a special case in which dispersal has a mean drift 
% to the right, so that the drift term in $\DG$ is $d/dx$, and the effect of the drift is 
% just to shift the population profile to the right. If we are at a point where the 
% slope of the population density is positive, then the shift will {\em reduce} the
% population density at the point; the change in population density is proportional
% to $-d/dx$. Similarly, if dispersal tends to move offspring away from a region, that
% is if $\grad\cdot\meanq >0$, 
% then it will push down the population density, hence the term $-\grad\cdot\meanq$ in
% equation~(\ref{eqn:nonlocal_pde}).

\paragraph{Two-step convergence to PDE:}
Although the coefficients at $x$ in \eqref{eqn:nonlocal_pde} are nonlocal,
we can choose our kernels $\rho_\gamma$, $\rho_r$, and $\rho_F$
in such a way that they depend only on the population in a region close to $x$,
and so we expect that under rather general conditions
solutions of the nonlocal PDE will be close to the corresponding 
classical PDE.
The following propositions provide two concrete situations in which this is true.
In the first, the PDE is a reaction diffusion equation, and we borrow
an idea from~\cite{penington:2017} to express the solutions to both the nonlocal
equation and the classical PDE through a Feynman-Kac formula. 
\begin{proposition}
    \label{prop:nonlocal_to_local}
Recall the notation $\rho^\epsilon_F(x)=\rho_F\big(x/\epsilon)/\epsilon^d$.
Assume $\varphi_0\in L^2(\IR)$ is a positive, uniformly Lipschitz, and uniformly bounded function. 
Suppose that $\varphi^\epsilon\in L^2(\IR^d)$ is a weak solution to the equation
\begin{equation}
\label{nonlocalPDEv1} 
%\begin{cases}
\partial_t \varphi^\epsilon = \DG^* \varphi^\epsilon + 
\varphi^\epsilon F(\rho_F^\epsilon*\varphi^\epsilon),  
	%& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi^\epsilon(0,x) = \varphi_0(x) & x \in \mathbb{R}^d;
%\end{cases}
\end{equation}
with initial condition $\varphi_0(\cdot)$,
and that $\varphi$ is a weak solution to the equation
 \begin{equation}
\label{localPDE} 
%\begin{cases}
\partial_t \varphi = \DG^* \varphi + \varphi F(\varphi),  
%	& 
\qquad x \in \mathbb{R}^d,\, t >0, %\\ 
%	\varphi(0,x) = \varphi_0(x) & x \in \mathbb{R}^d.
%\end{cases}
\end{equation}
also with initial condition $\varphi_0(\cdot)$.
%Assume $\varphi_0\in L^2(\IR)$ is a positive, Lipschitz, and uniformly bounded function; 
Suppose further that $F$ is a Lipschitz function which is bounded above, and that
$b(x)$ and $C(x)$, the drift and 
covariance matrix of $\DG$, satisfy the conditions of 
Assumptions~\ref{def:model_setup}.
Then, for all $T>0$ there exists a constant $K=K(T, \Vert \varphi_0 \Vert_\infty)>0$,
and a function $\delta(\epsilon)$ (dependent on $\rho_F$) with $\delta(\epsilon)\to 0$ 
as $\epsilon\to 0$,
such that, for all $0 \leq t \leq T$, and $\epsilon$ small enough,
\[ 
\Vert \varphi_t(\cdot) - \varphi_t^\epsilon(\cdot) \Vert_\infty\leq K\delta(\epsilon). 
\]
In particular, as $\epsilon\to 0$, we have that $\varphi^\epsilon$ converges 
uniformly in compact intervals of time to $\varphi$.
\end{proposition}
\begin{remark}
Note that Theorem~\ref{thm:kurtzxiong} guarantees uniqueness of solutions
to equation~(\ref{nonlocalPDEv1}). 
\end{remark}
Our second example is a nonlocal version of a porous medium equation with
logistic growth. That is, 
we consider non-negative solutions to the equation
\begin{equation}
\label{mollified equation}
\partial_t \psi^\epsilon=
\Delta\left(\psi^\epsilon \, \rho^{\epsilon}_\gamma*\psi^\epsilon\right)
+\psi^\epsilon\left(1-\rho^{\epsilon}_\gamma*\psi^\epsilon\right).
\end{equation}
The case without the reaction term (and with $\IR^d$ replaced by a torus) is considered 
by~\cite{lions/mas-gallic:2001} who use
it as a basis for a particle method for numerical solution of the porous medium equation.
Of course this does not quite fit into our framework, since in the notation of our 
population models this would necessitate $\gamma(x,m)=\smooth{\epsilon}m$ which is
not bounded. However, this can be overcome by an additional layer of approximation
(c.f.~our numerical experiments of Section~\ref{beyond linear diffusion})
and we do not allow this to detain us here. Existence and uniqueness of solutions 
to~(\ref{mollified equation}) can be obtained using the approach
of~\cite{lions/mas-gallic:2001}.

We should like to prove that as $\epsilon\to 0$ we have
convergence to the solution to the porous medium equation with
logistic growth:
\begin{equation}
\label{PME}
\partial_t\psi=
\Delta\left(\psi^2\right)
+\psi\left(1-\psi\right).
\end{equation}

\begin{notation}
We use $\rightharpoonup$ to denote weak convergence in the sense of analysts;
that is, $\psi^\epsilon\rightharpoonup \psi$ in $L^1$ means 
$\int \psi^\epsilon v d x
\rightarrow\int \psi v d x$ for all $v\in L^\infty$.

We write $L_t^2(H^1)$ for functions for which the $H^1$ norm in
space is in $L^2$ with respect to time, i.e.
$$\int_0^T\int \left\{\psi_t(x)^2+\left\| \nabla \psi_t(x)\right\|^2\right\}
dx dt <\infty,$$
and $C_t(L^1)$
will denote functions for which the $L^1$ norm in space is continuous in time.
\end{notation}

\begin{proposition}
	\label{nonlocalPME to PME}
Suppose that
we can write $\rho_\gamma=\zeta*\check{\zeta}$, where $\check{\zeta}(x)=\zeta(-x)$ and
$\zeta\in\mathcal{S}(\IR^d)$ (the Schwartz space of rapidly decreasing functions).
Furthermore, suppose that %$\psi_0^\epsilon\in L^1(\IR^d)$, 
$\psi_0^\epsilon\geq 0$ is such that
there exist $\lambda\in (0,1)$ and $C\in (0,\infty)$ (independent
of $\epsilon$) such that 
\begin{equation*}
\int\exp(\lambda \|x\|)\psi_0^\epsilon(x)dx<C,\quad\mbox{ and }
\sup_\epsilon\int \psi_0^\epsilon|\log \psi_0^\epsilon|dx<\infty,
\end{equation*}
with
$\psi_0^\epsilon\rightharpoonup \psi_0$ as $\epsilon\to 0$. Then
writing $\psi^\epsilon$ for the solution to~(\ref{mollified equation})
on $[0,T]\times \IR^d$ with
initial condition $\psi_0^\epsilon$,
$\psi^\epsilon\rightharpoonup \psi$ as $\epsilon\to 0$ where
$\psi\in L_t^2(H^1)\cap C_t(L^1)$, $\int \psi|\log \psi| dx<\infty$, and
$\psi$ solves~(\ref{PME}) on $[0,T]\times \IR^d$. 
\end{proposition}
The example that we have in mind for the kernel $\rho_\gamma$ is a Gaussian
kernel.
\begin{remark}
Although it seems hard to formulate an all-encompassing result,
Propositions~\ref{prop:nonlocal_to_local} and~\ref{nonlocalPME to PME} are by no
means exhaustive. When the scaling limit is deterministic, one can 
expect analogous results under rather general conditions. However, when the limit
points are stochastic, they resemble ``nonlinear superprocesses'' and so one cannot
expect a density with respect to Lebesgue measure in $d\geq 2$. It is then not
reasonable to 
expect to be able to make sense of the limit if we scale the kernels in this way.
Moreover, in one dimension, where the classical superprocess does have a density with 
respect to Lebesgue measure, the form of~(\ref{eqn:limiting_mgale_problem})
suggests that even if one can remove the local averaging from $\gamma$, it
will be necessary
to retain averaging of $r$ in order to obtain a well-defined limit. 
\end{remark}

\paragraph{One-step convergence to PDE:}

Theorem \ref{thm:nonlocal_convergence}, combined with Proposition \ref{prop:nonlocal_to_local},
implies that we can take the limit $N \to \infty$
followed by the limit $\epsilon \to 0$
to obtain solutions to the PDE \eqref{localPDE}.
However, it is of substantial interest to know whether
we can take those two limits simultaneously.
The general case seems difficult,
but we prove such ``diagonal'' convergence in the following situation.

\begin{theorem}[Convergence to a PDE]
    \label{thm:local_convergence}
    Let $(\eta^N_t)_{t \geq 0}$
    be as defined in Definition~\ref{defn:mgale_construction} with
$r(x,m)\equiv 1\equiv \gamma(r,m)$, $F(x,m)\equiv F(m)$, 
$\rho^\epsilon_F$ a symmetric Gaussian density with variance
parameter $\epsilon^2$,
and $\DG=\Delta$. 
Further suppose that $F(m)$ is a polynomial with $F(m)\ind_{m\geq 0}$ bounded above.
We assume that $\langle 1,\eta_0\rangle$ is bounded, and that for all 
$x\in\IR^d$ and $k\in \IN$,
	\[
		\limsup_{\epsilon\to 0} 
		\IE\big[\rho_F^\epsilon*\eta_0(x)^k\big]<\infty,
	\]
	and 
\[
	\int \IE\big[\rho_F^\epsilon*\eta_0(x)^k\big] dx<\infty.
\]
Finally assume that $N \to \infty$, $\theta \to \infty$
and $\epsilon \to 0$ in such a way that
\begin{equation}
\frac{1}{\theta\epsilon^2}+\frac{\theta}{N\epsilon^d}\to 0.
\end{equation}
Then the sequence of 
${\cal D}([0,\infty), \measures)$-valued
stochastic processes $\big(\rho_F^\epsilon*\eta^N_t(x)dx\big)_{t \ge 0}$
converges weakly
to a measure-valued process with a density $\varphi(t, x)$
that solves
\begin{align}
	\label{target equation in local convergence}
        \partial_t \varphi(t, x)=\Delta \varphi (t,x) +\varphi(t,x) F(\varphi(t,x)).
\end{align}
\end{theorem}

\begin{remark}
In fact,
our proof goes through without significant change under the conditions that
$F(m)\ind_{m\geq 0}$ is bounded above (but not necessarily below), and that
for all $m,n\in[0,\infty)$
\[
|F(m)|\leq\sum_{j=1}^ka_jm^j, \quad\mbox{ and }
|F(n)-F(m)|
\leq|n-m|\sum_{j=1}^{k'}b_j\Big(n^j+m^j\Big),
\]
for some non-negative constants $\{a_j\}_{j=0}^k$, $\{b_j\}_{j=0}^{k'}$.
We take $F$ to be polynomial to somewhat simplify notation in the proof.
\end{remark}

% % % % % % % % % % % %
\subsection{Ancestral lineages in the scaling limit}
\label{sec:ancestral lineages}

Now that we have established what we can say about how 
population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t^N)_{t \ge 0}$,
begun at a spatial location $L_0^N = x$
where there is a focal individual in the present day
can be obtained by, for each time $t$,
setting $L_t^N$ to be the spatial location of the individual alive at time $t$
before the present from
whom the focal individual is descended.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$, since individuals are born one at a time, 
one could construct the distribution of a lineage $(L_t^N)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to formally retain such information when we 
pass to the scaling limit.

The \emph{lookdown construction} in Section~\ref{sec:lookdown}
will enable us to recover information about ancestry in the infinite
population limit.
Roughly speaking,
each particle is assigned a unique ``level'' from $[0,\infty)$
that functions as a label and thus allows reconstruction of lineages.
The key to the approach is that levels are assigned in such a way as to be exchangeable, 
so that sampling a finite number, $k$ say, of
individuals from a given 
region is equivalent to looking at the individuals in that region with the $k$ lowest levels.
Moreover, as we pass to 
the infinite population (and therefore infinite level) limit,
the collection of (individual, level) pairs 
converges, Theorem~\ref{thm:lookdown_convergence}.
See~\citet{etheridge/kurtz:2019} for an introduction to these ideas.
In particular, even in the infinite
population limit, we can sample an individual
from a region (it will be the individual in that region with the lowest level) 
and trace its line of descent. 
This will, in principle, allow us to calculate, for each $y\in\IR^d$, the proportion of
the population at location $x$ in the present day population that are descended from 
a parent who was at location $y$ at time $t$ in the past. To make
sense of this in our framework, in Section~\ref{sec:lineages_proof}, 
we justify a weak reformulation of this idea.

We are interested in two questions.
First, when is the motion of an ancestral 
lineage, given complete knowledge of the population process, a well-defined process?
In other words, is knowledge of the process $(\eta_t)_{t=0}^T$ that records 
numbers of individuals
but not their ancestry sufficient to define the distribution of $(L_t)_{t=0}^T$?
Second, does the process have a tractable description?

We focus on the simplest situation in which the population process is deterministic.
However, the results here apply when the population process solves 
either a nonlocal or a classical PDE.
There will be no 
coalescence of ancestral lineages in the deterministic limit,
but understanding motion of single lineages is useful in practice,
and our results can
be seen as a first step towards understanding genealogies for very high population 
densities. 
\begin{definition}[Ancestral lineage] 
\label{def:lineage_generator}
    Let $(\varphi_t(x))_{0 \le t \le T}$
    denote the density of the scaling limit of our population model, 
solving \eqref{eqn:nonlocal_pde},
    and let $y$ be a point with $\varphi_T(y) > 0$.
    We define $(L_s)_{s=0}^T$,
    the ancestral lineage of an individual sampled from the
population at $y$ at time $T$, by setting $L_s$
    to be the position of the unique ancestor of that individual at time $T - s$.
    We define
    $(Q_s)_{s \geq 0}$
    to be the time inhomogeneous semigroup satisfying
    \begin{align*}
        Q_s f(y) := \IE_y[ f(L_s) ] .
    \end{align*}
\end{definition}

Our next result identifies the ancestral lineage as a diffusion
by characterizing its generator.

\begin{theorem} \label{thm:lineages}
    For $\varphi: \IR^d \to \IR$, define
    \begin{align}
        \label{eqn:lineage_generator}
        \Lgen_\varphi f
        &=
        \frac{r}{\varphi}
        \left[
            \DG^*(\gamma \varphi f) 
            - f \DG^* (\gamma \varphi)
        \right] \\
        &= \label{eqn:lineage_generator2}
        r\gamma
        \left[
            \sum_{ij} \covq_{ij} \partial_{x_ix_j} f
            + \sum_j \vec{m}_j \partial_{x_j} f
        \right] ,
    \end{align}
    where $\vec{m}$ is the vector
    $$
    \vec{m}_j
    =
    2 \sum_i C_{ij} \partial_{x_i} \log(\gamma \varphi)
    + 2 \sum_i \partial_{x_i} C_{ij}
    - \meanq_j .
    $$
    Then the generator of the semigroup $Q_s$
    of Definition~\ref{def:lineage_generator} is given by
	$\partial_sQ_sf(y)= \Lgen_{\varphi_{T-s}}Q_sf(y).$ 
\end{theorem}
\begin{remark}
As usual,
to make the generator readable, we've written it in concise notation,
omitting the dependencies on location and population density,
which itself changes with time.
When interpreting this,
remember that everything depends on location and density at that location and time --
for instance, ``$r$'' is actually $r(x, \varphi(x))$ (in the classical case),
or $r(x, \smooth{r} \eta(x))$ (in the nonlocal case).

Moreover, we haven't proved any regularity of the population density process $\varphi$,
so, as written, the generator~(\ref{eqn:lineage_generator})
may not make sense. Instead, it should be interpreted in a weak sense which is made precise
in Section~\ref{sec:lineages_proof}.
\end{remark}

\begin{corollary} \label{cor:lineages_simple}
    In addition to the assumptions of Theorem~\ref{thm:lineages},
    if the covariance of the dispersal process is isotropic
    (i.e., $\covq = \sigma^2 I$),
    then
    \begin{equation}
        \Lgen_\varphi f
        =
        r \gamma
        \left(
            \sigma^2 \Delta f
            +
            \left(
                2 \sigma^2 \grad \log(\gamma \varphi)
                - \meanq
            \right)
            \cdot \grad f
        \right) .
    \end{equation}
    (However, $\meanq$ can still depend on location.)
\end{corollary}

In other words, 
the lineage behaves as a diffusion driven by Brownian motion 
run at speed $\sigma^2$ multiplied by the local per-capita production of 
mature offspring ($r \gamma$)
in a potential tilted by migration bias
($(\varphi_s \gamma)^2 \exp(-\meanq \cdot x / \sigma^2)$).
In particular, lineages are drawn to regions of high fecundity (production of juveniles), but their
speed is determined by the rate of production of mature offspring.
This can be compared to Remark~\ref{ancestral lineages: first guess}.

\begin{corollary} \label{cor:stationary_dist}
    In addition to the assumptions of Corollary~\ref{cor:lineages_simple},
    if the population process is stationary (so $\varphi_t \equiv \varphi$),
    and $\meanq(x) = \grad h(x)$ for some function $h$,
    then $Y$ is reversible with respect to
    \begin{align}
        \pi(x)
        =
        \frac{\gamma}{r} \varphi(x)^2 e^{-h(x) / \sigma^2} .
    \end{align}
\end{corollary}

\comment{I'm a bit surprised by the factor of 2 (that manifests as $\varphi^2$),
    but I've checked where it comes from pretty carefully, and don't see anything wrong.}
Long-term fitness of an individual
is proportional to the fraction of lineages from the distant future that pass through the individual,
and hence the total long-term fitness at a location is proportional to
the stationary distribution of $Y$ there, if it exists.
Therefore,
if $\pi$ is integrable then the per-capita long-term fitness of an individual at $x$
is proportional to $\pi(x) / \varphi(x)$.

\begin{corollary} \label{cor:wavefront}
    In addition to the assumptions of Corollary~\ref{cor:lineages_simple},
    suppose that the population process is described by a travelling wave with 
velocity $\wavespeed$,
    i.e., the population has density
    $\varphi(t, x) = w(x - t \wavespeed)$
    where $w$ solves
    \begin{align*}
        r \DG^* (\gamma w) + w F + \wavespeed \cdot \grad w = 
r\sigma^2\Delta (\gamma w) -\meanq \cdot\grad (\gamma w) +\wavespeed \cdot \grad w =0 .
    \end{align*}
    Then the semigroup $Q_s$
    of the motion of a lineages in this population 
    is time-homogeneous with generator
    \begin{align}
\label{Lgen}
        \Lgen f
        &=
        \sigma^2 r \gamma
        \left(
            \Delta f
            +
            2 \grad \log (\gamma w)
            \cdot \grad f
        \right)
        + (\wavespeed - r\gamma \meanq) \cdot \grad f .
    \end{align}
\end{corollary}

%%%%%%%%%%%%%%%%%%%%%%
\section{Examples and applications}
\label{sec:applications}

We now turn to some examples and applications.

% % % % % % % % % % % %
\subsection{Beyond linear diffusion}
\label{beyond linear diffusion}

\comment{Maybe check for more recent references for paragraph below.}
Equation~(\ref{eqn:nonlocal_pde}) is a nonlocal version of a reaction diffusion equation with nonlinear
diffusion; in other words the diffusivity of the population depends on the population
density. Passing to the classical limit, we recover equations
like~(\ref{PME}).
Such equations are widely used in a number of contexts in biology in which
it is clear that motility within a population varies with population density.
For example, density dependent dispersal is a common feature in spatial
models in ecology, eukaryotic cell biology, and avascular tumour growth;
see~\cite{sherratt:2010} and references therein for further discussion. 
In particular, it has been suggested as a model for
the expansion of a certain type of bacteria % of the type {\em Paenbacillus dendritiformis}
on a thin layer of agar in a petri dish 
\citep{cohen/golding/kozlovsky/benjacob/ron:1999}. 
We shall pay particular attention to the case in which the equation can be 
thought of as modelling the density of an expanding population. 
\cite{li/buenzli/simpson:2022} write down an interacting particle
system on a hexagonal
lattice (with at most one individual per site) that incorporates an effect
of local crowding on individual motion. Their equation
corresponds to a reaction-diffusion equation with nonlinear diffusion and a 
bistable reaction term of the form~(\ref{Birzu nonlinearity}) and they
investigate the interplay between nonlinear diffusion and long-term 
survival or extinction of the population. We focus instead on 
the monostable reaction of~(\ref{PME}).

Comparing %\eqref{PME} 
with \eqref{eqn:nonlocal_pde},
we see that to set up a limit in which the population density $\varphi$ follows the 
porous medium equation with logistic growth of~(\ref{PME}),
we need $r=1$,
$\gamma = \varphi$, and $F = 1 - \varphi$.
Consulting equation~\eqref{eqn:mu_defn},
this implies that $\mu_\theta = \max\big(0, (1 + 1/\theta) \varphi - 1/\theta\big)$.
In other words,
establishment is certain
and birth rates increase linearly with population density,
but to compensate, death rates increase slightly faster (also linearly),
as shown in Figure~\ref{fig:pme_waves}.
Alert readers will notice that
the condition from Assumptions~\ref{def:model_setup} 
that $\gamma(x,m)$ be uniformly bounded 
is violated.
This can be corrected by use of a cut-off, and in fact the downwards drift
provided by the logistic control of the population size prevents $m$ from 
getting too big. 
In practice in our simulations we
take discrete time steps of length $dt$ (with $dt$ suitably small),
and have each individual with local density $m$ reproduce and die, respectively, with probabilities
\begin{align*}
    p_\text{birth}(m) = \left(1 - e^{- m dt}\right)
    \qquad 
    p_\text{death}(m) = \left(1 - e^{- (m (1+1/\theta) - 1/\theta) dt} \right) ,
\end{align*}
so that $\gamma(x, m) = p_\text{birth}(m)/dt \approx m$
and
\begin{align*}
    F(x, m) = \theta(\gamma(x, m) - \mu(x, m)) / dt \approx 1 - m .
\end{align*}
Birth and death rates are equal at density $m = 1$,
corresponding to an unscaled density of $N$ individuals per unit area.

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/ex2a/pme_123.steps}
        \includegraphics[width=0.45\textwidth]{figures/ex2b/pme_123.steps}
    \end{center}
    \caption{
        Two panels, showing simulated 1D populations under a
porous medium equation with logistic growth~(\ref{PME}), comparing the simulated
        profile to the analytical solution;
        $\theta/N$ not small on left; small on right.
        % (Note: noisier wave should move slower, although maybe not much since it's pushed.)
        \label{fig:pme_waves}
    }
\end{figure}

In one dimension, equation~(\ref{PME}) has an explicit travelling wave solution
\begin{align} \label{eqn:pme_wave}
    w^P(t, x)
    :=
    \left( 1 - e^{ \frac{1}{2} (x - x_0 - t) } \right)_+ .
\end{align}
Notice that the wave profile has a sharp boundary at $x = x_0 + t$.
There are also travelling wave solutions with $c>1$ \citep{gilding/kersner:2005},
which lack this property.
However, for initial conditions that decay sufficiently rapidly at infinity,
such as one might use in modelling a population invading new territory,
the solution converges to \eqref{eqn:pme_wave} \citep{kamin/rosenau:2004}.
In Figure~\ref{fig:pme_waves} we show simulations of the individual based model
described above. They
display travelling wave solutions
similar to numerical solutions of~(\ref{PME}),
with increasingly good agreement for larger $\theta$ and $N$.
\comment{TODO: does it?}
As expected, noise slows down the wave.
\comment{Not sure that noise will slow things drastically because the 
biggest production of
mass happens behind the wavefront.}

% % % % % % % % % % % %
\subsection{Ancestry in different types of travelling waves}
\label{ancestral lineages for travelling waves}

% \begin{figure}
%     \begin{center}
%         \includegraphics{figures/ex3_fkpp/fkpp_123.lineages}
%         \includegraphics{figures/ex3_fkpp/fkpp_123.lineagehist}
%         \includegraphics{figures/ex3_pme/pme_123.lineages}
%         \includegraphics{figures/ex3_pme/pme_123.lineagehist}
%         \includegraphics{figures/ex3_allen-cahn/allen-cahn_123.lineages}
%         \includegraphics{figures/ex3_allen-cahn/allen-cahn_123.lineagehist}
%     \end{center}
%     \caption{
%         Six panels: one pair for FKPP, one for the porous medium 
% equation with logistic growth~(\ref{PME}), and one for 
% the Allen-Cahn equation~(\ref{eqn:allen_cahn}).
%     \comment{clarify this is in a moving frame of reference}
%         Each pair shows (top) lineage traced back in wavefront,
%         and (bottom) stationary distribution of lineage location,
%         compared to analytical solution for the porous medium case.
%         \label{fig:pme_vs_fkpp}
%     }
% \end{figure}

Although it remains challenging to establish the distribution of genealogical
trees relating individuals sampled from our population model, as described in
the introduction, we can gain some insight by investigating the motion of a
single ancestral lineage. Here we do that in the context of a one-dimensional
population expanding into new territory as a travelling wave. 
We focus on three cases in which we have 
explicit information about the shape of the travelling wave profile:
the Fisher-KPP equation,
a special case of the Allen-Cahn equation with a bistable nonlinearity,
and the porous media equation with logistic growth, equation~(\ref{PME}).
We work here in one dimension,
and take $\sigma^2 = 1$ and $\meanq = 0$.

%%%%%%%%%%

%%%%%%%%%%
\paragraph{Fisher-KPP equation:}
Consider the classical Fisher-KPP equation,
\begin{align} \label{eqn:fkpp}
    \partial_t\varphi = \partial_{xx}\varphi + \varphi (1-\varphi) .
\end{align}
Even though we do not have an explicit formula for the wave shape in this case,
our methods provide information about ancestral lineages.
The equation has non-negative travelling wave solutions of 
speed $c$ for all $c \geq 2$, 
but, started from any compact perturbation of a Heaviside function, the 
solution will converge to the profile $w^F$ with the minimal wavespeed, $c=2$
\citep{kolmogorov/petrovsky/piscounov:1937,fife1977approach,bramson:1983}.
No matter what initial condition,
for any $t>0$ the support of the 
solution will be the whole real line. 
In this case, we must have $r = \gamma = 1$,
and $F(x, m) = 1 - m$ so $\mu_\theta(x, m) = 1 + (m-1)/\theta$.
By Corollary~\ref{cor:wavefront},
the generator of the motion of an ancestral lineage is
\begin{equation} \label{eqn:fkpp_generator}
    \Lgen_1  f
    =
    \partial_{xx}f + 2 \frac{\partial_xw^F}{w^F} \partial_x f + 2 \partial_xf .
\end{equation}
Near the tip of the wave (for $x$ large), $w^F(x) \sim e^{-x}$,
so \eqref{eqn:fkpp_generator} implies that
the motion of a lineage is close to unbiased Brownian motion.
On the other hand, in the ``bulk'', %for $x < 0$ 
a lineage behaves approximately as
Brownian motion with drift at rate two to the right.
This implies that
ancestral lineages are pushed into the tip of the wave,
and there is no stationary distribution,
so that long-term dynamics of genetic inheritance
depend on the part of the wave not well-approximated by a smooth profile,
in agreement with the results described in Section~\ref{introduction}.

%%%%%%%%%%
\paragraph{Allen-Cahn equation:}
Now take the Allen-Cahn equation:
\begin{align} \label{eqn:allen_cahn}
    \partial_t\varphi = \partial_{xx}\varphi + \varphi(1-\varphi)(2\varphi-1+s),
\end{align}
for a given $s \in (0,2)$.
Once again we have taken $r=\gamma=1$, but now the reaction term is bistable.
This equation can be used to model the motion of so-called 
hybrid zones in population genetics; see, for example,
\cite{barton:1979, gooding:2018, etheridge/gooding/letter:2022}.
This equation has an explicit travelling wave solution with speed $s$
and shape
\[ w^A(x) = (1+e^{x})^{-1} , \]
i.e., $\phi_t(x)=w^A(x-st)$ solves~(\ref{eqn:allen_cahn}).
Substituting $w^A$ in place of $w^F$ in~(\ref{eqn:fkpp_generator}),
we find that the generator of an ancestral lineage is now, 
\begin{align*}
\Lgen_2 f
    &=
    \partial_{xx}f
    + 
    2 \frac{\partial_xw^A}{w^A} \partial_xf
    +
    s \partial_x f\\
    \qquad &=
    \partial_{xx}f
    -
    2 \frac{e^x}{1+e^x} \partial_xf 
    + 
    s \partial_xf,
\end{align*}
so lineages in the tip are pushed leftwards into the bulk of the 
wave at a rate $-2e^x/(1+e^{x})$.
The strength of this drift decreases exponentially as $x\to -\infty$.
The density of the speed measure for this diffusion is
$$
    m(x) \propto e^{sx}(1+e^x)^{-2},
$$
which is integrable, and so determines the unique stationary distribution.
Thus the position of the ancestral lineage relative to the wavefront will converge to a stationary 
distribution which is maximised away from the extreme tip of the wave.
This is consistent with~\cite{etheridge/penington:2022}, who consider an analogous
stochastic population model, although the stronger result there (that the genealogy
of a sample from behind the wavefront is approximately a Kingman coalecsent) requires
the stronger condition $s<1$.

% % % % % % % % % % % % % % % %

\paragraph{Porous Medium equation with logistic growth:}
Finally, consider equation~(\ref{PME}). Setting $x_0=0$ (for definiteness) and substituting
the form of $w^P$ from equation~\eqref{eqn:pme_wave}
into Corollary~\ref{cor:wavefront},
with $c=1$,
$\gamma(x, m) = m$,
$r(x,w) = 1$,
and $F(x, m) = (1 - m)$,
the generator of the diffusion governing the position of the ancestral lineage 
is, for $x < 0$,
%with $f_x$ denoting the spatial derivative of $f$,
\begin{align*}
    \Lgen_3 f
    &=
        w^P
        \left(
        \partial_{xx} f
         +
         2 \frac{\partial_x((w^P)^2)}{(w^P)^2} \partial_xf
        \right)
        + \partial_xf \\
    &=
        \left(1 - e^{\frac{1}{2} x} \right)
        \partial_{xx}f
        -
        2 e^{\frac{1}{2} x} \partial_xf
        +
        \partial_xf .
\end{align*}
The speed measure corresponding to this diffusion has density
%again for $\xi < 0$,
\begin{align*}
    m(\xi)
    &\propto
        \frac{ 1 }{ 2 (1 - e^{\xi/2}) }
        \exp\left(
            \int_\eta^\xi \left\{
                1 - \frac{e^{x/2}}{1 - e^{x/2}}
            \right\} dx
        \right) \\
    &\propto
        e^\xi\left(1-e^{\xi/2}\right),
        \quad
        \text{for } \xi < 0 ,
\end{align*}
which is integrable and so when suitably normalised gives the unique stationary distribution.
Notice that even though we have the same reaction term as in the Fisher-KPP equation, with this 
form of nonlinear diffusion, at stationarity
the lineage will typically be significantly behind the front, 
suggesting a different genealogy.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Clumping from nonlocal interactions}

Simulating these processes and exploring parameter space,
one sooner or later comes upon a strange observation:
with certain parameter combinations, the population spontaneously forms a regular grid
of more or less dicrete patches, separated by areas with nearly no individuals.
The phenomenon is discussed in Section~16.10 of \citet{haller2022evolutionary},
and has been described in similar models, e.g.,
by \citet{britton1990spatial,sasaki1997clumped,hernandezgarcia2004clustering,young2001reproductive,berestycki2009nonlocal}.
Heuristically, this can occur if density-dependent effects of individuals
extend sufficiently farther than the typical dispersal distance:
the patches are stable because new offspring landing between two clumps
effectively compete with \emph{both} neighboring clumps,
while individuals within a clump effectively compete with only one.

The implication here is that the equilibrium solution to~\eqref{general deterministic limit}
may not be unique,
and a stable solution may have oscillations on a scale determined by the interaction distance.
We can see this easily in the simple case that
$\DG = \sigma^2 \Delta$,
the interaction kernels $\rho_r$, $\rho_\gamma$, and $\rho_F$
are all equal to $p_{\epsilon^2}$ (the heat kernel at time $\epsilon^2$),
and all parameters are spatially homogeneous,
so that $r(x, \eta) = r(\smooth{r}\eta(x))$, and similarly for $\gamma$ and $F$.
Let $\varphi_0$ be a value so that $F(\varphi_0) = 0$ and $F'(\varphi_0) < 0$,
so that we would expect $\varphi(x) \equiv \varphi_0$ to be a stable equilibrium solution.
Let us linearize~\eqref{general deterministic limit} about $\varphi_0$:
let $\varphi_t(x) = \varphi_0 + \psi_t(x)$,
and (informally) $r(x) \approx r(\varphi_0) + r'(\varphi_0) \smooth{r} \psi(x)$.
Writing $r_0 = r(\varphi_0)$ and $r'_0 = r'(\varphi_0)$,
and similarly for $\gamma$ and $F$,
\begin{align*}
    \partial_t \psi
    &\approx
    \sigma^2 \varphi_0 r_0 \gamma_0' \Delta \smooth{\gamma} \psi
    + \sigma^2 r_0 \gamma_0 \Delta \psi
    + \varphi_0 F'_0 \smooth{F} \psi .
\end{align*}
Letting $\hat \psi(u) = \int e^{2\pi i u x} \psi(x) dx$,
\begin{align*}
    \partial_t \hat \psi(u)
    &\approx
    \left(
        - \sigma^2 \varphi_0 r_0 \gamma_0' u^2 e^{-\epsilon^2 u^2 /2}
        - \sigma^2 r_0 \gamma_0 u^2
        + \varphi_0 F'_0 e^{-\epsilon^2 u^2 /2} 
    \right) \hat \psi(u) .
\end{align*}
Since $F'_0 < 0$, the last two terms are negative,
but on the other hand if we take $\gamma_0' < 0$
(i.e., per capita birth rate decreases with density near equilibrium),
then the stability of the constant solution $\varphi \equiv \varphi_0$
depends on the balance between the first and latter two terms.
Writing
$A_1 = - \sigma^2 \varphi_0 r_0 \gamma'_0$,
$A_2 = \sigma^2 r_0 \gamma_0$,
and $A_3 = - \varphi_0 F'_0$
(with signs chosen so each is positive),
and defining
$$ g(v) = A_1 v - A_2 v e^{\epsilon^2 v /2} - A_3 , $$
we have $\partial_t \hat \psi(u) = g(u^2) e^{\epsilon^2 u^2 /2} \psi(u)$.
The term $g(v)$ is maximized at $v_*$ solving
$e^{\epsilon^2 v_*/2} = A_1 / (A_2 (\epsilon^2 v_*/ 2 + 1))$,
which is positive if $A_1 > A_2$,
i.e., if $-\varphi_0 \gamma_0' > \gamma_0$.
If so, then the constant solution is unstable with respect to perturbations
that oscillate with frequency around $\sqrt{v_*}$.
\comment{This seems wrong, since it doesn't involve $\sigma$?
    But at least note the unstable frequencies scale with $\epsilon$.
    We might say more by linearizing the exponential term,
    which gives $v_* = (A_1 - A_2 + A_3 \epsilon^2/2)/(A_1 \epsilon^2)$,
    although the resulting expression isn't terribly helpful.
}

\begin{figure}
    \begin{center}
        \includegraphics[width=0.45\textwidth]{figures/clumpy_2d/clumpy_123.density.3.0}
        \comment{TODO} FIGURE
    \end{center}
    \caption{
        Two panels: left is 2D picture of clumped population;
        right is a bumpy expanding wavefront.
        \label{fig:clumping}
    }
\end{figure}

% % % % % % % % % % % % % % % %
\subsection{Lineage motion is not uniquely determined by population density}

It is natural for applications to wonder about identifiability:
when can the observed quantities like population density
or certain summaries of lineage movement uniquely determine
the underlying demographic parameters?
Consider a deterministic,
continuous population generated by parameters $\gamma$, $\mu_\theta$, and $r$,
with $\meanq = 0$ and $\covq = I$.
Suppose it has a stationary profile $w(x)$, that must satisfy
$$
   r \Delta(\gamma w) + (r \gamma - \mu_\theta) w = 0 .
$$
It is easy to see that $w$ does not uniquely specify $\gamma$, $\mu_\theta$, and $r$:
let $\lambda(x)$ be a smooth, nonnegative function on $\IR^d$,
and let $r'(x, m) = \lambda(x) r(x, m)$ and $\mu_\theta'(x, m) = \lambda(x) \mu_\theta(x, m)$
(and, let $\gamma' = \gamma$).
Then the population with parameters $\gamma'$, $\mu_\theta'$, and $r'$
has the same stationary profile(s) as the original population.

Can these two situations be distinguished from summaries of lineage movement?
The first has lineage generator
\[
    f \mapsto \Lgen f = r \gamma \left( \Delta f + 2 \grad \log(\gamma w) \cdot \grad f \right),
\]
while the second has lineage generator $f \mapsto \lambda(x) \Lgen f(x)$.
In other words,
although the stationary profile of the population is unchanged when we scale
local establishment and death by $\lambda$,
the motion of lineages is sped up locally by $\lambda$.
This corresponds to making areas with $\lambda > 1$ more ``sink-like'' 
and $\lambda < 1$ more ``source-like'':
if $\lambda(x) > 1$, then at $x$ both the death rate and probability of 
establishment of new individuals are higher.
As a result, lineages in the second model spend more time in areas with $\lambda < 1$,
i.e., those areas have higher reproductive value, %long-term fitness
something that is, in principle, discernible from genetic data
(because, for instance, making reproductive value less evenly distributed
reduces long-term genetic diversity).

%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Heuristics}
    \label{sec:heuristics}

In this section we perform some preliminary calculations and use them 
to provide heuristic arguments for our main results,
to build intuition before the formal proofs.

% % % % % % % % % % % % % % % % % %
\subsection{The population density}
    \label{sec:population_heuristics}

{\em We reiterate that in our prelimiting model,
the population is represented by a point measure in which each
individual is assigned a mass $1/N$. We use the term ``population density'' for
this process, as it is supposed to measure population size relative to a 
nominal occupancy of $N$ individuals per unit area, but it is not
absolutely continuous with respect to Lebesgue measure.}

We write $\Pgen^N$
for the generator of the scaled population process $\eta^N$
of Definition~\ref{defn:mgale_construction}
acting on test functions of the form $G( \langle f, \eta \rangle )$,
where $f \geq 0$ is smooth and bounded on $\IR^d$ and 
$G \in C^\infty ([0,\infty))$.
Recall that $\theta =\theta(N) \to \infty$ as $N\to\infty$ in such 
a way that $\theta(N)/N\to\alpha$.
%,
%although we don't remind the reader with extra superscripts.

A Taylor expansion allows us to write
\begin{multline}
	\label{generator prelimit}
    \Pgen^N
    G(\langle f,\eta \rangle)
    =
    G'(\langle f, \eta \rangle)
    \lim_{\delta t\downarrow 0} \frac{1}{\delta t}
    \IE\left[
        \left. \langle f, \eta_{\delta t} \rangle
        -
        \langle f, \eta \rangle
        \right| \eta_0=\eta
    \right]
    \\
    \qquad {}
    + \frac{1}{2}
        G''(\langle f,\eta\rangle)
    \lim_{\delta t\downarrow 0}\frac{1}{\delta t}
    \IE\left[
        \left.\big(\langle f,\eta_{\delta t}\rangle
        -
        \langle f, \eta\rangle\big)^2 \right|\eta_0=\eta
    \right]
    +
    \epsilon_N(f, G, \eta),
\end{multline}
where the terms that make up 
$\epsilon_N(f, G, \eta)$
will be negligible in our scaling limit (at least if $G'''<\infty$).

% % % % % % % 
\subsubsection*{Mean measure}

Recall that in our parameterization only death rates $\mu_\theta$
and the dispersal kernel $q_\theta$ depend on $\theta$.
For a suitable test function $f$, we find
\begin{equation} \label{mean measure}
    \begin{split}
    \Pgen^N \langle f, \eta \rangle
    &=
    \lim_{\delta t\downarrow 0} \frac{1}{\delta t}
    \IE\left[ \left.
        \langle f, \eta_{\delta t} \rangle
        -
        \langle f, \eta\rangle
        \right| \eta_0 = \eta
    \right]
    \\
    &=
    \theta \int
        \int f(z) r(z,\eta) q_\theta(x,dz)
    \gamma(x, \eta) \eta(dx)
    -
    \theta \int f(x)\mu_\theta(x, \eta)
    \eta(dx).
    \end{split}
\end{equation}
The first term is the increment in $\langle f,\eta\rangle$
resulting from a birth event (recalling that
we don't kill the parent) integrated against the rate of such events,
and the second reflects death events.
The factor of $\theta$ appears from the time rescaling.
In both terms,
the rate of events has a factor of $N$ (because events happen at a rate 
proportional to the number of individuals,
whereas $\eta$ has mass $1/N$ for each individual)
which is offset by the fact that  
the birth or loss of a single 
individual at the point $y$, say, changes $\langle f,\eta\rangle$
by $f(y)/N$.

We use the fact that $\int q_\theta(x,dz)=1$ to rewrite~(\ref{mean measure})
as 
\begin{equation}
\label{eqn:rewritten mean measure}
\begin{split}
    \int\left(
        \int \theta \left( f(z) r(z,\eta)- f(x) r(x,\eta) \right) q_\theta(x,dz)
    \right)
    \gamma(x,\eta)
    \eta(dx)
    \\
    + \int \int f(x) \theta \Big(
        r(x,\eta) \gamma(x,\eta)
        - \mu_\theta(x,\eta)
    \Big) \eta(dx).
\end{split}
\end{equation}
We have defined $\mu_\theta$ so that the second term is simple:
\begin{align*}
    \theta \Big( r(x,\eta) \gamma(x,\eta) - \mu_\theta(x,\eta) \Big)
    = F(x, \eta) .
\end{align*}
Furthermore, recall from Definition~\ref{def:dispersal_generator} that
\begin{align} \label{eqn:near_critical}
    \int \theta \Big(
        r(z,\eta) f(z)
        -
        r(x,\eta) f(x)
    \Big) q_\theta(x,dz) 
    \qquad \stackrel{\theta\to\infty}{\longrightarrow} \qquad  
    \DG \big(r(\cdot,\eta)f(\cdot)\big)(x) .
\end{align}
In particular,
if dispersal is determined by a standard multivariate Gaussian
with mean zero and covariance $\sigma^2 I / \theta$,
then $\DG = \sigma^2 \Delta$, where $\Delta$ denotes the Laplacian.

In summary, equation~\eqref{eqn:rewritten mean measure} converges to
\begin{equation} \label{limit of mean measure equation}
\int \gamma(x,\eta)
\DG \big(f(\cdot)r(\cdot,\eta)\big)(x)
\eta(dx)
+
\int f(x)
F(x,\eta)
\eta(dx) .
\end{equation}
This explains the form of the martingale of Theorem~\ref{thm:nonlocal_convergence}.

% % % % % % % 
\subsubsection*{Quadratic variation}

% Note: this bit in firstversionplus.tex has things written out in detail for
% the Poisson offspring number case.

We now look at the second order term in~(\ref{generator prelimit}).
An individual at location $x$ gives birth 
to a surviving offspring at $y$ at rate
$$
\gamma(x,\eta) r(y,\eta) q_{\theta}(x, dy) ,
$$
and since this increments $\langle f, \eta \rangle$ by $f(y) / N$,
the contribution to the quadratic variation from birth events,
which occur at rate $\theta$ per individual 
(so, rate $N\theta |\eta|$ overall), is
$$
\int_{\IR^d}
    N \theta \gamma(x,\eta)
    \int \frac{1}{N^2} f^2(y) r(y,\eta)
    q_\theta(x,dy) 
\eta(dx) .
$$
Similarly, the increment in $\langle f, \eta\rangle$ resulting from 
the death of an individual at $x$ is $-f(x)/N$, and so combining with the 
above, the second order term in the generator takes the form
\begin{align*}
& G''(\langle f,\eta\rangle)
\frac{1}{2} N \theta
\left\{
    \int_{\IR^d}
        \gamma(x,\eta)
        \int \frac{1}{N^2}f^2(y)r(y,\eta)q_\theta(x,dy) 
    \eta(dx)
    +
    \int_{\IR^d}
        \mu_\theta(x,\eta)\frac{1}{N^2}f^2(x) 
    \eta(dx)
\right\} \\
&\qquad
= \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
    \int_{\IR^d}
        \gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta),
    \eta(dx) .
\end{align*}
Since $\int f^2(y) r(y, \eta) q_\theta(x, dy) \to f^2(x) r(x, \eta)$
and $r\gamma  + \mu_\theta = 2 r \gamma  - F / \theta \to 2 r \gamma$
as $\theta \to \infty$,
this converges to
\begin{align*}
\frac{\alpha}{2} G''(\langle f, \eta \rangle)
    \big\langle
	2 r(x, \eta) \gamma(x, \eta),
        \eta(dx)
    \big\rangle .
\end{align*}

An entirely analogous argument shows that if $G'''$ is bounded, then
the term $\epsilon_{\theta,N}(f, G, \eta)$ 
in~(\ref{generator prelimit}) will be $\bigO(\theta/N^2)$.

If we hold $\rho_\gamma$, $\rho_r$, $\rho_F$ fixed, then 
by taking $\theta/N \rightarrow 0$, the second order term in the generator 
will vanish and we expect a deterministic limit,
for which $\partial_t \langle f, \eta_t \rangle$ is equal to~\eqref{limit of mean measure equation}.
In other words, the limit is a weak solution to the deterministic equation
\begin{equation}
\label{deterministic limit}
\partial_t\varphi_t(x)
=
    r(x,\varphi_t)
    \DG\big(
        \gamma(\cdot,\varphi_t) \varphi_t(\cdot)
    \big)(x)
    + F(x, \varphi_t) \varphi_t(x) 
\end{equation}
in the sense of Definition~\ref{defn:weak_solutions},
where $\varphi_t$ is the density of $\eta_t$, if it has a density.
On the other hand, if $N = \alpha \theta$ for some $\alpha > 0$,
the second order term remains, and we expect a ``generalised superprocess'' limit.
The limiting quadratic variation
is exactly as seen in Theorem~\ref{thm:nonlocal_convergence}.

\paragraph{One-step convergence:}
In order to pass directly to a classical PDE limit
in Theorem~\ref{thm:local_convergence}
we impose the stronger condition
that $\theta/N\epsilon^d\to 0$ and also require that
$\theta\epsilon^2\to\infty$. 
Recall that in this case, we take $\rho_F^\epsilon$ 
to be a symmetric Gaussian density with variance 
$\epsilon^2$. The condition $\theta\epsilon^2\to\infty$ 
ensures that $\epsilon^2$ is large enough relative to $1/\theta$
that the regularity gained by smoothing our population density by convolution with
$\rho_\epsilon$ is preserved under the dynamics dictated by $q_{\theta}$.
To understand the first condition, note that we are aiming to obtain a 
deterministic expression for the limiting population density. 
It is helpful to think
about a classical Wright-Fisher model (with no spatial structure and just two types, say). 
We know then that if the
timescale $\theta$ is on the same order as population size $N$, we see stochastic
fluctuations in the frequencies of the two types in the limit as $N\to\infty$; to 
obtain a deterministic limit, we look over timescales that are short relative to population
size. In our setting, the total population size is replaced by the local population
size, as measured by convolution with $\rho_{\epsilon}$, which we expect to be 
of order $N\epsilon^d$, and so in order to ensure a deterministic limit we 
take $\theta/(N\epsilon^d)\to 0$.

% % % % % % % % % % % % % % % % % %
\subsection{Motion of ancestral lineages}

Although our proof of Theorem~\ref{thm:lineages}
uses the explicit representation in terms of the lookdown process,
the result can be understood through informal calculations.
Suppose that we have traced a lineage back to an individual 
at location $y$ at time $t$.
Looking further back through time, at the time of the birth of that 
individual, the lineage will jump to the location of the parent of the
individual.
Now, the rate at which new individuals are born to parents at $x$ 
and establish at $y$ is
$$
\theta N\eta^N_t(dx)    \gamma(x, \eta^N_t) q(x, dy) r(y, \eta^N_t) .
$$
Suppose that $\eta^N$ did have a density (in the prelimit it does not),
say $\eta^N_t(dx) = \varphi^N_t(x) dx$. Informally, the 
probability that a randomly chosen individual near $y$
is a new offspring from a parent at $x$ in $[t, t+dt)$ is
\begin{equation} \label{eqn:informal_rates}
\frac{\theta
    \varphi^N_t(x) \gamma(x, \eta^N_t) r(y, \eta^N_t)
}{
    \varphi^N_t(y)
} \frac{ q_\theta(x, dy) }{ dy } dx dt .
\end{equation}
Leaving aside questions of whether a lineage can be treated 
as a randomly chosen individual, we define
a continuous-time jump process 
whose transition rates, conditional on $(\varphi^N_t)_{t=0}^T$, 
are given by \eqref{eqn:informal_rates}. Because we are tracing the
lineage backwards 
in time we make the substitution $s=T-t$ and write
$(L^N_s)_{s=0}^T$ for the location of a lineage that moves 
according to these jump rates.
Then, abusing notation to write 
$q_\theta(x, y)$ for the density of $q_\theta(x,dy)$,
\begin{align} \label{eqn:lineage_generator_Edt}
    \begin{split}
    &\IE[f(L^N_{s+ds}) - f(y) \;|\; L^N_s = y]
    \\&\qquad 
    =
    ds\, \theta \int \left(f(x) - f(y)\right)
    \frac{
        \varphi^N_{T-s}(x) \gamma(x, \eta^N_{T-s}) r(y, \eta^N_{T-s})
    }{
        \varphi^N_{T-s}(y)
    }
    q_\theta(x, y) dx .
    \end{split}
\end{align}
(Note that this integral is with respect to $x$.)
Referring back to Definition~\ref{def:dispersal_generator},
a quick calculation shows that as $N \to \infty$,
\begin{align*}
    &
    \theta \int \big(f(x) - f(y)\big) g(x) q_\theta(x, y) dx \\
    &\qquad =
    \theta \int \big\{
        (f(x) g(x) - f(y) g(y)) - f(y) (g(x) - g(y))
    \big\} q_\theta(x, y) dx \\
    &\qquad \to
        \DG^*(fg)(y) - f(y) \DG^* g(y) . 
\end{align*}
Applying this to~\eqref{eqn:lineage_generator_Edt} with $g = \varphi_{T-s} \gamma$,
this suggests that the generator of the limiting process is
\begin{align} \label{eqn:heuristic_lineage_generator}
    \Lgen_s f
    &=
    \frac{r}{\varphi_{T-s}}
    \left\{
        \DG^*(\gamma \varphi_{T-s} f) - f \DG^*(\gamma \varphi_{T-s})
    \right\} .
\end{align}
This agrees with Theorem~\ref{thm:lineages}.

%%%%%%%%%%%%%%%%%%%
\section{The lookdown process}
    \label{sec:lookdown}

Our characterisation of the motion of lines of descent (from which we establish
that of ancestral lineages) when we pass
to the scaling limit in our model will be justified via a lookdown construction.
In this section 
we present a lookdown construction for the general population model 
of Definition~\ref{defn:mgale_construction}. It will be
in the spirit of~\cite{kurtz/rodrigues:2011}. 
The general set-up is as follows.
Each individual will be labelled with a ``level'',
which will be a number in $[0, N]$.
We will still encode the process embellished by these levels
as a point measure:
if the $i^\mathrm{th}$ individual's spatial location is $x_i$
and level is $u_i$, then we will write
\[
    \lp^N = \sum_i \delta_{x_i, u_i} ,
\]
which is a measure on $\IR^d \times [0, N]$.
Note that each individual contributes mass 1 to the measure,
not $1/N$ as above. If we assign mass $1/N$ to each individual and 
ignore the levels we will recover our population model. 
Moreover, at any time, the levels of 
individuals in a given spatial region 
will be exchangeable and conditionally uniform on $[0, N]$:
in particular, choosing the $k$ individuals with the lowest levels 
in that region
is equivalent to taking a uniform random sample of size $k$ 
from the population in the region.
However, this exchangeability is only as regards the \emph{past}:
an individual's level encodes information about their future reproductive output,
since individuals with lower levels tend to live longer, and have more offspring.
For more explanation of the set-up and how this is possible,
see \citet{kurtz/rodrigues:2011} and \citet{etheridge/kurtz:2019} 
(and note that our $N$ corresponds to the $\lambda$ of those papers).
The power of this approach is that we can pass to a limit under the
same scalings as described in Theorem~\ref{thm:nonlocal_convergence}, 
and 
the limiting ``spatial-level'' process will still be a point measure.
In this way, 
we explicitly retain the notion of individuals and lineages as we pass to the 
infinite-population limit.

% % % % % % % % % % % % % % % %
\subsection{Lookdown representation of the model of 
Definition~\ref{defn:mgale_construction}}
\label{sec:lookdown_defn}

{\em For the remainder of this section, when there is no risk of ambiguity we shall
suppress the superscript $N$ on the processes $\eta$ and $\xi$.}

In this subsection,
we'll define the process $(\lp_t)_{t \ge 0}$ in terms of the 
dynamics of labelled particles,
and write down its generator.
The dynamics depend on the spatial locations of particles,
and in this section $\eta_t$ is the corresponding spatial measure,
i.e.,
$$
    \eta_t(\cdot) = \frac{1}{N} \lp_t(\cdot \times [0, N])  .
$$
A nontrivial consequence of the way we define $\lp_t$ will be that
this process has the same distribution as the process $(\eta_t)_{t \ge 0}$ 
of Definition~\ref{defn:mgale_construction}.
(This provides our justification for using the same notation for both.)

Following~\cite{etheridge/kurtz:2019}, we build the generator step by step
from its 
component parts. 
Suppose that the initial population is composed of $O(N)$ particles
with levels uniformly distributed on $[0, N]$,
and that the current state of the population is $\lp$,
with spatial projection $\eta$.

An individual at spatial location $x$ with level $u$
produces one juvenile offspring at rate 
$$
2 \theta \left(1 - \frac{u}{N}\right) \gamma(x, \eta) ,
$$
which disperses to a location relative to $x$ drawn 
from the kernel $q_{\theta}(x, \cdot)$.
Averaging over the uniform distribution of the level $u$,
we recover the birth rate $\theta\gamma(x, \eta)$.
This juvenile -- suppose its location is $y$ --
either survives, with probability $r(y, \eta)$, or immediately dies.
(As before, ``maturity'' is instantaneous.)
If it survives, 
a new level $u_1$ is sampled independently and uniformly from $[u,N]$,
and the parent and the offspring are assigned in random order to the 
levels $\{u, u_1\}$.
This random assignment of levels to parent
and offspring will ensure that assignment of individuals to levels remains exchangeable.

Evidently this mechanism increases the proportion
of individuals with higher levels.
To restore the property that
the distribution of levels is conditionally uniform
given $\eta$,
we impose that 
the level $v$ of an individual at location $x$
evolves according to the differential equation
$$
    \dot{v}
    =
    -\theta \frac{v}{N} \left(N - v\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) .
$$
Since $v \in [0, N]$, this moves levels down;
see~\cite{etheridge/kurtz:2019}, Section~3.4 for a detailed explanation.

Levels never cross below 0,
while particles whose levels move above $N$ are regarded as dead
(and are removed from the population).
Therefore, in order to incorporate death,
the level of the individual at location $x$ with level $u$
moves upwards at an additional rate $\theta\mu_\theta(x,\eta) u$.
Since levels are uniform,
it is easy to check that if $\mu_\theta$ were constant,
this would imply an exponential lifetime for each individual;
see~\cite{etheridge/kurtz:2019}, Section~3.1
for more general justification.

Putting these together,
the level $u$ of an individual at $x$ evolves according to:
\begin{equation} \label{eqn:dot_u}
    \dot u
    =
    - \theta\frac{u}{N} \left(N - u\right)
    \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) 
    +
    \theta\mu_\theta(x,\eta) u .
\end{equation}
We shall write 
\begin{align} \label{b_defn}
    b_\theta(x, \eta)
    :=
    \theta\left(
    \gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy)
    -
    \mu_\theta(x,\eta)
    \right) ,
\end{align}
which captures the local net difference between reproduction and death,
and
\begin{align} \label{c_defn}
    c_\theta(x, \eta)
    :=
    \frac{\theta}{N} \gamma(x, \eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy) ,
\end{align}
which captures the local rate of production of successful offspring.
Recall from equation~\eqref{eqn:mu_defn} that
$F(x,\eta) = \theta(r(x,\eta)\gamma(x,\eta) - \mu_\theta(x,\eta))$,
and so 
\begin{equation*} 
b_\theta(x, \eta)
=    \theta \gamma(x, \eta) \int_{\IR^d} \left( r(y, \eta) - r(x, \eta) \right) 
	q_\theta(x, dy)
    + F(x, \eta). 
\end{equation*}
Under Assumptions~\ref{def:model_setup}, as $\theta \to \infty$
$c_\theta(x, \eta)$ will tend to $\alpha \gamma(x, \eta) r(x, \eta)$
and $b_\theta(x, \eta)$ will tend to
\begin{equation}
\label{eqn:b_limit}
    \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta) .
\end{equation}

We can then rewrite the differential equation 
governing the dynamics of the level of each individual as
\begin{align}
\dot{u}
    &=
    \theta\gamma(x,\eta) \int_{\IR^d} r(y, \eta) q_\theta(x, dy)
    \left\{
        -\frac{u}{N}\left(N - u\right)
        + u
    \right\}
    -
	b_\theta(x,\eta) u
    \nonumber \\
    &=
    c_\theta(x, \eta) u^2
    -
	b_\theta(x, \eta) u
    . \label{differential equation for level}
\end{align}

Now, we can write down the generator for $(\lp_t)_{t \ge 0}$,
the lookdown process.
In what follows, we will write sums (and, products) over ``$(x, u) \in \xi$''
to mean a sum over the (location, level) pairs of each individual in the population.
Test functions for $\lp$ will take the form
\begin{equation} \label{eqn:test_functions}
    f(\lp) = \prod_{(x,u)\in \lp}g(x,u)=\exp\left(\int \log g(x,u)\lp(dx, du)\right),
\end{equation}
where
$g(x,u)$ is differentiable in $u$ and 
smooth in $x$.
We will also assume that $0\leq g(x,u) \leq 1$ for all $u\in [0,N]$,
and $g(x,u)\equiv 1$ for $u\geq N$.
In the expressions that follow,
we shall often see one or more factor of $1/g(x,u)$;
it should be understood that if $g(x,u)=0$,
then it simply cancels 
the corresponding factor in $f(\lp)$.

First consider the terms in the generator that come from birth events.
When a birth successfully establishes,
a new level is generated above the parent's level,
and this new level is assigned to either the offspring or the parent.
Since the probability of each is 1/2,
the contribution of birth to the generator maps $f(\lp)$ to
\begin{align}
\label{partition over level assignment}
    &
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \frac{\theta}{N} \gamma(x, \eta)
    \int_u^N
    \int_{\IR^d}
    \left(
    \frac{1}{2}
    \bigg\{
            g(y, u_1)
        + \frac{ g(y, u) g(x, u_1) }{ g(x, u) }
    \bigg\}
        - 1
    \right)
    r(y, \eta) q_\theta(x, dy)
    du_1
    \\
    \begin{split} \label{eqn:birth_generator}
&=
    f(\lp)
    \sum_{(x, u) \in \lp}
    2 \gamma(x, \eta)
    \bigg\{
        \frac{1}{2 N}
        \int_u^N
        g(x, u_1) du_1
        \frac{
            \theta \int_{\IR^d} (g(y, u) - g(x, u)) r(y, \eta) q_\theta(x, dy)
        }{
            g(x, u)
        }
    \\ & \qquad \qquad \qquad {}
        + \frac{\theta}{N}
        \int_u^N \int_{\IR^d}
        \left( \frac{g(y, u_1) + g(x, u_1)}{2} - 1 \right)
        r(y, \eta) q_\theta(x, dy)
        du_1
    \bigg\}
    .
    \end{split}
\end{align}
In (\ref{partition over level assignment}),
$u_1$ is the new level and $y$ is the offspring's location,
and so the two terms in the integral correspond to the two situations:
in the first, we have added an individual at $(y, u_1)$,
while in the second, we replace an individual at $(x, u)$
by one at $(x, u_1)$ and another at $(y, u)$.
We've rewritten it in the form~(\ref{eqn:birth_generator})
because each of the two pieces
naturally converges to a separate term in the limit.

The remaining term in the generator is due to the motion of particles' levels.
Reading off from~(\ref{differential equation for level}),
it takes the form
\begin{align} \label{eqn:level_generator}
    f(\lp)
    \sum_{(x, u) \in \lp}
    \left(
    \frac{\theta}{N}
        c_\theta(x, \eta) u^2
        -
        b_\theta(x, \eta) u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)} .
\end{align}

We can now define the spatial-level process
explicitly as a solution to a martingale problem,
whose generator is just the sum of
\eqref{eqn:birth_generator} and \eqref{eqn:level_generator}.
We need some notation. 
Write ${\mathcal C}={\mathcal C}(\IR^d\times [0,\infty)$ for the counting measures 
on $\IR^d\times [0,\infty)$ and ${\mathcal C}_N$ for the subset consisting of 
counting measures on $\IR^d\times [0,N]$.

\begin{definition}[Martingale Problem Characterisation]
    \label{defn:lookdown_mgale}
For given positive values of $N$ and $\theta$, define the generator $A^N$
by
\begin{equation}
	\label{eqn:lookdown_generator}
\begin{split}
& A^{N}f(\lp ) \\
&\quad =
    f(\lp)
    \,\sum_{(x,u)\in \lp}\,
    2 \gamma(x, \eta)
    \Bigg\{ \frac{1}{2N} \int_u^N g(x,u_1) du_1
            \frac{
                \theta \int_{\IR^d}
                (g(y,u) - g(x,u))
                r(y, \eta) q_{\theta}(x,dy)
            }{ g(x,u) }
        \\
    &\qquad\qquad\qquad\qquad\qquad\qquad\qquad {} +
        \frac{\theta}{N} \int_u^N
        \int_{\IR^d}\left(
            \frac{ g(y,u_1) + g(x,u_1) }{ 2 } - 1
        \right)
        r(y, \eta) q_{\theta}(x,dy)
        du_1
    \Bigg\}\\
    &\qquad\qquad {} +
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left(
        c_\theta(x,\eta) u^2 - b_{\theta}(x,\eta)u
    \right)
    \frac{\partial_u g(x,u)}{g(x,u)}
    ,
\end{split}
\end{equation}
where $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$ is as defined in \eqref{eqn:test_functions},
and $\eta(\cdot) = \lp(\cdot \times [0, N]) / N$
as before. 
Given $\lp_0 \in {\mathcal C}_N$, 
we say that a ${\mathcal D}_{[0,\infty)}({\mathcal C}_N)$-valued process 
$(\xi_t)_{t \ge 0}$
   is a solution to the $(A^N, \lp_0)$ martingale problem
    if 
$f(\lp_t) - f(\lp_0)-\int_0^t A^N f(\lp_s) ds$ is a martingale (with 
respect to the natural filtration) for all 
test functions $f$ 
    as defined above.
\end{definition}

The martingale problem for finite $N$ has a unique solution.
Next we state the limiting martingale problem,
for which we do not necessarily have uniqueness.
As before, the parameter $\alpha$ will correspond to 
$\lim_{N\to\infty}\theta(N)/N$. Whereas for finite $N$, conditional
on the population process $\eta_t^N$, the levels of particles are
independent and 
uniformly distributed on $[0,N]$, in the infinite population limit, 
conditional on $\eta_t$, the process $\xi_t$ is Poisson distributed on 
$\IR^d\times [0,\infty)$ with mean measure $\eta_t\times\lambda$, where 
$\lambda$ %%%place changed notation for Lebesgue measure
is Lebesgue measure.

\begin{definition}[Martingale Problem Characterisation, scaling limit]
    \label{defn:limiting_lookdown_mgale}
    Fix $\alpha \in [0, \infty)$, and define  
    test functions $f$ by 
    $f(\xi) = \prod_{(x, u) \in \xi} g(x, u)$ 
	with $g$ differentiable in $u$, smooth in $x$, satisfying 
$0\leq g(x,u)\leq 1$ and
    such that there exists a $u_0$ with $g(x, u) = 1$ for all $u > u_0$.
    Then, define the operator $A$ on such test functions by 
    \begin{align} \begin{split} \label{eqn:limiting_lookdown_generator}
    A f(\lp)
    &=
        f(\lp)  \sum_{(x, u) \in \xi}
        \gamma(x, \eta)
            \frac{
                \DG(g(\cdot, u) r(\cdot, \eta))(x) - g(x,u) \DG r(x,\eta)
            }{
                g(x, u)
            }
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        2 \alpha \gamma(x, \eta) r(x, \eta) \int_u^\infty (g(x, u_1) - 1) du_1
    \\ &\qquad {} +
        f(\lp) \sum_{(x, u) \in \xi}
        \left(
            \alpha \gamma(x, \eta) r(x, \eta) u^2
            -
            \left\{
                \gamma(x, \eta) \DG r(x, \eta) + F(x, \eta)
            \right\} u
        \right)
        \frac{\partial_u g(x, u)}{ g(x,u) }  .
    \end{split} \end{align}
    We say that a ${\cal D}([0,\infty), \mathcal{C})$-valued process $(\xi_t)_{t \ge 0}$
    is a solution to the $(A, \lp_0)$ martingale problem
    if it has initial distribution $\lp_0$
    and $f(\lp_t) - f(\lp_0)-\int_0^t Af(\lp_s) ds$ is a martingale (with
respect to the natural filtration) for all 
test functions $f$
    as defined above.
\end{definition}

The lookdown processes have been carefully constructed so that
observations about the past spatial positions of individuals in 
the population do not give us any information about the assignment of 
individuals to levels.
In other words, the dynamics of the lookdown
process preserve the conditionally uniform (or in the limit, conditionally Poisson) structure
-- if started with uniform levels,
levels are uniform at all future times.
Moreover, if we average over levels in the expression for the generator
(equation~\eqref{eqn:lookdown_generator} or~\eqref{eqn:limiting_lookdown_generator})
we recover the generator for the population process.
%(visible in equation~\eqref{eqn:eta_martingale}
%or~\eqref{eqn:limiting_mgale_problem}, respectively).
Once this is verified (along with some boundedness conditions)
the Markov Mapping Theorem (Theorem~\ref{thm:mmt}; also see \citep{etheridge/kurtz:2019})
tells us that by ``removing labels'' from the lookdown process $\lp$
we recover the population process $\eta$.
% In particular, for any
% solution to the martingale problem of Definition~\ref{defn:mgale_construction}
% % of equations~\eqref{eqn:eta_martingale} and \eqref{eqn:prelimit_martingale_variation}
% there is a % lookdown representation $(\lp^N_t)_{t \ge 0}$
% solution to the martingale problem of Definition~\ref{defn:lookdown_mgale})
% for which $\eta^N_t(\cdot) = \lp^N_t(\cdot \times [0, N])/N$
% and such that at each $t$, conditioned on $(\eta^N_s)_{s \le t}$, the levels of $\lp_t^N$
% are independently and uniformly distributed on $[0, N]$.
% Similarly, in the scaling limit:
% for any solution $(\eta_t)_{t \ge 0}$ to the martingale problem of 
% Theorem~\ref{nonlocal_convergence}
% %there is a lookdown process $(\lp^N_t)_{t \ge 0}$
% there is a solution to the martingale problem of Theorem~\ref{lookdown_convergence}
% for which $\eta_t(\cdot) = \lim_{u_0 \to \infty} \frac{1}{u_0} \lp_t(\cdot \times [0, u_0))$
% and such that at each $t$, conditional on $(\eta_s)_{s\leq t}$, 
% $\lp_t$ is conditionally Poisson with mean measure $\eta_t(dx) \times du$.
% Furthermore, convergence of the lookdown processes would imply convergence of the population processes
% \citep[Thm A.9 in][]{kurtz/rodrigues:2011}.

To make this precise,
define the spatial projection maps
$\kappa^N: \mathcal{M}(\IR^d \times [0, N]) \to \mathcal{M}(\IR^d)$
by $\kappa^N(\xi^N)(\cdot) = \xi^N(\cdot \times [0, N]) / N$,
and $\kappa: \mathcal{M}(\IR^d \times [0, \infty)) \to \mathcal{M}(\IR^d)$
by $\kappa(\xi)(\cdot) = \lim_{u_0 \to \infty} \xi(\cdot \times [0, u_0]) / u_0$.
We will also need an inverse notion:
for a measure $\lp^N$ on $\IR^d \times [0, N]$ and a sigma-field $\mathcal{F}$,
we say that \emph{$\lp^N$ is conditionally uniform given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_uniform}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle H_f^N, \kappa^N(\lp) \rangle} ,
\end{align}
where
$$ H_f^N(x) = - N \log \frac{1}{N} \int_0^N e^{-f(x, u)} du . $$
In other words, the $[0, N]$ components of $\lp$
are independent, uniformly distributed on $[0, N]$, and independent of $\kappa(\lp)$.
Similarly, for a measure $\lp$ on $\IR^d \times [0, \infty)$
we say that \emph{$\lp$ is a conditionally Poisson random measure given $\mathcal{F}$}
if for all compactly supported $f$,
\begin{align} \label{eqn:conditionally_poisson}
    \IE[ e^{-\langle f, \lp \rangle} \;|\; \mathcal{F} ]
    &=
    e^{-\langle \int_0^\infty (1 - e^{-f(x, u)}) du, \kappa(\lp)(dx) \rangle} .
\end{align}
In other words, $\lp$ is conditionally Poisson with Cox measure $\kappa(\lp) \times \lambda$,
where $\lambda$ is Lesbegue measure.

\begin{proposition}
    \label{thm:mmt_application}
    If $\tilde \eta^N$ is a solution
    of the martingale problem of Definition~\ref{defn:mgale_construction}
    with initial distribution $\eta^N_0$
    then there exists a solution $\lp^N$
    of the $(A^N, \lp^N_0)$-martingale problem of Definition~\ref{defn:lookdown_mgale}
    such that $\eta^N = \kappa^N \circ \lp^N$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta^N$.
    Furthermore, for each $t$, $\lp^N_t$ is conditionally uniform given given $\mathcal{F}^{\eta^N}_t$
    in the sense of \eqref{eqn:conditionally_uniform}.

    Similarly,
    if $\tilde \eta$ is a solution
    of the limiting martingale problem of Theorem~\ref{thm:nonlocal_convergence}
    with initial distribution $\eta_0$
    then there exists a solution $\lp$
    of the martingale problem of of Definition~\ref{defn:limiting_lookdown_mgale}
    such that $\eta = \kappa \circ \lp$
    has the same distribution on $D_{\measures}[0, \infty)$
    as $\tilde \eta$.
    Furthermore, $\lp_t$ is conditionally Poisson given $\mathcal{F}^{\eta}_t$
    in the sense of \eqref{eqn:conditionally_poisson}.
\end{proposition}

Now we can present the main convergence theorem that is analogous
to Theorem~\ref{thm:nonlocal_convergence} for the population process.

\begin{theorem}
    \label{thm:lookdown_convergence}
    Let $(\lp_t^N)$ satisfy Definition~\ref{defn:lookdown_mgale}
    and assume that as $N \to \infty$,
    $\theta \to \infty$ in such a way that $\theta/N \to \alpha$.
    Let $\eta_0^N = \kappa(\lp_0^N)$ and
    suppose also that $\eta^N_0 \to \eta_0$ in $\measures$,
    and that for each $N$, $\lp^N_0$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}.
    Then, any sequence of stochastic processes $(\lp_t^N)_{t \ge 0}$
    has a subsequence which converges in distribution as $N \to \infty$
    to a measure-valued process $(\lp_t)_{t \ge 0}$
    with $\lp_0$ conditionally Poisson given $\eta_0 = \kappa(\lp_0)$
    in the sense of \eqref{eqn:conditionally_poisson},
    that is a solution to the martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}.
\end{theorem}

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Explicit construction of lines of descent}
    \label{sec: individual lines of descent}

The main interest in using a lookdown construction for our population
processes is that it allows us to retain information about the 
relatedness of individuals as we pass to the infinite population limit.
In order to exploit this, in this section we write down stochastic 
equations for the locations and levels of individuals in the prelimiting lookdown model.
We will then be able to pass to the scaling limit.
This provides an explicit description of the solution to
the limiting martingale problem of Definition~\ref{defn:limiting_lookdown_mgale}
which will enable us to identify all 
individuals in the current population that are descendants of a
given ancestor at time zero. In theory at least, this allows us to
recover all the information about genealogies relating individuals 
sampled from the present day population. This idea draws on the notion
of ``tracers'', popular in statistical physics and used in population
genetics by a number of authors including 
\citet{hallatschek/nelson:2008}, \citet{durrett/fan:2016}, and \citet{biswas/etheridge/klimek:2018}.

We will construct the process using a Ulam-Harris indexing scheme.
First, we assign each individual alive at time 0 a unique label from $\IN$.
Suppose an individual with label $a$ and level $u$ reproduces,
and as a result there are two individuals, one with level $u$ and one with a new level $u_1 > u$.
The parent individual, previously labeled $a$, might be assigned either level.
We will track chains of descendant individuals forwards through time by
following levels, rather than individuals, and will call this a \emph{line of descent}.
So, after reproduction, we give a new label to \emph{only} the individual
that is given the new level $u_1$,
retaining the label $a$ for the individual with the old level $u$.
In this way, at each birth event, a unique label is 
assigned to the resulting individual with the higher level,
and the label of an individual may change throughout its lifetime.

Concretely, then: for each label $a$ in
$\labelspace = \bigcup_{k \ge 1} \IN^k$,
let $\Pi_a$ be an independent Poisson process on 
$[0, \infty)^2 \times \IR^d \times \{0,1\}$.
The mean measure of each $\Pi_a$ is a product of Lebesgue measure on $[0, \infty)^2$,
the density of the standard Gaussian on $\IR^d$, and 
$(\delta_0 + \delta_1)/2$ on $\{0, 1\}$.
It will also be convenient
to suppose that for each label $a$ we have an enumeration of the points in $\Pi_a$,
so we may refer to ``the $j^\text{th}$ point in $\Pi_a$'',
although the precise order of this enumeration is irrelevant.
If $(\tau, v, z, \kappa)$ is the $j^\text{th}$ point in $\Pi_a$,
then $\tau$ will determine a possible birth time,
$v$ will determine the level of the offspring,
$z$ will determine the spatial displacement of the offspring relative to the parent,
$\kappa$ will be used to determine whether parent or offspring is assigned the new level,
and the new label produced will be $a \concat j$,
i.e., the label $a$ with $j$ appended
(so, if $a = (a_1, \ldots, a_k)$ then $a \concat j = (a_1, \ldots, a_k, j)$).
Each label $a$ has a birth time $\tau_a$,
when it is first assigned,
and a (possibly infinite) death time $\sigma_a$, when its level first hits $N$.
For any $\tau_a \le t \le \sigma_a$ we denote by $X_a(t)$ and $U_a(t)$ the spatial location and level
of the individual carrying label $a$ at time $t$, respectively.
Furthermore, define
\begin{align*}
    \eta^N_t = \frac{1}{N} \sum_{a : \tau_a \le t < \sigma_a} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp^N_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} .
\end{align*}

Now, since we have defined labels so that the level does not jump,
$U_a$ satisfies~\eqref{differential equation for level}
	for $\tau_a \le t \le \sigma_a$, i.e.,
\begin{equation} \label{eqn:U_line_of_descent}
    \begin{split}
& U_a(t)
    =
    U_a(\tau_a) \\
&\qquad {}   
    + \int_{\tau_a}^{t}
    \left(
        c_\theta(X_a(s),\eta_s) U_a(s)^2
        -
        b_\theta(X_a(s),\eta_s) U_a(s)
    \right)
    ds ,
\end{split}
\end{equation}
and, of course, $\sigma_a = \inf\{t \ge \tau_a : U_a(t) > N\}$.

Potential reproduction events occur at times $\tau$
for each point $(\tau, v, z, \kappa) \in \Pi_a$ with $\tau_a \le \tau < \sigma_a$.
(We say ``potential'' since if the level of the resulting offspring is greater than $N$,
the event does not happen.)
If this is the $j^\text{th}$ point in $\Pi_a$,
the potential new label is $a \concat j$, the birth time is $\tau_{a \concat j} = \tau$,
and the spatial displacement of the potential ogfspring is $y(X(\tau-), z)$, where
$$
    y(x, z)
    :=
    \frac{1}{\theta}\meanq(x)
    +
    \frac{1}{\sqrt{\theta}}K(x) z,
$$
and $K(x)K^{T}(x) = \covq(x)$.

Next we must choose the new level created at the birth event.
We would like an individual with level $u$ and at spatial position $x$
to produce offspring at $y$ at instantaneous rate
\begin{equation}
	\label{rate of birth}
	2 \Big(1 - \frac{u}{N}\Big) \theta \gamma(x, \eta) r(x + y, \eta).
\end{equation}
To do this
we will associate the point $(\tau, v, z, \kappa) \in \Pi_a$ with 
level $u + v \ell$,
where $\ell$ is chosen so that the rate of appearance of points in 
$\Pi_a$ with level below $N$, that is points with
$v \ell <N-u$, is given by~(\ref{rate of birth}).
Since the mean measure of $\Pi_a$ is Lebesgue measure
in the $t$ and $v$ directions, we must take
\begin{equation}
	\label{expression for ell}
    \ell(x, y, \eta)
    =
    \frac{N-u}{2 (1-u/N) \theta \gamma(x, \eta) r(x + y, \eta) } 
    =\frac{1}{2N^{-1} \theta \gamma(x, \eta) r(x + y, \eta) } 
    ,
\end{equation}
and, using this, the (potential) new level is
\begin{equation*}
    U_{a \concat j}(\tau)
    =
    U_a(\tau)
    +
    v \ell\big(X_a(\tau-), y(X_a(\tau-), z), \eta_{\tau-} \big) .
\end{equation*}
If $U_{a \concat j}(\tau) < N$,
the new individual labeled $a \concat j$ is produced,
and $\kappa$ determines which label, $a$ or $a\concat j$,
is associated with the new location,
so
\begin{align*}
    X_{a \concat j}(\tau)
    &=
    X_a(\tau-) + (1 - \kappa) y\big(X_a(\tau-), z\big) .
\end{align*}
On the other hand
if $U_{a \concat j}(\tau) \ge N$, then $X_a$ is unchanged and 
$X_{a \concat j}$ is undefined,
so
\begin{align} \label{eqn:X_line_of_descent}
    X_a(\tau)
    &=
    X_a(\tau-) + \kappa y(X_a(\tau-), z) \ind_{U_{a \concat j}(\tau) < N} .
\end{align}
Recall that the parental \emph{individual} always retains their spatial location,
so that $\kappa = 0$ corresponds to the parent being assigned a new level,
and our line of descent switching to the offspring.
Combining these observations, $X_a$, for $\tau_a \le t < \sigma_a$, solves the equation 
\begin{align*}
    X_a(t)
    &=
    X_a(\tau_a) + 
    \int_{[\tau_a, t) \times [0, \infty) \times \IR \times [0, 1]}
    y(X_a(\tau-), z)
    \kappa
    \ind_{
        U_a(\tau) + v \ell(X_a(\tau-), y(X_a(\tau-), z), \eta_{\tau-})<N
    }
    d\Pi_a(\tau, v, z, \kappa) .
\end{align*}

Although we have described the evolution of a line of descent only for a given label
(i.e., for $\tau_a \le t < \sigma_a$),
we can extend the definition to times
$0 \le t < \sigma_a$ by setting
$X_a(t)$ equal to $X_{[a]_t}(t)$,
where $[a]_t$ is the label of the ancestor of label $a$ alive at time $t$,
and similarly for $U_a(t)$.
It is then straightforward, albeit tedious,
to write down the time evolution of $(X_a(t), U_a(t))$ for all time back 
to $t=0$
in terms of the driving Poisson processes.

\begin{remark}
Although we have a single construction that couples the processes across all $N$,
unlike in \citet{kurtz/rodrigues:2011},
the actual trajectories, $X_a(\cdot)$, do not necessarily coincide for 
different values of $N$, since they are affected by the whole population process.
However, this does suggest approximating the genealogies in the infinite density limit
by simulating up until a sufficiently high level
that we have a good approximation to the population process.
\end{remark}

% % % % % % % % % % % % %  % % %

%% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Limiting processes for lines of descent}
\label{sec:limiting_lines_of_descent}

The previous section constructed the lookdown process
using the same underlying Poisson processes $\{\Pi_a\}_{a \in \labelspace}$ for
different values of $N$.
As a result, if the spatial projections $\eta$ converge, then
individual lines of descent converge pointwise as $N \to \infty$.
To see this, first note that if the Poisson processes are fixed
then the set of events to which a given label $a \in \labelspace$ is associated
is also fixed -- this is the sequence $(\tau_k, v_k, z_k, \kappa_k)$ associated 
with the label $a$.
To conclude that the lines of descent converge, 
first, we clearly need that the spatial projections $\eta$ converge.
%
Supposing that they do,
consider how a line of descent $(X_a(t), U_a(t))$ evolves.
It throws off a new line of descent at a higher level
when there is a point $(\tau, v, z, \kappa)$ in $\Pi_a$ with $\tau > \tau_a$ and
\begin{equation} \label{eqn:line_points}
    v < 2 \frac{\big(N - U_a(\tau)\big)}{N} \theta \gamma(X_a(\tau-), \eta_{\tau-}) 
	r\Big(X_a(\tau-) + y\big(X_a(\tau-),z\big), \eta_{\tau-}\Big) .
\end{equation}
Since the mean measure of the $v$ coordinate is Lebesgue measure,
$\theta/N \to \alpha$,
and $q_\theta(x, dy) \to \delta_x(dy)$,
this corresponds in the limit to new lines of descent being thrown off 
according to a Poisson process with intensity
$$
2 \alpha \gamma(X_a(t), \eta_t) r(X_a(t), \eta_t)dt \times du.
$$
Now consider the location of the line of descent:
at each birth event, with probability one half
the line of descent jumps to $X_a(t) + y$.
Taking $g$ to be a suitable test function on $\IR^d$, 
and rewriting~\eqref{eqn:line_points}, when the level is $u$
and the state of the population is $\eta$, the generator of the
spatial motion of the line of descent
applied to $g(x)$ is
\begin{align*}
    &
    \left(1 - \frac{u}{N}\right) \gamma(x, \eta)
    \theta \int_{\IR^d} r(x+y, \eta) (g(x+y) - g(x)) q_\theta(x, dy) \\
    &\qquad {}
    =
    \left(1 - \frac{u}{N}\right) \gamma(x, \eta)
    \bigg\{
        \theta \int_{\IR^d} (r(x+y, \eta) g(x+y) - r(x, \eta) g(x)) q_\theta(x, dy) \\
        &\qquad \qquad {}
        -
        \theta \int_{\IR^d} (r(x+y, \eta) - r(x, \eta) ) g(x) q_\theta(x, dy) 
    \bigg\} \\
    &\qquad {}
    \to
    \gamma(x, \eta)
    \left(
        \DG(rg)(x) - g(x) \DG(r)(x)
    \right) ,
    \qquad \text{as } N, \theta \to \infty .
\end{align*}
Notice that the factors of 2 have cancelled,
and that the result is independent of $u$.
Also recall that $r(x, \eta)$ depends on $\eta$ only
through $\smooth{r}\eta(x)$,
which is guaranteed to be smooth, so that $\DG(r)$ 
and $\DG(gr)$ are well-defined.

We write out the differential operator above in more detail.
Recall that $\DG g(x) = \sum_i \meanq_i \partial_i g(x) + \sum_{ij} \covq_{ij} \partial_{ij} g(x)$,
and for the moment write $r(x)$ for $r(x, \eta)$, $\meanq(x) = \meanq$, and $\covq(x) = \covq$
so that
\begin{align}
\DG(rg)(x) - g(x) \DG(r)(x)
    &= \nonumber
    r(x) \sum_i \meanq_i \partial_i g(x)
    + 2 \sum_{ij} \partial_i r(x) \covq_{ij} \partial_j g(x)
    + r(x) \sum_{ij} \covq_{ij} \partial_{ij} g(x) \\
    &= \label{eqn:limiting_generator}
    r(x) \left\{
        \left(
        \meanq
        + 2 \covq \grad \log r(x)
        \right)
        \cdot
        \grad g(x)
        +
        \sum_{ij} \covq_{ij} \partial_{ij} g(x)
    \right\} .
\end{align}

The only thing that remains is to describe how the levels change,
but this is immediate from applying limit~\eqref{eqn:b_limit}
to equation~\eqref{differential equation for level}.

We summarize the results in a proposition.

\begin{proposition}[Line of descent construction]
    \label{prop:limiting_construction}
Let 
$J(x,\eta)$ be such that $r(x,\eta)\gamma(x,\eta)\covq(x) = J(x,\eta) J(x,\eta)^T$, and set
$\beta(x, \eta) = r(x,\eta)\gamma(x,\eta)\big(\meanq(x) + 2 \covq(x) \grad \log r(x, \eta)\big).$
Associate with each label $a \in \labelspace = \cup_{k\geq 1}\IN^k$
an independent $d$-dimensional Brownian motion $W_a$
and an independent Poisson process $R_a$ on $[0, \infty)^2$
with Lebesgue mean measure,
and with points ordered in some way.
Given $\eta_0 \in \measures$,
let $(x_i, u_i)$ be the points of a Poisson process $R_0$
on $\IR^d \times [0, \infty)$ 
with mean measure $\eta_0\times \lambda$
(the product of $\eta_0$ and Lebesgue measure).
For each $i$, with $(x_i, u_i)$ the $i^\text{th}$  point of $R_0$, begin a line of descent
with label $i$, location $X_i(0) = x_i$, 
level $U_i(0) = u_i$, and birth time $\tau_i = 0$.

Write $\tau_a$ for the birth time of the label $a$ and 
$\sigma_a = \lim_{K \to \infty} \inf\{t \ge 0: U_a(t) > K\}$
the time the level hits $\infty$.
Suppose that the spatial locations and level of each line of descent $a$
solve, for $\tau_a \le t < \sigma_a$,
\begin{align}
    \label{eqn:limiting_construction}
    \begin{split}
X_a(t)
    &=
    X_a(\tau_a)
    + \int_{\tau_a}^{t}
        \beta(X_a(s), \eta_s) ds
    + \int_{\tau_a}^{t}
        J(X_a(s),\eta_s) dW_a(s)
    \\
U_a(t)
    &=
    U_a(\tau_a)
    + \int_{\tau_a}^{t}
    \bigg(
        \alpha \gamma(X_a(s),\eta_s)
        r(X_a(s), \eta_s) U_a(s)^2
\\ &\qquad \qquad \qquad {}   
        -
        \big\{
            \gamma(X_a(s),\eta_s) \DG r(X_a(s),\eta_s)
            + F(X_a(s), \eta_s)
        \big\}
        U_a(s)
    \bigg)
    ds .
    \end{split}
\end{align}
Each point in each $R_a$ denotes a potential birth time for $a$:
if the $j^\text{th}$ point in $R_a$ is $(\tau, v)$, with 
$\tau_a \le \tau < \sigma_a$,
then a new line of descent with label $a \concat j$ is produced,
with birth time $\tau_{a \concat j} = \tau$,
    location $X_{a \concat j}(\tau) = X_a(\tau)$, and level
$$
    U_{a \concat j}(\tau) = U_a(\tau)
    + \frac{v}{ 2 \alpha \gamma(X_a(\tau), \eta_\tau) r(X_a(\tau), \eta_\tau) } ,
$$
if this is finite.
For any solution to the equations above, the processes defined by
\begin{align*}
    \eta_t
    =
    \lim_{u_0 \to \infty} \frac{1}{u_0}
        \sum_{a : \tau_a \le t < \sigma_a \; ; \; U_a(t) < u_0} \delta_{X_a(t)}
    \qquad \text{and} \qquad
    \lp_t = \sum_{a : \tau_a \le t < \sigma_a} \delta_{(X_a(t), U_a(t))} 
\end{align*}
are solutions to the martingale problems of
Theorems~\ref{thm:nonlocal_convergence} and~\ref{thm:lookdown_convergence},
respectively.
\end{proposition}

In particular, note that if $\alpha=0$,
no new lines of descent are produced.
More precisely, comparing with~(\ref{expression for ell}),
they are produced, but ``at infinity'',
and their trace is seen in the 
spatial motion of the line of descent which results from
the production of these lineages.

\begin{proof}[of Proposition~\ref{prop:limiting_construction}]
    The fact that a solution to the system of equations~\eqref{eqn:limiting_construction}
    is a solution to the martingale problem of Theorem~\ref{thm:lookdown_convergence}
    is an application of It\^o's theorem.
    Furthermore, in Proposition~\ref{thm:mmt_application} we showed that
    the conditional Poisson property of $\lp_0$ is preserved
    (i.e., holds for $\lp_t$ for all $t$), and so
    $(\eta_t)_{t \ge 0}$ is well-defined,
    and furthermore that $\eta_t$ is a solution
    to the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Proofs of convergence for nonlocal models}
\label{sec:proofs}

In the following sections we present formal proofs of our three scaling limits.
In Subsection~\ref{sec:population_density_proof} 
we prove Theorem~\ref{thm:nonlocal_convergence}, to obtain a limit in
which interactions between individuals in the population are nonlocal.
In Subsection~\ref{subsec:nonlocal to local}
we show how, in two important examples in which the nonlocal
limit is respectively a deterministic solution to a non-local equation of 
reaction diffusion type and a deterministic solution to a nonlocal 
porous medium equation with an additional logistic growth term, 
one can pass to a further 
limit to obtain a classical PDE.
In subsequent sections (Sections~\ref{subsec:one step convergence proof} and~\ref{sec:lookdown_proofs})
we prove our third scaling limit, in which 
we pass from our population model 
directly to a reaction diffusion equation without going via a nonlocal limit,
and finally our results concerning both convergence of the lookdown representation
and characterization of the lineage process in the deterministic case.
\comment{TODO: revisit this paragraph}

\subsection{Preliminaries}

Below we will have frequent use for the quantity
\begin{equation}
\label{generator Bf}
    B^\theta_f(x, \eta) = \theta \int_{\IR^d} (f(y) r(y, \eta) - f(x) r(x, \eta)) q_\theta(x, dy) .
\end{equation}

First, we prove Lemma~\ref{lem:conditions on r}.

\begin{proof}[of Lemma~\ref{lem:conditions on r}]
Here, we need to prove that $|\gamma(x, \eta) B^\theta_f(x, \eta)|$ is bounded,
uniformly over $x$ and $\eta$.
First suppose that
Condition~\ref{control through r} of  
Lemma~\ref{lem:conditions on r} is satisfied.
We write
\begin{align*}
    r\big(y,\rho_r*\eta(y)\big) f(y) - r\big(x,\rho_r*\eta(x)\big) f(x)
    &=
    \sum_i (y_i - x_i) \partial_{x_i} \Big( r\big(x,\rho_r*\eta(x)\big) f(x) \Big) 
    \\ &\qquad {}
    +
    \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
        \partial_{x_i x_j} \Big( r\big(z, \smooth{r}\eta(z)\big) f(z) \Big) ,
\end{align*}
where $z = \kappa x + (1-\kappa) y$ for some $0 \le \kappa \le 1$.
Condition~\ref{control through r} 
gives uniform bounds on the derivatives of $r(x, \rho_r*\eta(x))$
in this expression and so, provided $f$ also has uniformly bounded first and second derivatives,
we have a bound of the form
\begin{multline*}
|B^\theta_f|
\leq
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| 
\end{multline*}
for suitable constants $K_1$, $K_2$.
This is bounded uniformly because
$q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
are bounded uniformly in $x$ under Assumptions~\ref{def:model_setup}.
Since $\gamma$ is bounded as well, so is $|\gamma B^\theta_f|$.

Now suppose that instead
Condition~\ref{control through gamma} of Lemma~\ref{lem:conditions on r} is satisfied.
First note that
    \begin{align}
	    \nonumber
    & |B_f^\theta| =
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x, \smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
 \label{eqn:q_bound_split}
	    &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(y) r\big(y, \smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                \big\}
            q_{\theta}(x,dy)
        \right|
    \\ 
\nonumber
	    &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \big\{
                    f(x) r\big(x,\smooth{r}\eta(y)\big)
                    -
                    f(x) r\big(x,\smooth{r}\eta(x)\big)
                \big\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

 Writing
    $K_3 = \sup_{x,m} \max_i |\partial_{x_i} f(x)r(x, m)|$
    and 
    $K_4 = \sup_{x,m} \max_{i,j} |\partial_{x_i x_j} f(x)r(x, m)|$, since
\[\big|\partial_{x_i}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_3\]
and
\[\big|\partial_{x_ix_j}\big(f(x)r(x,\rho_r*\eta(y))\big|\leq K_4,\]
the first term
is bounded exactly as above. For the second, 
    \begin{multline*}
        r(x, \smooth{r}\eta(y)) - r(x, \smooth{r}\eta(x))
        \\
=
        (\smooth{r}\eta(y) - \smooth{r}\eta(x)) r'(x,\smooth{r}\eta(x))
        + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 r''(x, \overline{m}),
    \end{multline*}
    where $\overline{m}= \kappa'\smooth{r}\eta(x) +(1-\kappa') 
\smooth{r}\eta(y)$ for some $0\leq \kappa'\leq 1$, and
    we have used $r'$ and $r''$ to denote the first and second derivatives of $r(x,m)$
with respect to the second argument.
    So, writing $K_5=\|f\|_\infty\|r'\|_\infty$ and $K_6=\|f\|_\infty\|r''\|_\infty$,
    the second term in~\eqref{eqn:q_bound_split} is bounded by
    \begin{align*}
       K_5
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))
            q_\theta(x, dy)
        \right|
        +
       K_6
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
            q_\theta(x, dy)
        \right| .
    \end{align*}
Under Condition~\ref{control through gamma} of 
Assumptions~\ref{def:model_setup}, the right hand side is bounded by 
a constant times $\smooth{\gamma}\eta(x)+(\smooth{\gamma}\eta(x))^2$ and 
$m^2\gamma(m)$ is bounded, so therefore $|\gamma B^\theta_f|$ is bounded.
\end{proof}

% % % % % % % % % % % % %
\subsection{Proof of Theorem~\ref{thm:nonlocal_convergence}: limit points for the nonlocal process}
    \label{sec:population_density_proof}

In this section we prove 
Theorem~\ref{thm:nonlocal_convergence}.
This would be implied by convergence of the lookdown process
(see \citet{kurtz/rodrigues:2011} and \citet{etheridge/kurtz:2019});
however in our setting,
because the parameters in the lookdown process
depend on the empirical distribution,
we actually use tightness of the sequence of population processes
in the proofs of tightness for the corresponding lookdown processes.

\begin{proof}[Proof of Theorem~\ref{thm:nonlocal_convergence}.]
The proof follows a familiar pattern. First we extend $\IR^d$ 
to its one-point compactification $\overline{\IR}^d$ and 
establish, in Lemma~\ref{lem:eta_compact_containment},
compact containment of the 
sequence of scaled population
processes in $\cmeasures$ 
(for which, since we have compactified $\IR^d$,
it suffices to consider the sequence of total masses); armed with this, 
%Theorem~3.9.1 of~\cite{ethier/kurtz:1986} tells us that 
tightness of the population processes in $\mathcal{D}_{[0,\infty)}(\cmeasures)$
follows from tightness of the real-valued processes $(H(\eta_t))_{t\geq 0}$ for a sufficiently large
class of test functions $H$, which we establish through an application of the
Aldous-Rebolledo criterion in Lemma~\ref{lem:eta_projections_tightness}.
These ingredients are gathered together in 
Proposition~\ref{tightness in one point compactification}
to deduce tightness of the scaled population processes in 
the larger space $\mathcal{D}_{[0,\infty)}(\cmeasures)$.

We then characterise
limit points as solutions to a martingale problem in
Lemma~\ref{lem:limit_mgale}; finally in Lemma~\ref{no mass at infty}
we check that 
in the process of passing to the limit, no mass `escaped to infinity', so that
in fact the limit points take values in $\mathcal{D}_{[0,\infty)}(\measures)$.
\end{proof}

% % % % % % % % % % % % %
\subsubsection{Compact Containment Condition}

As advertised, we work with the one-point compactification of $\IR^d$ and
consider $(\eta^N_t)_{t\geq 0}$ as a sequence of 
$\cmeasures$-valued processes. Since, for each $K>0$,
$\{\eta: \langle 1,\eta\rangle\leq K\}$
is a compact set in $\cmeasures$, we shall focus on 
controlling $(\langle 1,\eta^N_t\rangle)_{t\geq 0}$.
The key is that Assumptions~\ref{def:model_setup}
are precisely chosen to guarantee boundedness
of the net per-capita reproduction rate.

\begin{lemma}
    \label{lem:eta_f_bound}
    Under Assumptions~\ref{def:model_setup},
    for all $f \in C^2_b(\IR^d)$ with uniformly bounded first and 
	second derivatives, and all $T>0$, there exists a $C=C(f,T) < \infty$, 
	independent of $N$,
    such that
    \begin{equation}
\label{eta_f_bound}
        \IE[\langle f, \eta^N_t \rangle]
        \le
	    C\IE[\langle 1, \eta^N_0 \rangle]
        %e^{C_f t}
    \end{equation}
    for all $N \geq 1$.
\end{lemma}

\begin{proof}
Consider the semimartingale decomposition from equation~\eqref{eqn:eta_martingale}:
\begin{align} \label{eqn:eta_f_mgale_decomp}
    \langle f, \eta^N_t \rangle
    &=
    \langle f, \eta^N_0 \rangle
    + \int_0^t \int_{\IR^d} \big\{
        \gamma(x, \eta^N_s) B^\theta_f(x, \eta^N_s)
        + f(x) F(x, \eta^N_s)
        \big\} \eta^N_s(dx) ds
    + M^N_t(f) ,
\end{align}
where $M^N_t(f)$ is a martingale and $B^\theta_f$ is defined in \ref{generator Bf}.
First note that Condition~\ref{gamma_B_condition} of Assumptions~\ref{def:model_setup}
stipulates that $|\gamma B^\theta_f|$ is uniformly bounded,
and so recalling that $F$ is bounded above, we can now conclude that
under Assumptions~\ref{def:model_setup}
$\gamma(x, \eta) B^\theta_f(x, \eta) + f(x) F(x, \eta) \le C_f $ for some $C_f$.

Now, taking expectations in~(\ref{eqn:eta_f_mgale_decomp}),
    \begin{align}
\label{bound on intfdeta}
        \IE\left[ \langle f, \eta^N_t \rangle \right]
        &\le
        \IE\left[ \langle f, \eta^N_0 \rangle \right]
        + C_f \int_0^t \IE\left[ \langle 1, \eta^N_s \rangle \right] dt .
    \end{align}
The bound~(\ref{eta_f_bound}) then follows by first applying Gronwall's
inequality in the case $f=1$, which yields
\[
	\IE\big[\langle 1,\eta_t^N\rangle\big]\leq e^{Ct}\IE\big[\langle 1,\eta_0^N\rangle\big],
\]
with $C$ independent of $N$, and then substituting 
the resulting bound on $\IE\big[\langle 1,\eta_s^N\rangle\big]$ into the
expression above.
\end{proof}

With a bound on per-capita net growth rate in hand,
bounds on the expectation of the supremum of 
the total population size over a finite time interval also follow easily.

\begin{lemma}[Compact containment for the population process]
    \label{lem:eta_compact_containment}
    Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
    for each $T>0$, there exists some constant $C_T$,
independent of $N$,
    such that
    \begin{align}
        \label{eqn:eta_mass_bound}
        \IE\left[
            \sup_{0 \le t \le T}
            \langle 1, \eta^{N}_t \rangle
        \right]
        \le C_T \IE[\langle 1,\eta_0\rangle].
    \end{align}
    In particular, for any $\delta > 0$, there exists $K_{\delta}>0$ such that
    \begin{equation}
    \limsup_{N \to \infty}
        \IP \left\{ \sup_{s \in [0,T]}
            \langle 1 ,\eta^{N}_{s}\rangle
            > K_\delta \right\}
        \leq \frac{C_T}{K_{\delta}}
        < \delta.
    \end{equation}
\end{lemma}

\begin{proof}
    First note that by Lemma~\ref{lem:eta_f_bound},
    $\IE[\langle 1, \eta^N_t \rangle] \le \IE[\langle 1, \eta^N_0 \rangle] e^{Ct}$
    for some $C$ (independent of $N$).
    Now, let $M^*_t(f) = \sup_{0 \le s \le t} M_t(f)$,
    and as before let $[M(f)]_t$ be the quadratic variation of $M_t(f)$,
    so that by the Burkholder-Davis-Gundy inequality, there is a $C'$ so that,
in the notation of Lemma~\ref{lem:eta_f_bound},
    \begin{align*}
        \IE\left[ M^*_t(1) \right]
        &\le
        C' \IE\left[ \sqrt{[M(1)]_t} \right]
        \\ &\le
        C'\left( 1 + \IE\left[ [M(1)]_t \right] \right)
        \\ &=
        C'\left( 1 + \frac{\theta}{N} \IE\left[
            \int_0^t
            %\int_{\IR^d} 
	    \Big\langle\left\{
                \gamma(x, \eta^N_s)
                \int_{\IR^d} r(y, \eta^N_s) q_\theta(x, dy)
                + \mu_\theta(x, \eta^N_s)
            \right\}, \eta_s^N(dx)\Big\rangle
            ds
            \right] \right)
        \\ &=
        C'\left( 1 + \IE\left[
            \int_0^t
            %\int_{\IR^d} 
	    \Big\langle\left\{
                \frac{2\theta}{N} \gamma(x, \eta^N_s) r(x, \eta^N_s)
		+ \frac{\gamma(x,\eta_s^N)}{N} 
		B^\theta_1(x, \eta^N_s)-\frac{1}{N}F(x,\eta^N_s)
            \right\}, \eta_s^N(dx) \Big\rangle
            ds
            \right] \right) .
    \end{align*}

	We have not assumed that $F$ is bounded below, but to see that the 
	term involving $-F$ does not cause us problems, we rearrange
	equation~(\ref{eqn:eta_f_mgale_decomp}) 
	with $f=1$ to see that
	\begin{equation}
		\label{integral of -F}
\IE\left[\int_0^t \Big\langle -F(x, \eta^N_s), \eta^N_s(dx)
	\Big\rangle ds \right]
	=\IE[\langle 1, \eta^N_0\rangle] -\IE[\langle 1,\eta^N_t\rangle]
	+\IE\left[\int_0^t\Big\langle %int_{\IR^d}
\gamma(x,\eta_s^N)
	B^\theta_1(x, \eta_s^N),\eta_s^N(dx)\Big\rangle ds\right],
	\end{equation}
which is bounded since $\gamma (x,\eta)$ and $B^\theta_1(x,\eta)$ are both bounded and 
$\langle 1,\eta_t\rangle\geq 0$. 
    Since $\theta/N \to \alpha < \infty$,
    combining constants, we obtain that
    \begin{align*}
        \IE\left[ M^*_t(1) \right]
        &\le
        C' + C'' \IE[ \langle 1, \eta_0^N \rangle ] e^{tC} .
    \end{align*}
    Taking suprema and expectations on both sides of equation~\eqref{eqn:eta_f_mgale_decomp},
    then again using the fact that $\gamma(x, \eta) B^\theta_1(x, \eta) + F(x, \eta) \le C$,
    \begin{align*}
        \IE\left[\sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle \right]
        &\le
        \IE[\langle 1, \eta^N_0 \rangle]
        + \IE\left[
            \sup_{0 \le s \le t}
            \int_0^t 
	    %\int_{\IR^d} 
	    \Big\langle\left\{
                \gamma(x, \eta^N_s)
                B^\theta_1(x, \eta^N_s)
                + F(x, \eta^N_s)
            \right\}, \eta^N_s(dx) \Big\rangle ds
        \right]
        \\
 & + \IE[M^*_t(1)] 
        \\
        &\le
        \IE[\langle 1, \eta^N_0 \rangle]
        + C \IE\left[
            \int_0^t \sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle ds
        \right]
        + C' + C'' \IE[ \langle 1, \eta_0^N \rangle ] e^{tC} .
    \end{align*}
    Once again applying Gronwall's inequality,
    \begin{align*}
        \IE\left[\sup_{0 \le s \le t} \langle 1, \eta^N_s \rangle \right]
        &\le
        C''' 
        \left(1+ \IE[\langle 1, \eta^N_0 \rangle]\right) e^{2tC} .
    \end{align*}
    For any $T$,
    the inequality on the right is bounded above by a constant $C(T)$ independent of $N$.
    As a result, for any $K > 0$,
    \begin{equation*}
    \limsup_{N \to \infty}
        \IP\left[ \sup_{0 \le s \le T} \langle 1, \eta^{N}_{s} \rangle \geq K \right]
        \leq
        \frac{C(T)}{K}.
    \end{equation*}
\end{proof}

Our next task is to show tightness of 
$(\langle f,\eta_t^N\rangle)_{t\geq 0}$ for 
$f\in C_b^\infty(\overline{\IR}^d)$.

\begin{lemma}[Tightness of $(\langle f, \eta^{N}_t \rangle )_{t>0}$]
    \label{lem:eta_projections_tightness}
	For each $f \in C^{\infty}_{b}(\overline{\IR}^d)$, 
the collection of processes
$(\langle f, \eta^{N}_t \rangle)_{t \geq 0}$
for $N = 1, 2, \ldots$ is tight
as a sequence of c\`adl\`ag, real-valued processes.
\end{lemma}
\begin{proof}
The Aldous-Rebolledo criterion (found in the form we use here in \cite{rebolledo:1980}),
applied to the semimartingale representation of $\langle f, \eta^N_t\rangle$
of equation~\eqref{eqn:eta_f_mgale_decomp}, tells us
that it suffices to show that for each $T>0$,
(a) for each fixed $0\leq t\leq T$, the sequence $\{\langle f, \eta^N_t \rangle\}_{N \ge 1}$ is tight,
and (b) for any sequence of stopping times $\tau_N$ bounded by $T$,
and for each $\nu > 0$, there exist $\delta > 0$ and $N_0 > 0$ such that 
\begin{gather}
        \label{eqn:eta_projections_goal1}
    \sup_{N > N_0}
    \sup_{t \in [0, \delta]}
    \IP\left\{\left|
            \int_\tau^{\tau + t}
            \int_{\IR^d}
            \left\{
                \gamma(x, \eta^N_s) B^\theta_f(x, \eta^N_s)
                + f(x) F(x, \eta^N_s)
            \right\} 
            \eta^N_s(dx)
            ds
        \right|> \nu \right\}
        < \nu ,
    \\ \text{and} \qquad
        \label{eqn:eta_projections_goal2}
    \sup_{N > N_0}
    \sup_{t \in [0, \delta]}
    \IP\left\{\big|
        [M^{N}(f)]_{\tau + t} 
            - [M^{N}(f)]_\tau \big|
        > \nu
    \right\}
    < \nu.
\end{gather}
Tightness of $\langle f, \eta^N_t\rangle$ for fixed $t$
follows from Lemma~\ref{lem:eta_f_bound} and Markov's inequality,
so we focus on the remaining conditions.

The proof of Lemma~\ref{lem:eta_f_bound}
provides a uniform bound on $\gamma B^\theta_f$, but we only know
that $F$ is bounded above. However, since, by 
assumption, for each fixed value of
$m$, $\sup_{k\leq m}|F(x,k)|$ is uniformly bounded as a function of $x$, 
noting that $\smooth{F}\eta\leq
\langle 1,\eta\rangle\|\rho_F\|_\infty$,
and using Lemma~\ref{lem:eta_compact_containment} 
to choose $N_0$ and $K$ such that
if $N > N_0$, then
$$
    \IP\left\{\sup_{0\leq s\leq T}\langle 1, \eta_s^N\rangle\geq K\right\} < \nu,
$$
we now choose $\delta_1$ so that
$$
    \delta_1 \|f\|_\infty \sup\big\{ \sup_{x}|F(x,k)| : k\leq K\|\rho_F\|_\infty \big\}
    <
    \nu/2,
    \qquad
    \sup_{x,\eta}\gamma(x,\eta)B^\theta_f(\,\eta)\delta_1<\nu/2,
$$
so that~\eqref{eqn:eta_projections_goal1} is satisfied with $\delta = \delta_1$.

Similarly,
\begin{align*}
    &
 \big|   [M^{N}(f)]_{\tau + t} 
        - [M^{N}(f)]_\tau \big|
    \\ & \qquad =
  \Big|  \int_{\tau}^{\tau + t}
        \frac{\theta}{N}
        \int_{\IR^d}
        \left\{
            \gamma(x, \eta^N_s)
            \int_{\IR^d} f^2(y) r(y, \eta^N_s) q_\theta(x, dy)
            +
            \mu_\theta(x, \eta^N_s) f^2(x)
        \right\}
        \eta^N_s(dx)
    ds \Big|
    \\ & \qquad =
   \Big| \int_{\tau}^{\tau + t}
        \frac{\theta}{N}
        \int_{\IR^d}
        \left\{
            \gamma(x, \eta^N_s)
            \left(
                2 f^2(x) r(x, \eta^N_s)
                +
                B^\theta_{f^2}(x, \eta^N_s)
            \right)
            -
            f^2(x) \frac{F(x, \eta^N_s)}{ \theta}
        \right\}
        \eta^N_s(dx)
    ds\Big| ,
\end{align*}
and so using the fact that $\theta/N \to \alpha < \infty$,
an argument entirely analogous to 
that for \eqref{eqn:eta_projections_goal1}
yields a $\delta_2$ for which \eqref{eqn:eta_projections_goal2} is
satsified. Taking $\delta=\min\{\delta_1,\delta_2\}$, the result follows.

\end{proof}

% % % % % % % % %

We collect the results of the last two subsections into a proposition.
\begin{proposition}[Tightness of $(\eta^{N}_t)_{t \geq 0}$]
\label{tightness in one point compactification}
    The collection of measure-valued processes 
    $\{(\eta^{N}_t)_{t \geq 0}: N \geq 1\}$
    is tight in $\mathcal{D}_{[0,\infty)}(\cmeasures)$.
\end{proposition}

\begin{proof}
    Theorem 3.9.1~in \cite{ethier/kurtz:1986}
    says that if the collection of $E$-valued processes
    satisfies a compact containment condition
    (for any $\epsilon > 0$ and $T>0$, there is a compact set
    such that the processes stay within that set up to time $T$ with probability at least $1-\epsilon$),
    then the collection is relatively compact (which is equivalent to tightness
since we are working on a Polish space)
    if and only if
    $\{(f(\eta^{N}_t))_{t \geq 0}: N \geq 1\}$ is relatively compact
    for all $f$ in a dense subset of $C_b(E)$
    under the topology of uniform convergence in compact sets.

Since $\{ \nu: \langle 1, \nu \rangle \leq K \} $ is compact in $\cmeasures$,
    Lemma~\ref{lem:eta_compact_containment}
    gives compact containment.
    Lemma \ref{lem:eta_projections_tightness}
    shows that the real-valued processes $\langle f, \eta^{N}_t \rangle$
    are relatively compact for all 
$f \in \mathcal{C}^{\infty}_{b}(\overline{\IR}^d)$.
    Since by the Stone-Weierstrass theorem,
    the algebra of finite sums and products of terms of this form
    is dense in the space of bounded continuous functions on $\cmeasures$,
    we have relative compactness in $\mathcal{D}_{[0,\infty)}(\cmeasures)$.
\end{proof}

We wish to characterise the limit points of
$\{(\eta^{N}_t)_{t>0}\}_{N\geq 1}$ as solutions to a martingale problem 
with generator $\Pgen^{\infty}$ which we now identify. Most of the work
was done in Section~\ref{sec:heuristics}.
First, we record an equivalent formulation
of the martingale problems,
which were essentially laid out in Subsection~\ref{sec:population_heuristics}.
% that follows by application of Dynkin's formula.

\comment{TODO: the previous definition of $\Pgen^N$ was the following wrong thing; fixup below}
\begin{equation}
\begin{aligned}
    \text{ONLY RIGHT TO $O(1/N)$} \qquad
    \Pgen^N G_f(\eta)
    &:=
    G'(\langle f, \eta \rangle)
        \theta \bigg\langle
        \gamma(x, \eta)
            \int f(z) r(z,\eta) q_\theta(x,dz)
        -
        f(x)\mu_\theta(x, \eta),
        \eta(dx) \bigg\rangle
    \\&\qquad {}
    + \frac{1}{2} G''(\langle f, \eta \rangle)
    \frac{\theta}{N}
    \bigg\langle
        \gamma(x, \eta) \int f^2(y) r(y, \eta) q_\theta(x, dy) + f^2(x) \mu_\theta(x, \eta),
    \eta(dx) \bigg\rangle .
\end{aligned}
\end{equation}

\begin{lemma}
    \label{def: MP definition of limit}
For $G \in \mathcal{C}^{\infty}(\IR)$ with $\|G'''\|_\infty<\infty$,
and 
$f \in \mathcal{C}_{b}^\infty(\overline{\IR}^d)$, define the 
function $G_f$ by 
$G_f(\eta):=G (\langle f, \eta \rangle)$.
Let $\Pgen^N$ be the generator given by
\begin{equation}
\begin{aligned}
    \Pgen^N G_f(\eta)
    &:=
    \theta N \bigg\langle
    \gamma(x, \eta)
        \int 
        \left(G(\langle f, \eta \rangle + f(z)/N) - G(\langle f, \eta \rangle)\right)
        r(z,\eta) q_\theta(x,dz)
    \\ & \qquad \qquad {}
    +
    \left(G(\langle f, \eta \rangle - f(x)/N) - G(\langle f, \eta \rangle)\right)
    \mu_\theta(x, \eta),
    \eta(dx) \bigg\rangle .
\end{aligned}
\end{equation}
A process $(\eta^N_t)_{t \geq 0}$ 
satisfies the martingale characterization of Definition~\ref{defn:mgale_construction}
if and only if it is a solution 
to the $(\Pgen^N, \eta_0)$-martingale problem, i.e.,
if for all such test functions 
$$
    M_t :=
    G_f(\eta^N_t) - G_f(\eta^N_0)
    - \int_{0}^{t}\Pgen^N G_f(\eta^N_s)ds
$$
is a martingale (with respect to the natural $\sigma$-field).

Furthermore, let $\Pgen^\infty$ be the generator given by
\begin{equation}
    \label{eq: Limit Generator Definition}
\begin{aligned}
\Pgen^{\infty} G_f(\eta)
    &:= G'(\langle f, \eta \rangle)
                   %\times 
\left\{
                   \big\langle
                        \gamma(x, \eta)
                            \mathcal{B}\left(
                            f(\cdot) r(\cdot, \eta)
                            \right)(x)
                    +
                    f(x) F(x, \eta),
                    \eta(dx)
                    \big\rangle
                   \right\}\\
               &\qquad {}
               + \frac{\alpha}{2} G''(\langle f, \eta \rangle)
                  %\times 
\left\{
                  \big\langle
                    2\gamma\left( x, \eta \right)
                    r\left(x,\eta \right)
                    %+\mu\left(x,\eta \right)
                    f^2(x),
                    \eta (dx)
                    \big\rangle 
                  \right\}.
\end{aligned}    
\end{equation}
A process $(\eta^{\infty}_t)_{t \geq 0}$ 
satisifes the martingale characterization of equations~\ref{eqn:limiting_mgale_problem}
and~\ref{eqn:limiting_mgale_variation}
if and only if it is a solution 
to the $(\Pgen^{\infty}, \eta^\infty_0)$-martingale problem, i.e.,
if for all such test functions 
$$
    M_t :=
    G_f(\eta^{\infty}_t) - G_f(\eta^\infty_0)
    - \int_{0}^{t}\Pgen^{\infty}G_f(\eta^{\infty}_s)ds
$$
is a martingale (with respect to the natural $\sigma$-field). 
\end{lemma}

\begin{lemma}[Characterisation of limit points]
    \label{lem:limit_mgale}
Suppose that $(\eta^{N}_0)_{N\geq 1}$ converges weakly to $\eta_0$ as 
$N\to\infty$.
Then any limit point of $\{(\eta^{N}_t)_{t \geq 0}\}_{N\geq 1}$ 
in $\mathcal{D}_{[0,\infty)}(\cmeasures)$ 
is a solution to the martingale problem for 
$(\Pgen^\infty, \eta_0)$.
\end{lemma}
\begin{proof}
We use Theorem 4.8.2 in \cite{ethier/kurtz:1986}.
First observe that the set of functions
$\{G_f(\eta):= G(\langle f, \eta \rangle ),~
G \in \mathcal{C}^{\infty}(\IR), \|G'''\|_\infty<\infty,~
f \in \mathcal{C}_{b}^\infty(\overline{\IR}^d)\}$
is separating 
on $\mathcal{M}_F(\overline{\IR}^d)$.
Therefore, it suffices to show that for any $t>0$ and $\tau > 0$ that
\begin{equation}
    \label{eq: Convergence Condition}
\lim_{N \to \infty}
\mathbb{E}\left[
\left(
G_f(\eta^{N}_{t+\tau})-G_f(\eta^{N}_t)
-\int_{t}^{t+\tau}\Pgen^{\infty}G_f(\eta^{N}_s)ds
\right)
\prod_{i=1}^{k} h_i(\eta^{N}_{t_i})
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<\ldots,t_k \leq t < t+\tau$,
and 
bounded continuous functions 
$h_1,\ldots,h_k$ on $\cmeasures$.

Since $(\eta^N_t)_{t\geq 0}$ is Markov, the tower property gives that,
for each choice of $G_f$ and for each $N$,
\begin{equation}
    \label{eq: Prelimit MP Application}
\mathbb{E}\left[
\left(
G_f(\eta^{N}_{t+\tau})-G_f(\eta^{N}_t)
-\int_{t}^{t+\tau}\Pgen^{N}G_f(\eta^{N}_s)ds
\right)
\prod_{i=1}^{k}h_i(\eta^{N}_{t_i})
\right]=0
\end{equation}
for all $k\geq 0$, $0\leq t_1<t_2<...,t_k \leq t < t+\tau$,
and $h_1,...,h_k \in \mathcal{C}_{b}(\cmeasures)$.
%
Therefore, it suffices to show that 
\begin{equation}
    \label{eq: Convergence of Spatial Generator}
\lim_{N \to \infty}
\mathbb{E}\left[
\int_{t}^{t+\tau}
\left|
\Pgen^{N}G_f(\eta^{N}_s)
-\Pgen^{\infty}G_f(\eta^{N}_s)
\right|
ds
\prod_{i=1}^{k}h_i(\eta^{N}_{t_i})
\right]=0,
\end{equation}
and, again using the tower property, since the functions $h_i$ are
bounded, this will follow if
\begin{equation}
\lim_{N \to \infty}
\mathbb{E}\left[\left.
\int_{t}^{t+\tau}
\left|
\Pgen^{N}G_f(\eta^{N}_s)
-\Pgen^{\infty}G_f(\eta^{N}_s)
\right|
ds\right| {\cal F}_t\right]=0
\end{equation}
(where $\{{\cal F}_t\}_{t\geq 0}$ is the natural $\sigma$-field).

We rewrite $\Pgen^NG_f(\eta^N_s)$ using a 
Taylor series expansion up to third order for
$G\big(\langle f,\eta\rangle\pm f(y)/N\big)$ around
$G(\langle f, \eta \rangle )$.
As in Section~\ref{sec:heuristics} (except that now we are more explicit about
the error term), we find
\begin{equation} 
    \label{eq: Pre-Limit Generator Expanded}
\begin{aligned}
\Pgen^{N} G_f(\eta):=& G'(\langle f, \eta \rangle)\int_{\IR^d}  
\theta\Big\langle\int_{\IR^d} \big\{f(y)\gamma(x, \eta) r(y,\eta) 
-f(x)\mu_\theta(x, \eta)\big\}q_{\theta}(x,dy), \eta(dx)\Big\rangle\\
&+\frac{1}{2}\frac{\theta}{N}G''(\langle f, \eta \rangle)\int_{\IR^d} 
 \Big\langle\gamma(x, \eta)\int_{\IR^d} f^2(y) r(y,\eta)q_{\theta}(x,dy)+f^2(x)
\mu_\theta(x, \eta), \eta(dx)\Big\rangle\\
&+\frac{1}{6}\frac{\theta}{N^2}\Big\langle \big\{
G'''(w) \gamma(x, \eta)\int_{\IR^d} f^3(y) r(y,\eta) 
q_{\theta}(x,dy)
-G'''(v)f^3(x)\mu_\theta(x, \eta)\big\}, \eta(dx)\Big\rangle
\end{aligned}    
\end{equation}
%\normalsize
for some $w,v \in [\langle f,\eta \rangle - \|f\|_\infty/N, 
\langle f,\eta \rangle + \|f\|_\infty/N]$.

Combining with equation~\eqref{limit of mean measure equation}, and the fact
that $\mu_\theta(x,\eta) \to r(x,\eta)\gamma(x,\eta)$ as $\theta\to\infty$, we have
pointwise convergence:
\begin{equation}
\lim_{N\to \infty} |\mathcal{P}^{N}G(\langle f, \eta \rangle) 
- \mathcal{P}^{\infty}G(\langle f, \eta \rangle)| = 0 .
\end{equation}
To conclude convergence of the expectation, 
we would like to apply the Dominated Convergence 
Theorem in~(\ref{eq: Convergence of Spatial Generator}).
Recall that $f$ and $G$ and their derivatives are bounded, $\gamma(x,\eta)$ is
bounded, and, in the notation of Lemma~\ref{lem:eta_f_bound}
$\gamma(x,\eta) B^\theta_f(x,\eta)$ is bounded. Moreover $\theta/N^2\to 0$ as $N\to\infty$. Thus we can 
dominate 
$\left|
\Pgen^{N}G_f(\eta^{N}_s)
-\Pgen^{\infty}G_f(\eta^{N}_s)
\right|$ by 
a constant multiple of
$\theta \big\langle |r(x,\eta^N_s)\gamma(x,\eta^N_s)-\mu_\theta(x,\eta^N_s)|,
\eta^N_s(dx)\big\rangle$, 
which we  
recognise as $\langle |F(x,\eta^N_s)|,\eta^N_s(dx)\rangle$.
Since $F$ is bounded above, there is a constant $K$ such that $|F|\leq K-F$ so
that, exactly as in equation~(\ref{integral of -F}), 
we can check that
\[
\IE\Big[\left.\int_t^{t+\tau}\big\langle |F(x,\eta^N_s)|,\eta^N_s(dx)\big\rangle ds
\right| {\cal F}_t\Big]<\infty,
\]
which concludes our proof.
\end{proof}

The last step in the proof of Theorem~\ref{thm:nonlocal_convergence}
is to check that
any limit point $(\eta_t)_{t\geq 0}$ of $\{(\eta^N_t)_{t\geq 0}\}_{N\geq 1}$
actually takes its values in $\measures$, that is, ``no mass has escaped to infinity''.

\begin{lemma}
\label{no mass at infty}
Under the assumptions of Theorem~\ref{thm:nonlocal_convergence},
let $(\eta_t)_{t\geq 0}$ be a limit point of $\{(\eta^N_t)_{t\geq 0}\}_{N\geq 1}$. 
For any $\delta>0$, 
\[
\IP\big[\eta_t\big(\{\|x\|>R\}\big)>\delta\big]\to 0\qquad\mbox{ as }R\to\infty.
\]
\end{lemma}
\begin{proof}[Sketch]
To avoid the idea being buried in notation, we sketch the proof in the 
special case of $d=1$ and $\DG=\Delta$.
For any $f\in C_b^\infty(\IR)$,
\begin{multline}
\label{special case first moment equation}
\IE[\langle f,\eta_t\rangle] 
=\IE[\langle f,\eta_0\rangle]
\\
+\IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
\Delta \big(f(\cdot)r(\cdot,\smooth{r}\eta_s(\cdot))\big)(x)
+F(x,\smooth{F}\eta_s(x)) f(x),
\eta_s(dx)\Big\rangle ds\Big].
\end{multline} 
We take a sequence of test functions $f_n$ increasing to
$\exp(\lambda x)$, with $\lambda>0$. Having passed
to the limit,
substituted in~(\ref{special case first moment equation}) and 
expanded the term involving the Laplacian, we obtain
\begin{multline}
\label{exponential moment}
    \IE[\langle \exp(\lambda x),\eta_t\rangle]
    =
    \IE[\langle \exp(\lambda x),\eta_0\rangle]
    +
    \IE\Big[\int_0^t\Big\langle\gamma(x, \smooth{\gamma}\eta_s(x))
    \exp(\lambda x)
    \Big\{
        \lambda^2 r(x, \smooth{r}\eta_s(x))
    \\ {}
        + 2\lambda\partial_x\big(r(\cdot,\smooth{r}\eta_s(\cdot)\big)(x)
        + \partial_{xx} r(\cdot,\smooth{r}\eta_s(\cdot)\big)(x)
    \Big\} 
    + F(x,\smooth{F}\eta_s(x)),
    \eta_s(dx)\Big\rangle ds\Big] .
\end{multline} 
    \comment{alison to look at this again; check whether it depends on just bounding $B^\theta_1$}
Evidently, in much the same way as in the proof of
Lemma~\ref{lem:eta_f_bound},
our conditions on $r$ and $F$ ensure
that the integrand on the right hand side 
of~(\ref{exponential moment})
is bounded by $C\langle\exp(\lambda x),\eta_s(dx)\rangle$ for some constant
$C$. Gronwall's inequality then implies that 
$\IE[\langle\exp(\lambda x),\eta_t(dx)\rangle]$ 
is uniformly bounded for $t\in [0,T]$, and an application of Markov's inequality
tells us that for any $\delta >0$,
$\IP[\eta_t([R,\infty])>\delta]\to 0$ as $R\to\infty$. A symmetric argument
yields the corresponding bound for $\IP[\eta_t([-\infty, -R])>\delta$ and the 
result is proved.
\end{proof}

% % % % % % % % % % % % %
\subsection{Convergence of some nonlocal equations to classical PDEs}
\label{subsec:nonlocal to local}

It is natural to conjecture that when the limit of the rescaled 
population process that we obtained in the previous section solves a nonlocal
PDE, then if we further scale the kernels 
$\rho_r$, $\rho_\gamma$, and $\rho_F$ by setting 
$\rho^\epsilon(\cdot)=\rho(\cdot/\epsilon)/\epsilon$, as $\epsilon\to 0$, the 
corresponding solutions should converge to a limiting
population density that solves the corresponding ``classical'' PDE. We verify
this in two examples; in the first the nonlocal equation is a reaction-- diffusion equation
with the ``nonlocality'' only appearing in the 
reaction term; in the second the nonlocal PDE is a special 
case of a nonlinear porous medium equation. These, in particular,
capture the examples that we explored in 
Section~\ref{ancestral lineages for travelling waves}.

\subsubsection{Reaction--diffusion equation limits}
\label{two-step convergence to FKPP}

In this subsection we prove Proposition~\ref{prop:nonlocal_to_local}. {\em The conditions of the
proposition are in force throughout this subsection.} 
The proof rests on a Feynman-Kac representation.
We write $(Z_t)_{t\geq 0}$ for a diffusion with generator $\DG^*$ and denote 
its transition density by $f_t(x,y)$. The first step is a regularity
result for this density.

\begin{lemma} \label{regularityForX1}
Fix $T>0$. There exists a constant $K= K(T)>0$ such that, for any 
$x, y\in \mathbb{R}^d$ and $t \in [0,T]$,
\begin{equation} \label{eq:boundDensityXt}
\int |f_t(x,z)-f_t(y,z)| dz \leq \frac{\|x-y\|}{\sqrt{t}} K.
\end{equation}
\end{lemma}
\begin{proof}
We first use the Intermediate Value Theorem to obtain the bound
\begin{align*}
\int |f_t(x,z)-f_t(y,z)| dz & \leq \int \|x-y\| \|\nabla f_t(w,z)\| dz
\end{align*}
where $\nabla$ acts on the first coordinate only and $w$ is in the line 
segment $[x,y]$ joining $x$ to $y$. 
Under our assumptions on $b$ and $C$, equation~(1.3) 
of~\cite{sheu:1991}, gives existence of constants 
$\lambda=\lambda(T)>0$ and $K$ such that,
\[ \|\nabla f_t(w,z)\|  \leq  \frac{K }{\sqrt{t}}p_{ \lambda  t}(w,z), \]
where $p_s(x,y)$ is the Brownian transition density.
Hence, 
\begin{align*}
\int |f_t(x,z)-f_t(y,z)| dz & \leq K 
\frac{\|x-y\|}{\sqrt{t}} \int p_{\lambda  t}(w,z)dz = 
K  \frac{\|x-y\|}{\sqrt{t}},
\end{align*}
which proves the result.
\end{proof}

\begin{lemma} \label{regularityForX2}
Fix $T>0$. Let $x,y \in \mathbb{R}^d$, $t \in [0,T]$. Let us denote by 
$(Z_t^y)_{t\geq 0}$ and $(Z_t^x)_{t\geq 0}$ independent copies of the 
diffusion $(Z_t)_{t\geq 0}$ starting 
from $y$ and $x$ respectively. There exists a constant $K=K(T)>0$ such that, 
\[ \IE[\|Z_t^y-Z_t^x\|] \leq K(\sqrt{t} + \|y-x\|). \]
\end{lemma}
\begin{proof}
First we write, 
\[ \IE[\|Z_t^y-Z_t^x\|]  = \int \int \|u-v\| f_t(y,u) f_t(x,v) du dv . \]
Under our regularity assumptions on $C$, $b$, using equation~(1.2)
of~\cite{sheu:1991}, there exist constants $K$, $\lambda =\lambda (T)>0$ 
for which,
\[ f_t(y,u) \leq K  p_{\lambda  t}(y,u). \]
It then follows that,
\begin{align}
  \IE[\|Z_t^y-Z_t^x\|]  \leq \int \int \|u-v\| K^2 p_{\lambda  t}(y,u) 
p_{\lambda  t}(x,v) dv du = K^2 \IE[\|B_{\lambda  t}^y-B_{\lambda  t}^x\|], 
\label{eq:ExpDif1}
\end{align}
where $(B_t^y)_{t\geq 0}$ and $(B_t^x)_{t\geq 0}$ are 
independent Brownian motions starting at $y$ and $x$ respectively. 
Using the triangle inequality, and writing $(B^0_t)_{t\geq 0}$ for a 
Brownian motion started from the origin,
\begin{equation}
\IE[\|B_{\lambda  t}^y-B_{\lambda  t}^x\|]
	\leq \|y-x\|+\IE[\|B^0_{2\lambda t}\|] 
\leq \|y-x\|+C\sqrt{t}. \label{eq:ExpDif2}
\end{equation}
Substituting (\ref{eq:ExpDif2}) in (\ref{eq:ExpDif1}) gives the result.
\end{proof}

We use the representations of the solutions to equations \eqref{localPDE} and \eqref{nonlocalPDEv1} respectively:
\begin{align}
\varphi_t(x) &= \IE_x\Big[\varphi_0(Z_t) 
+ \int_0^t \varphi_s(Z_s)F(\varphi_s(Z_s)) ds\Big], \label{FK:varphi} \\
\varphi^\epsilon_t(x) &= \IE_x\Big[\varphi_0(Z_t) 
+ \int_0^t \varphi^\epsilon_s(Z_s)F(\rho^\epsilon_F*\varphi^\epsilon_s(Z_s)) ds\Big],
\end{align}
from which
\begin{equation} 
\varphi_t(x) - \varphi^\epsilon_t(x) = 
\IE_x\left[ \int_0^t \Big(\varphi_s(Z_s)F\big(\varphi_s(Z_s)\big)
-\varphi^\epsilon_s(Z_s)F\big(\rho_\epsilon*\varphi^\epsilon_s(Z_s)\big)\Big) ds \right] ,
\label{FK:diferencesVarphis}
\end{equation}
where $\IE_x$ denotes expectation for $Z$ with $Z_0 = x$.
The key to our proof of Proposition~\ref{prop:nonlocal_to_local}
will be to replace $F(\varphi_s(Z_s))$ by $F(\rho^\epsilon_F*\varphi_s(Z_s))$ 
in this expression. 
We achieve this through three lemmas.

First we need a uniform bound on $\varphi$ and $\varphi^\epsilon$.  
\begin{lemma} 
\label{BoundednessVarphis}
For any $T>0$ there exists $M = M(T, \Vert \varphi_0 \Vert) >0$ such that, 
for all $0 \leq t \leq T$:
\[ \max\{ \| \varphi_t(\cdot) \|_\infty, \| \varphi^\epsilon_t(\cdot) \|_\infty\} 
< M.
 \]
\end{lemma}
\begin{proof}
Using that $\varphi_0$ and $F$ are bounded above, from the
representation~(\ref{FK:varphi}), we have
\begin{align*}
\varphi_t(x) \leq \| \varphi_0 \|_\infty + 
K \IE\Big[\int_0^t \varphi_s(Z_s) ds\Big].
\end{align*}
In particular, % since $\varphi_s$ is non-negative,
\[ \| \varphi_t(\cdot) \|_\infty \leq \| \varphi_0 \|_\infty
+  K \int_0^t \| \varphi_s(\cdot) \|_\infty ds, \]
so, by Gronwall's inequality,
\[ \| \varphi_t(\cdot) \|_\infty \leq \| \varphi_0 \|_\infty \exp\left( K T \right). \]
Similarly,
$\| \varphi^\epsilon_t(\cdot) \|_{\infty} \leq \| \varphi_0 \|_\infty \exp\left( K T \right) $.
\end{proof}

We also need a continuity estimate for $\varphi$.
\begin{lemma} 
\label{ContinuityVarphi}
Let $T>0$. There exists a constant 
$K=K(T, \| \varphi_0\|_\infty)>0$ and 
$\delta_0=\delta_0(T, \| \varphi_0\|_\infty) >0$ such that 
for all $0 < \delta<\delta_0$ and $0 \leq t \leq T$,
\[ 
 \|x-y\|<\delta^3 \Rightarrow |\varphi_t(x)-\varphi_t(y)|< K\delta.
\]
\end{lemma}
\begin{proof}
First we need some notation. 
Fix $T>0$ and write $M$ for the corresponding constant from 
Proposition~\ref{BoundednessVarphis}.
Let $\| F \|_{M} = \sup_{m \in [0,M]} |F(m)|$.
We reserve $\widehat{K}$ for the constant on the right hand side of 
equation~(\ref{eq:boundDensityXt})
and $\widetilde{K}$ for the constant in Lemma~\ref{regularityForX2},
and write $K_{\varphi_0}$ for the Lipschitz constant of 
$\varphi_0$. Set 
\[ 
\delta_0=\min\Big( \frac{1}{\| F \|_M^{2}},\frac{1}{M e\big(2\| F \|_{M}+ \widehat{K}\big)}, 
\frac{1}{K_{\varphi_0} + 2 \| F \|_M M},  1\Big).
\]
In what follows we take $0< \delta<\delta_0$.

We first prove that the result holds if $t < \delta^2$. As before
let $Z_t^x$ and $Z_t^y$ be 
independent copies of the diffusion $Z_t$ starting at $x$ and $y$ respectively. 
From our representation (\ref{FK:varphi}) and 
Lemma~\ref{BoundednessVarphis}, we can write:
\begin{align*}
|\varphi_t(x)- \varphi_t(y)| 
&\leq \big|\IE_x[\varphi_0(Z_t)]-\IE_y[\varphi_0(Z_t)]\big| 
    + 2 \| F \|_M M t \\ 
& \leq \IE[|\varphi_0(Z_t^x)-\varphi_0(Z_t^y)|] 
    + 2 \| F \|_M M t  \\ 
&\leq K_{\varphi_0}  \IE[\|Z_t^x-Z_t^y\|] 
    + 2 \| F \|_M M t \\ 
& \leq \widetilde{K} K_{\varphi_0}  (\sqrt{t} + \|y-x\|) 
    + 2 \| F \|_M M t \\ 
& \leq \widetilde{K} K_{\varphi_0} (\delta + \delta^3) 
    + 2 \| F \|_M M \delta^2 
\leq (\widetilde{K} K_{\varphi_0} +1)\delta,
\end{align*}
where we have used Lemma~\ref{regularityForX2} in the fourth inequality and the definition of $\delta_0$ in the last inequality.

Suppose now that $\delta^2<t$.
We will follow the pattern in Lemma~2.2 of~\cite{penington:2017}. 
First, note that by the Feynman-Kac formula we have an alternative 
representation for $\varphi_t(x)$: for any $t'<t$,
\[ 
\varphi_t(x) = \IE_x\Big[ \varphi_{t-t'}(Z_{t'}) 
\exp\big( \int_0^{t'} F(\varphi_{t-s}(Z_s)) ds \big) \Big]. 
\]
Therefore, setting $t'=\delta^2$ and using Lemma~\ref{BoundednessVarPhis}, for all $z$,
\[ 
e^{-\delta^2\| F \|_{M}}\IE_z\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\leq  
\varphi_t(z) 
\leq e^{\delta^2\| F \|_{M}} \IE_z\big[  \varphi_{t-\delta^2}(Z_{\delta^2})\big]. 
\]
We can then deduce that
\begin{align}
\nonumber
\varphi_t(x)-\varphi_t(y) 
&\leq e^{\delta^2\| F \|_{M}}\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
- e^{-\delta^2\| F \|_{M}}\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\\ 
\nonumber
& = e^{\delta^2\| F \|_{M}}
\Big(\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]
- \IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]\Big) 
\\ 
\nonumber
&\qquad \qquad 
+ \big(e^{\delta^2\| F \|_{M}} - e^{-\delta^2\| F \|_{M}}  \big)
\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] 
\\
\label{difference}
&\leq e^{\delta^2\| F \|_{M}} 
\Big(\IE_x\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]- 
\IE_y\big[ \varphi_{t-\delta^2}(Z_{\delta^2})\big]) 
+ M\big(e^{\delta^2\| F \|_{M}}-e^{-\delta^2\| F \|_{M}}\big) .
\end{align}

To bound the differences of the expected values in the last equation 
note that, by using again Lemma~\ref{BoundednessVarphis},
\begin{align*}
\IE_x\big[ &\varphi_{t-\delta^2}(Z_{\delta^2})\big]
- \IE_y[ \varphi_{t-\delta^2}(Z_{\delta^2})\big] & \\ 
& = \int \varphi_{t-\delta^2}(z) (f_{\delta^2}(x,z)-f_{\delta^2}(y,z) ) dz  \\ 
&\leq M \int \big|f_{\delta^2}(x,z)-f_{\delta^2}(y,z) \big| dz \\ 
&\leq M \widehat{K} \frac{\|x-y\|}{\delta}  
\leq M \widehat{K} \delta^2, 
\end{align*}
where we have used Lemma~\ref{regularityForX1} and that $\|x-y\| < \delta^3$.
Substituting in~(\ref{difference}),
\begin{align*}
\varphi_t(x)-\varphi_t(y)
&\leq e^{\delta^2 \| F \|_{M}} 
\left(M \widehat{K}\delta^2 +M - M e^{-2\delta^2\| F \|_{M}}  \right) \\ 
& \leq e^{\delta^2 \| F \|_{M}}\left(M \widehat{K}\delta^2 
+ 2 M \delta^2\| F \|_{M}  \right) \\ 
&\leq e \left( M \widehat{K}+2 M \| F \|_M  \right) \delta^2 \leq  \delta,
\end{align*}
where the last two inequalities follow from the definition of $\delta$. 
Interchanging $x$ and $y$ yields the same bound for $\varphi_t(y)-\varphi_t(x)$, and the result 
follows.
\end{proof}

We proceed to control the difference between $F(\varphi)$ and 
$F(\rho^\epsilon_F*\varphi)$. 
Note first that since $\rho_F\in L^1$, 
\[
I(\epsilon):=\int_{\{\|y\|>\epsilon^{3/4}\}}\rho^\epsilon_F(y)dy
=\int_{\{\|y\|>\epsilon^{-1/4}\}}\rho_F(y)dy
\to 0\qquad\mbox{ as }
\epsilon\to 0.
\]

\begin{lemma} 
\label{ContinuityConvolution}
Let $T>0$. There exists a constant 
$C=C(T,\| \varphi_0 \|_\infty)>0$ such that, for all $0 \leq t \leq T$, 
for all $\epsilon$ small enough,
\begin{equation} 
\label{ContinuityConvolution1} 
\| \varphi_t(\cdot) - \rho^\epsilon_F*\varphi_t(\cdot) \|_\infty 
\leq C (I(\epsilon)+\epsilon^{1/4}). 
\end{equation}
Furthermore, there is a constant $\widetilde{C}(T, \| \varphi_0 \|_\infty) = \widetilde{C} >0 $ 
such that, for all $0 \leq t \leq T$,
\begin{equation} 
\label{ContinuityConvolution2} 
\| F(\varphi_t(\cdot)) - F(\rho^\epsilon_F*\varphi_t(\cdot)) \|_\infty 
\leq \widetilde{C}\big(I(\epsilon)+\epsilon^{1/4}\big). 
\end{equation}
\end{lemma}
\begin{proof}
Let $\epsilon<\delta_0^4$, with $\delta_0$ from Lemma~\ref{ContinuityVarphi}. Then,
\begin{align*}
|\varphi_t(x) - \rho^\epsilon_F*\varphi_t(x)| & \leq 
\int_{\|x-y\| > \epsilon^{3/4} } \rho^\epsilon_F(x-y)|\varphi_t(y)-\varphi_t(x)| dy  \\ 
& + \int_{\|x-y\| \leq \epsilon^{3/4} } \rho^\epsilon_F(x-y)|\varphi_t(y)-\varphi_t(x)| dy \\ 
& \leq 2 M \int_{\|x-y\| > \epsilon^{3/4}  } \rho^\epsilon_F(x-y) dy 
+ \int_{\|x-y\| \leq \epsilon^{3/4}} \rho^\epsilon_F(x-y) K \epsilon^{1/4} dy \\
&\leq 2M I(\epsilon) + K \epsilon^{1/4},
\end{align*}
where we used 
the estimates of Lemma~\ref{BoundednessVarphis}
and Lemma~\ref{ContinuityVarphi}.
This proves (\ref{ContinuityConvolution1}). For (\ref{ContinuityConvolution2}),
let $L_M$ be the (uniform) Lipschitz constant of $F$ on $[0,M]$, with
$M$ still taken from Lemma~\ref{BoundednessVarphis}. Then,
\begin{align*}
 \| F(\varphi_t(\cdot)) - F(\rho^\epsilon_F*\varphi_t(\cdot)) \|_\infty 
&\leq L_M \| \varphi_t(\cdot)) - (\rho^\epsilon_F*\varphi_t(\cdot)) \|_\infty \\ 
& \leq L_M ( 2 M I(\epsilon) + K\epsilon^{1/4}),
\end{align*}
which proves (\ref{ContinuityConvolution2}). 
\end{proof}

\begin{proof}[of Proposition \ref{prop:nonlocal_to_local}]
Let $\epsilon$ be small enough that Lemma~\ref{ContinuityConvolution} holds. 
We use the notation $\widehat{\delta}(\epsilon)$ for the quantity on the right hand side 
of~(\ref{ContinuityConvolution2}).
Then from the representation~(\ref{FK:diferencesVarphis}) 
and Lemma~\ref{ContinuityConvolution} we can write,
\begin{align*}
| &\varphi_t(x) - \varphi^\epsilon_t(x)|  \\ 
& \leq \IE_x\left[ \int_0^t\Big| \varphi_s(Z_s)
F(\rho^\epsilon_F*\varphi_s(Z_s))
-\varphi^\epsilon_s(Z_s)
F\big(\rho^\epsilon_F*\varphi^\epsilon_s(Z_s)\big)\Big| ds \right] 
+M t \widehat{\delta}(\epsilon)   \\ 
& \leq \IE_x\left[ \int_0^t \big|F(\rho^\epsilon_F*\varphi^\epsilon_s(Z_s))\big|\cdot
\big|\varphi^\epsilon_s(Z_s)-\varphi_s(Z_s)\big| ds  \right] \\ 
& \qquad +  \IE_x\left[\int_0^t |\varphi_s(Z_s)| 
\big|F(\rho^\epsilon_F*\varphi^\epsilon_s(Z_s))
-F(\rho^\epsilon_F*\varphi_s(Z_s))\big| ds \right] 
+ M t \widehat{\delta}(\epsilon)  \\ 
& \leq \| F \|_M \int_0^t \| \varphi^\epsilon_s(\cdot) 
- \varphi_s(\cdot) \|_\infty ds
  + M L_M \int_0^t \| \rho^\epsilon_F*\varphi^\epsilon_s(\cdot)
-\rho^\epsilon_F*\varphi_s(\cdot) \|_\infty ds 
+  M  t \widehat{\delta}(\epsilon)  \\ 
& \leq (\| F \|_M + M L_M) 
\int_0^t \| \varphi^\epsilon_s(\cdot) - \varphi_s(\cdot) \|_\infty ds 
+ M  t \widehat{\delta}(\epsilon),
\end{align*}
where the second inequality is the triangle inequality, 
and the third is Lemma~\ref{BoundednessVarphis}. 
An application of Gronwall's inequality then yields,
\begin{align*}
\| \varphi^\epsilon_t(\cdot)-\varphi_t(\cdot)\|_\infty 
&\leq M  t \widehat{\delta}(\epsilon)  
\exp(t (\| F \|_M + M L_M)) \\ 
& \leq  M  T \widehat{\delta}(\epsilon)  \exp(T(\| F \|_M + M L_M)),
\end{align*}
	giving the result, since $\widehat{\delta}\to 0$ as $\epsilon\to 0$.
\end{proof}

\subsubsection{Porous Medium Equation}

In this subsection we prove Proposition~\ref{nonlocalPME to PME}.
To ease notation, we drop the subscript $\gamma$ on $\rho_\gamma$ and we present 
the proof in $d=1$ (although we retain the notation $\nabla$).
It should be clear that it extends almost without change
to higher dimensions.

Recall that 
we are concerned with non-negative solutions to the equation~(\ref{mollified equation}):
\begin{equation*}
	\partial_t\psi_t^\epsilon(x) =
	\Delta\left(\psi_t^\epsilon \, \rho^\epsilon *\psi_t^\epsilon\right)(x)
	+\psi_t^\epsilon(x)\left(1-\rho^\epsilon *\psi_t^\epsilon(x) 
\right),
\end{equation*}
and 
we assume that $\rho=\zeta*\check{\zeta}$ with $\zeta$ a rapidly decreasing function. 
The example we have in mind is $\zeta$ (and therefore $\rho$) being the density of
a mean zero Gaussian random variable.  
We shall prove that under the 
assumptions of Proposition~\ref{nonlocalPME to PME},
as $\epsilon\to 0$, we have
convergence to the solution to the porous medium equation with
logistic growth:
\begin{equation}
\label{PMEold}
	\partial_t \psi_t(x)=
	\Delta\left(\psi_t^2\right)(x)
	+\psi_t(x)\left(1-\psi_t(x)\right).
\end{equation}

We work on the time interval $[0,T]$.
We will require a lower bound on $\int \psi_t^\epsilon(x)\log \psi_t^\epsilon(x) dx$ 
which we record as a lemma.
\begin{lemma}
\label{lower bound on ulogu}
	Suppose that there exists $\lambda\in (0,1)$ and $C<\infty$, both independent of
$\epsilon$, such that $\int \exp(\lambda |x|) \psi_0^\epsilon(x) dx<C$.
Then there exists
a constant $K<\infty$, independent of $\epsilon$, such that  
 $\int \psi_t^\epsilon(x)\log \psi_t^\epsilon (x)dx>-K$ for all $t\in [0,T]$. 
\end{lemma}
\begin{proof}
First observe that, since $x\log x$ is bounded below, 
$\int_{-1}^1\psi_t^\epsilon(x)\log\psi_t^\epsilon(x) dx$ is bounded below.

Now consider 
\begin{multline}
\frac{d}{dt}\int \exp(\lambda x) \psi_t^\epsilon (x)dx
	= \int \exp(\lambda x)\Delta \big(\psi_t^\epsilon \, \rho^{\epsilon}*\psi_t^\epsilon\big)(x) dx
	\\
	+ \int \exp(\lambda x)\psi_t^\epsilon(x)\big(1-\rho^{\epsilon}*\psi_t^\epsilon(x)\big) dx
    \\
	= \int (\lambda^2-1)\exp(\lambda x) \psi_t^\epsilon(x) \smooth{\epsilon} \psi_t^\epsilon(x) dx 
	+ \int \exp(\lambda x)\psi_t^\epsilon(x) dx
    \\
    \leq \int \exp(\lambda x)\psi_t^\epsilon(x) dx,
\end{multline}
and so, by Gronwall's inequality, 
$\int \exp(\lambda x)\psi_t^\epsilon (x)dx$ is uniformly bounded on $[0,T]$.
In particular, combining with the Mean Value Theorem, we find 
\begin{equation*}
\int_x^{x+1}\psi_t^\epsilon(y)dy\leq C\exp(-\lambda x),
\end{equation*}
where the constant $C$ is independent of $x\geq1$.
A fortiori, 
\begin{equation}
\label{exp decay of u} 
\int_x^{x+1}\psi_t^\epsilon(y)\ind_{\psi_t^\epsilon(y)\leq 1}dy\leq C\exp(-\lambda x).
\end{equation}
Now the function $\psi\mapsto \ind_{0\leq \psi\leq 1}\psi|\log \psi|$ is concave, and so 
using Jensen's inequality and~(\ref{exp decay of u}),
\begin{equation*}
\int_x^{x+1}\psi_t^\epsilon(y)
|\log \psi_t^\epsilon (y)|\ind_{\psi_t^\epsilon(y)\leq 1}dy
\leq C'x \exp(-\lambda x).
\end{equation*}
Evidently a symmetric argument applies for $x\leq -1$.
Summing over $x$, we find
\begin{equation*}
\int \psi_t^\epsilon (x) \log \psi_t^\epsilon (x) dx
\geq -C''\sum_{x=1}^\infty x\exp(-\lambda x)>-K>-\infty,  
\end{equation*}
as required.
\end{proof}

\begin{proof}[Proof of Proposition~\ref{nonlocalPME to PME}]

First observe that
\begin{multline*}
	\int \psi_t^\epsilon(x)\, \rho^{\epsilon}*\psi_t^\epsilon(x) dx 
=
\int\int\int \psi_t^\epsilon(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z) dz dy dx 
\\
=
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{z})
\psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z}) d\tilde{z} d\tilde{y} d \tilde{x}
	=\int\left(\zeta^{\epsilon}*\psi_t^\epsilon(x)\right)^2 dx,
\end{multline*}
where we have set $\tilde{x}=x-z$, $\tilde{y}=y-z$, $\tilde{z}=-z$.

Now note that
\begin{eqnarray*}
\frac{d}{dt}\int \psi_t^\epsilon (x) dx
	&=&
\int \Delta \big(\psi_t^\epsilon \, \rho^{\epsilon}*\psi_t^\epsilon\big)(x) dx 
	+\int \psi_t^\epsilon(x)\big(1-\rho^{\epsilon}*\psi_t^\epsilon(x)\big) dx
\\	
&=&
 \int \psi_t^\epsilon (x) d x
	-\int \big(\zeta^{\epsilon}*\psi_t^\epsilon(x)\big)^2  dx.
\end{eqnarray*}
Thus, Gronwall's inequality implies that $\int \psi_t^\epsilon(x) dx$ is uniformly bounded above in
$\epsilon$ and $t\in [0,T]$. Note that this also then gives a uniform
bound on the rate of change of $\int \psi_t^\epsilon (x) d x$, and since
we are working on $[0,T]$ this will be enough to give continuity in time of the
$L^1$ norm of the limit
when we pass to a convergent subsequence.

Now consider
\begin{eqnarray}
	\frac{d}{dt}\int \psi_t^\epsilon\log \psi_t^\epsilon dx &=&
	\int (1+\log \psi_t^\epsilon)\left[\Delta \big(\psi_t^\epsilon\,
	\rho^{\epsilon}*\psi_t^\epsilon\big)
	+\psi_t^\epsilon\big(1-\rho^{\epsilon}*\psi_t^\epsilon\big)\right]
dx
\nonumber
\\
&=&
	\int (1+\log \psi_t^\epsilon)\left[\nabla\Big(\psi_t^\epsilon\,
	\nabla(\rho^{\epsilon}*\psi_t^\epsilon)+
	\nabla \psi_t^\epsilon\, \rho^{\epsilon}*\psi_t^\epsilon\Big)
	+\psi_t^\epsilon\big(1-\rho^{\epsilon}*\psi_t^\epsilon\big)\right]  dx
\nonumber
	\\
&=&
\int \left[-\frac{\nabla \psi_t^\epsilon}{\psi_t^\epsilon}
\Big(\psi_t^\epsilon\, \nabla(\rho^{\epsilon}*\psi_t^\epsilon)
+\nabla \psi_t^\epsilon\, \rho^{\epsilon}*\psi_t^\epsilon\Big)+
(1+\log \psi_t^\epsilon)\psi_t^\epsilon(1-\rho^{\epsilon}*\psi_t^\epsilon)\right] d x
\nonumber
	\\
&=&
-\int \left(\nabla (\zeta^{\epsilon}*\psi_t^\epsilon)\right)^2  dx
-\int (\nabla \psi_t^\epsilon)^2\frac{\rho^{\epsilon}*\psi_t^\epsilon}{\psi_t^\epsilon}  dx
\nonumber
	\\
&&+\int
\left[\psi_t^\epsilon+\psi_t^\epsilon\log \psi_t^\epsilon
\big(1-\rho^{\epsilon}*\psi_t^\epsilon\big)
-\psi_t^\epsilon\, \rho^{\epsilon}*\psi_t^\epsilon\right] dx
\nonumber
	\\
&=&
-\int \left(\nabla (\zeta^{\epsilon}*\psi_t^\epsilon)\right)^2  dx
-\int (\nabla \psi_t^\epsilon)^2\frac{\rho^{\epsilon}*\psi_t^\epsilon}{\psi_t^\epsilon} dx
-\int (\zeta^{\epsilon}*\psi_t^\epsilon)^2  dx
\nonumber
	\\
&&+\int
\left[\psi_t^\epsilon+\psi_t^\epsilon\log \psi_t^\epsilon
\big(1-\rho^{\epsilon}*\psi_t^\epsilon\big)
\right] dx .
	\label{change in ulogu}
\end{eqnarray}
The first three terms are negative; and we already
saw that the $L^1$ norm
of $\psi_t^\epsilon$ is uniformly bounded.
%
Moreover, since $\psi_t^\epsilon\log \psi_t^\epsilon$ is uniformly bounded below,
$$
    -\int \psi_t^\epsilon \log \psi_t^\epsilon \, \rho^{\epsilon}*\psi_t^\epsilon dx
    \leq
    C\int \rho^{\epsilon}*\psi_t^\epsilon dx
    =
    C \int \psi_t^\epsilon dx.
$$
From this and~(\ref{change in ulogu}),
we see immediately that $\int \psi_t^\epsilon \log \psi_t^\epsilon dx$
is uniformly bounded
above in $\epsilon$ and $t\in [0,T]$. 
Combining with Lemma~\ref{lower bound on ulogu},
we deduce that
we have a uniform bound on $\int \psi_s^\epsilon(x)|\log \psi_s^\epsilon(x)| d x$.
From \eqref{change in ulogu}, this in turn means that both
$\int_0^t \int(\zeta^{\epsilon}*\psi_s^\epsilon(x))^2  dx ds$ and
$\int_0^t \int \big(\nabla (\zeta^{\epsilon}*\psi_s^\epsilon(x))\big)^2  dx ds$
are uniformly bounded in
$\epsilon$ and $t\in [0,T]$.
Now observe that
\begin{equation}
\label{total equation}
\int\Delta\left((\rho^{\epsilon}*\psi_t^\epsilon)\psi_t^\epsilon\right)\phi  dx =
\int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\,
\psi_t^\epsilon\,\nabla\phi dx +\int \rho^{\epsilon}*\psi_t^\epsilon
\,\nabla \psi_t^\epsilon
\, \nabla\phi d x.
\end{equation}
For the first term
\begin{eqnarray}
\nonumber
\int \nabla(\rho^{\epsilon}*\psi_t^\epsilon)\, \psi_t^\epsilon\, \nabla\phi  dx
&=&
\int\int\int\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z)
\psi_t^\epsilon(x)\nabla\phi(x) dz  dy  dx
\\
\nonumber
&=&
\int\int\int\nabla \psi_t^\epsilon(\tilde{x}-\tilde{y})
\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d\tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
&=&\int(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\psi_t^\epsilon\nabla\phi)\right)
d x\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\,\nabla\phi  dx
\\
\label{error 1}
&&+
\int\nabla(\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\psi_t^\epsilon\,\nabla\phi)-
\nabla\phi\,(\zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
where, as before, we have substituted $\tilde{x}=x-z$, $\tilde{y}=y-z$,
$\tilde{z}=-z$.

To control the term~(\ref{error 1})
we use the Intermediate Value Theorem to see that
\begin{multline*}
\left|\int\left[\psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)-
\nabla\phi(x)\psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
\leq C\|\Delta\phi\|_\infty\int \psi_t^\epsilon (x-y)y\zeta^{\epsilon}(y) d y.
\end{multline*}
Since $\zeta\in\mathcal{S}(\IR)$, 
the integral in this expression is
$\mathcal{O}(\epsilon)$.

Similarly, for the second term in~(\ref{total equation}),
\begin{eqnarray}
\nonumber
\int (\rho^{\epsilon}*\psi_t^\epsilon)\, \nabla \psi_t^\epsilon\, \nabla\phi dx
&=&
\int\int\int \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y-z)
\check{\zeta}^{\epsilon}(z)\nabla \psi_t^\epsilon(x)\nabla\phi(x) dz dy  dx
\\
\nonumber
&=&
\int\int\int \psi_t^\epsilon(\tilde{x}-\tilde{y})\zeta^{\epsilon}(\tilde{y})
\zeta^{\epsilon}(\tilde{z})\nabla \psi_t^\epsilon(\tilde{x}-\tilde{z})
\nabla\phi(\tilde{x}-\tilde{z}) d \tilde{z} d\tilde{y} d\tilde{x}
\\
\nonumber
	&=&\int(\zeta^{\epsilon}*\psi_t^\epsilon)\,
\left(\zeta^{\epsilon}*(\nabla \psi_t^\epsilon \, \nabla\phi)\right)  dx\\
\nonumber
&=&
\frac{1}{2}\int\nabla ((\zeta^{\epsilon}*\psi_t^\epsilon)^2)
\, \nabla\phi  dx
\\
\label{error 2}
&&+
\int (\zeta^{\epsilon}*\psi_t^\epsilon)
\left[\zeta^{\epsilon}*(\nabla \psi_t^\epsilon\, \nabla\phi)-
\nabla\phi\,(\nabla \zeta^{\epsilon}*\psi_t^\epsilon)\right] dx,
\end{eqnarray}
and~(\ref{error 2}) is controlled in the same way as~(\ref{error 1}):
using the Intermediate Value Theorem,
\begin{multline*}
\left|\int\left[\nabla \psi_t^\epsilon(x-y)\nabla\phi(x-y)\zeta^{\epsilon}(y)
-\nabla\phi(x)\nabla \psi_t^\epsilon(x-y)\zeta^{\epsilon}(y)
\right] dy\right|
\\
\leq C\|\Delta\phi\|_\infty\left|\int \nabla \psi_t^\epsilon(x-y)
y \zeta^{\epsilon}(y)
d y\right|,
\end{multline*}
which again is $\mathcal{O}(\epsilon)$.

We now have the ingredients that we need.
The calculations above yield both a uniform (in $\epsilon$) bound on
$\zeta^{\epsilon}*\psi_t^\epsilon$ in
$L^1\cap L^2\big([0,T]\times \IR\big)$, and
\begin{multline}
\int \psi_t^\epsilon(x)\phi(x) dx-\int \psi_0^\epsilon (x)\phi(x) dx
	=\int_0^t\int (\zeta^{\epsilon}*\psi_s^\epsilon(x))^2\Delta\phi(x) dx
\\
	+\int_0^t \int \zeta^{\epsilon}*\psi_s^\epsilon(x)
	\left(1- \zeta^{\epsilon}*\psi_s^\epsilon(x)\right)
\phi(x) dx
+\mathcal{O}(\epsilon)
\end{multline}
(for sufficiently regular $\phi$). Since
$\int \psi_t^\epsilon(x)\phi(x) d x-
\int \zeta^{\epsilon}*\psi_t^\epsilon (x)\phi(x) d x$ is order $\epsilon$, we see
that $\zeta^{\epsilon}*\psi^\epsilon\rightharpoonup \psi$ in $L^1$, where $\psi$ is the (unique)
solution
to equation~(\ref{PME})
and, so, therefore, does $\psi^\epsilon$. In fact, strong convergence, that is
$\int |\psi^\epsilon -\psi|\phi dx \to 0$, follows from the
uniform integrability of $\psi^\epsilon$ that we
can deduce from the uniform control of
$\int \psi^\epsilon|\log \psi^\epsilon| d x$ that we proved above.
\end{proof}

% % % % % % % % % % % % %
\section{Simultaneous scaling with interaction distance}
\label{subsec:one step convergence proof}

\comment{below $\eta_s^N$ stops having  an $N$ at some point}

In this section we prove Theorem~\ref{thm:local_convergence},
which simultaneously scales the parameters $\theta$ and $N$ and the interaction 
kernel $\rho_F$ in the special case in which $r\equiv 1\equiv\gamma$, $q_{\theta}(x,dy)$
is isotropic with zero mean, the kernel
$\rho_F$ is Gaussian, and the scaling 
limit is a reaction diffusion equation.

{\em To simplify notation, in this section we shall write 
$$\rho_\epsilon*\eta(x)=\rho_F^\epsilon*\eta(x)
=\langle p_{\epsilon^2}(x,y),\eta(dy)\rangle,$$
where $p_t(x,y)$ denotes the heat semigroup. 
The assumptions of Theorem~\ref{thm:local_convergence} will be in 
force throughout, in particular,
\begin{equation}
	\label{conditions for one step}
	\epsilon^2\theta\to\infty, \qquad\mbox{and }\frac{\theta}{N\epsilon^d}\to 0.
\end{equation}
That $N, \theta\to \infty$ and $\epsilon\to 0$ simultaneously will be 
implicit, so for example if we write $\lim_{\epsilon\to 0}$, it should be 
understood that $\theta, N\to\infty$ in such a way 
that~(\ref{conditions for one step}) is satisfied. 
Moreover, where there is no risk of confusion, except where it is helpful
for emphasis, we 
suppress dependence of $\eta$ on $N$.}

The first part of the proof mirrors that of Theorem~\ref{thm:nonlocal_convergence}:
we establish sufficient bounds on the moments of $\rho_\epsilon*\eta_t(x)$ to
establish tightness and then apply standard results on 
convergence of Markov processes from~\cite{ethier/kurtz:1986}. 
The challenge
comes in identifying the limit points. This is much more intricate than the case in 
which we do not scale the interaction kernel, as weak convergence will no longer
be sufficient to guarantee the form of the nonlinear terms in the limiting equation.
Identification of the limit will rest on regularity inherited from continuity 
estimates for a random walk with Gaussian jumps which we prove in 
Subsection~\ref{continuity for random walk}.
Various bounds on the moments of $\rho_\epsilon*\eta_t$, one consequence of 
which is tightness, are then provided in 
Subsection~\ref{bounds on rhoepsilon},
and we identify the limit points in Subsection~\ref{limit in onestep case}.

\subsection{Continuity estimates for a Gaussian random walk}
\label{continuity for random walk}

Let us write $\mathcal{L}^\theta f(x) := \theta \int (f(y)-f(x))q_\theta(x,y) dy$ 
where $q_\theta$ is a Gaussian kernel of mean $0$ and variance $1/\theta$. 
We note that $\mathcal{L}^\theta$ is the generator of a continuous (time and 
space) random walk, which makes jumps of mean $0$ and variance $1/\theta$ at
rate $\theta$.

In what follows we write $\psi_t^{\epsilon, x}(y)$ for the solution of
\begin{equation}
    \partial_t \psi_t^{\epsilon,x}
    =
    \mathcal{L}^\theta \psi_t^{\epsilon, x},
    \label{AlmostHeatEquation}
\end{equation}
with initial condition 
$\psi_0^{\epsilon,x}(y) = \rho_\epsilon(y-x) =p_{\epsilon^2}(x,y)$.

\begin{lemma} \label{PsiBoundHS}
    Fix $t>0$,
    let $(\Pi(s))_{s \ge 0}$ be a rate one Poisson process,
    and let $T(t) = \Pi(\theta t) / \theta$.
    Then
    \[
        \psi^{\epsilon,x}_t(y)
        =
        \IE\left[ p_{\epsilon^2+T(t)}(x,y)\right],
    \]
    and, moreover,
    since under our assumptions $\theta \epsilon^2$ is bounded below,
    there is a $C$ independent of $\epsilon$ or $t$ such that
    \[
        \| \psi^{\epsilon,x}_t \|_\infty
        \leq
        \frac{C}{(\epsilon^2 + t)^{d/2}} .
    \]
\end{lemma}

\begin{proof}
    The first claim is immediate from the definition of the random walk 
with generator $\mathcal{L}^\theta$.

For the second claim, first define $\tau(t) = T(t)-t$.
Since if $\tau(t) \ge -(\epsilon^2 + t)/2$,
then $1/(\epsilon^2 + T(t)) \le 2/(\epsilon^2 + t)$,
while $\epsilon^2 + T(t) \ge \epsilon^2$ always,
then partitioning over
$\{ \tau(t) \ge -(\epsilon^2 + t)/2 \}$ and its complement,
\begin{align}
   \| \psi^{\epsilon,x} \|_\infty
    &=
        \IE\left[ 
            \frac{1}{\big(2 \pi (\epsilon^2+T(t))\big)^{d/2}}
        \right] \nonumber
    \\ & \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        + 
        \frac{C}{\epsilon^d}
        \IP\left\{
            \tau(t) < - (\epsilon^2 + t) / 2
        \right\} \label{eqn:psi_infty_bound}
\end{align}
Now, observe that since $\IE[e^{-\Pi(\theta t)}] = \exp(-\theta t (1 - e^{-1}))$,
by Markov's inequality,
\begin{align}
    \IP\left\{
        \tau(t) < - \frac{\epsilon^2 + t}{2}
    \right\}
    &=
    \IP\left\{
        e^{-\Pi(\theta t)} > e^{-\theta (t - \epsilon^2) / 2}
    \right\}
\nonumber
\\
&\leq \frac{\IE[\exp\big(-\Pi(\theta t)\big)]}{\exp\big(-\theta(t-\epsilon^2)/2\big)}
\nonumber
    \\&=
    \frac{
        \exp(- \theta t(1 - e^{-1}))
    }{
        \exp(- \theta (t - \epsilon^2) / 2)
    }
\nonumber
    \\&=
    \exp\left\{ - \delta \theta t - \frac{\theta \epsilon^2}{2} \right\} ,
\label{bound for negative tau}
\end{align}
where $\delta = 1/2 - e^{-1} > 0$.
The second term in \eqref{eqn:psi_infty_bound} is therefore bounded by
\[
    C \left(1 + \frac{t}{\epsilon^2}\right)^{d/2} e^{-\delta \theta t - \epsilon^2 \theta / 2}
    \frac{1}{(\epsilon^2 + t)^{d/2}} .
\]
Now observe that the derivative (with respect to $t$) of 
$e^{-\delta \theta t - \epsilon^2 \theta / 2} (1 + t/\epsilon^2)^{d/2}$ 
is
\[
    \left( \frac{d}{2 \epsilon^2} - \left(1 + \frac{t}{\epsilon^2}\right) \delta \theta \right)
    \left( 1 + \frac{t}{\epsilon^2} \right)^{d/2 - 1} e^{-\delta \theta t - \epsilon^2 \theta / 2} ,
\]
which is negative if $\theta(\epsilon^2 + t) > d/2 \delta$.
At the maximum, $(\epsilon^2 + t) = d \delta / (2 \theta)$,
so $(1 + t/\epsilon^2) = d \delta/(2 \theta \epsilon^2)$,
and so this quantity is bounded uniformly over not only $t$ but also $\epsilon$
(since we've assumed that $\theta \epsilon^2$ is bounded below).
Therefore, the second term in \eqref{eqn:psi_infty_bound} 
is bounded by $C / (\epsilon^2 + t)^{d/2}$
(for a different constant $C$).
Substituting this into \eqref{eqn:psi_infty_bound} yields the result.
\end{proof}

We are going to approximate $\psi_t^{\epsilon,x}(\cdot)$ by $p_{\epsilon^2+t}(x, \cdot)$,
and to control the error that this introduces we 
need to control $T(t)-t$. We achieve this through a large deviations bound.

\begin{lemma}
    \label{lem:poisson_ld}
    Fix $A > e^3 - 1$ (so that $\log(1 + A) > 3$).
    In the notation of Lemma~\ref{PsiBoundHS},
    \begin{align*}
        \limsup_{\theta \to \infty}
        \frac{1}{\theta} \log
        \IP\left\{
            T(t) - t > A(\epsilon^2 + t)
        \right\}
    \le
        - (\epsilon^2 + t) \frac{A}{2}
        - \log(1 + A)\left( \frac{t + A(\epsilon^2 + t)}{2} \right) .
    \end{align*}
\end{lemma}

\begin{proof}
    We apply Cram\'er's Theorem.
    First note that
    \begin{align*}
        \IP\left\{
            T(t) - t > A(\epsilon^2 + t)
        \right\}
        =
        \IP\Big\{
            \frac{\Pi(\theta t)}{ \theta}
            >
            t + A(\epsilon^2 + t)
        \Big\} .
    \end{align*}
For convenience, suppose $\theta$ is an integer,
and so we may write $\Pi(\theta t) = Z_1(t) + \cdots + Z_\theta(t)$,
where the $Z_i(t)$ are i.i.d.~Poisson($t$) random variables.
    To apply Cram\'er's Theorem, we need the rate function for the 
Poisson($t$) distribution,
    which is
    \[
        \Lambda^*(x) = \sup_{\alpha > 0} \left[ \alpha x - \lambda(\alpha) \right],
    \]
    where $\lambda(\alpha) = \log \IE[\exp(\alpha Z_1(t))] = t(e^\alpha - 1)$.
    Now, the supremum of $\alpha x + t - te^\alpha$ is attained at $x = t e^\alpha$,
    and so
    \[
        \Lambda^*(x) = x \log(x/t) - x + t .
    \]
    Cram\'er's Theorem then gives
    \begin{align*}
        \limsup_{\theta \to \infty}
        \frac{1}{\theta} \log
        \IP\left\{
            \frac{\Pi(\theta t)}{\theta}
            > t + A(\epsilon^2 + t)
        \right\}
    &\le
        - \Lambda^*(t + A(\epsilon^2 + t))
    \\ &=
        A(\epsilon^2 + t)
        -
        \big(t + A(\epsilon^2 + t)\big) \log\left( 1 + \frac{A(\epsilon^2 + t)}{t} \right) 
    \\ &\le
        A(\epsilon^2 + t)
        -
        \big(t + A(\epsilon^2 + t)\big) \log(1 + A)
    \\ &\le
        A(\epsilon^2 + t) \left( 1 - \frac{1}{2} \log(1 + A) \right)
        -
        \frac{1}{2} \big(t + A(\epsilon^2 + t)\big) \log(1 + A)  .
    \end{align*}
    Since $(1/2) \log(1 + A) > 3/2$, from which $1 - (1/2) \log(1 + A) < - 1/2$,
    the proof is complete.
\end{proof}

As advertised, we wish to control the difference between
$\psi_t^{\epsilon,x}(y)$
and $p_{\epsilon^2 + t}(x, y)$.

\begin{lemma}
    \label{Lemma:BoundPsiHS2}
    In the notation of Lemma~\ref{PsiBoundHS},
    and with $A$ as in Lemma~\ref{lem:poisson_ld},
    \begin{multline}
\label{heat kernel estimate}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \\
\le
        \frac{C}{(\epsilon^2 \theta)^{d/2}}
        p_{2(A+1)(\epsilon^2+t)}(x, y)
        +
        C \exp(- \epsilon^2 \theta / 2)
        +
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        \IP\left\{
            \tau(t) > A(\epsilon^2 + t)
        \right\} .
    \end{multline}
\end{lemma}

\begin{proof}
    Still using the notation of Lemma~\ref{PsiBoundHS},
the three terms on the right hand side of 
equation~(\ref{heat kernel estimate})
    correspond to bounds obtained by partitioning
    into three events according to the value of $\tau(t)$.
    Let $A_1 = \{ \tau(t) < - (\epsilon^2 + t)/2 \}$,
    $A_2 = \{ \tau(t) > A(\epsilon^2 + t) \}$,
    and $A_3$ the remaining event, $\{ - (\epsilon^2 + t)/2 \le \tau(t) \le A(\epsilon^2 + t) \}$.
    Then,
    \begin{align*}
        \left|
            \psi_t^{\epsilon, x}(y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        &=
        \left|
        \IE\left[
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right]
        \right|
        \\&\le
        \IE\left[
            (1_{A_1} + 1_{A_2} + 1_{A_3})
            \left|
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \right] .
    \end{align*}
    For the first term, note that if $a < b$ then
    \begin{align*}
        |p_a(x, y) - p_b(x, y)| 
        &=
        \frac{1}{(2\pi)^{d/2}}
        \left|
            \frac{1}{a^{d/2}}
            e^{-\|x - y\|^2 / 2a}
            -
            \frac{1}{b^{d/2}}
            e^{-\|x - y\|^2 / 2b}
        \right|
        \\ &=
        \frac{1}{(2\pi a^2)^{d/2}}
            e^{-\|x - y\|^2 / 2b}
        \left|
            e^{-\|x - y\|^2 \left(\frac{1}{2a} - \frac{1}{2b}\right)}
            -
            \left(\frac{a}{b}\right)^{d/2}
        \right|
        \\ &\le
        C \left(\frac{b}{a}\right)^{d/2}
        p_b(x, y) ,
    \end{align*}
    where the inequality follows because both terms under the absolute value 
are less than 1.
Since, on the event $A_1$, $\tau(t)<0$, we can  
apply this with $a = \epsilon^2 + t + \tau(t)$ and $b = \epsilon^2 + t$,
and, using the bound~(\ref{bound for negative tau}),
    \begin{align*}
        \IE\left[
            1_{A_1} |p_a(x, y) - p_b(x, y)| 
        \right]
        & \le
            C \left(\frac{\epsilon^2 + t}{\epsilon^2}\right)^{d/2}
            p_{\epsilon^2 + t}(x, y) 
            \IP\left\{ \tau(t) < - \frac{\epsilon^2 + t}{2} \right\}
        \\ & \le
            C \frac{1}{\epsilon^d}
            \IP\left\{ \tau(t) < - \frac{\epsilon^2 + t}{2} \right\}
        \\ & \le
        C \exp\left(-\frac{\theta \epsilon^2}{2} \right) .
    \end{align*}

    For the third term, we will first collect some facts.
    Observe that on the event $A_3$,
    $\epsilon^2 + t + \tau(t)$ is between
    $(\epsilon^2 + t)/ 2$ and $(A + 1)(\epsilon^2 + t))$,
    and for any $s$ in this interval,
    \begin{align}
        p_{2s}(y)
        &\le \nonumber
        \left( \frac{
            2 (\epsilon^2 + t)(A + 1)
        }{
            \epsilon^2 + t
        } \right)^{d/2}
        p_{2(A+1)(\epsilon^2 + t)}(x, y)
        \\ &= \label{eqn:p_bounded_on_interval}
        (2(A+1))^{d/2}
        p_{2(A+1)(\epsilon^2 + t)}(x, y) .
    \end{align}
    Moreover, since $u e^{-u} \le e^{-1}$ for all $u \ge 0$,
    \begin{align}
        \frac{\|x - y\|^2}{s} p_s(x, y)
        &= \nonumber
        \frac{4}{(2 \pi s)^{d/2}}
        e^{- \frac{ \|x-y\|^2 }{ 4s }}
        \frac{\|x-y\|^2}{4s}
        e^{- \frac{ \|x-y\|^2 }{ 4s }}
        \\ &\le \label{eqn:p_deriv_term2}
        C p_{2s}(x, y) .
    \end{align}
    Now, by the Intermediate Value Theorem,
    \begin{align}
\label{deduction from IVT}
        \left|
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        &=
        \left| \tau(t) \right|
        \left| \frac{\partial p_s(x, y)}{\partial s} \right|
    \end{align}
    for some $s$ between $\epsilon^2 + t + \tau(t)$ and $\epsilon^2 + t$.
   Since 
    \begin{align*}
        \partial_ s p_s(x, y)
        &=
        \partial_s
        \left(
            \frac{1}{(2 \pi s)^{d/2}}
            \exp\left( - \frac{\|x - y\|^2}{2 s} \right)
        \right)
        \\ &=
        - \frac{d}{2s} p_s(x, y) + \frac{\|x - y\|^2}{2 s^2} p_s(x, y),
    \end{align*}
    applying the inequality~\eqref{eqn:p_deriv_term2},
    using the fact that $p_s(x, y) \le 2^{d/2} p_{2s}(x,y)$,
    and then \eqref{eqn:p_bounded_on_interval},
    we have that for any $s \in ((\epsilon^2 + t)/ 2, (A + 1)(\epsilon^2 + t))$,
    \begin{align*}
        \left| \frac{\partial}{\partial s} p_s(x, y) \right|
        \le
        \frac{C}{s} p_{2s}(x, y) 
        \le
        \frac{C}{\epsilon^2 + t} p_{2(A + 1)(\epsilon^2 + t)}(x, y) .
    \end{align*}
    Therefore, recalling that $\IE[\tau(t)^2] = t / \theta$, substituting 
into~(\ref{deduction from IVT}),
    \begin{align*}
        \IE\left[
            1_{A_3}
            \left|
                p_{\epsilon^2 + t + \tau(t)}(x, y)
                -
                p_{\epsilon^2 + t}(x, y)
            \right|
        \right]
        &\le
            \frac{C}{\epsilon^2 + t}
                p_{2(A + 1)(\epsilon^2 + t)}(x, y)
            \IE\left[|\tau(t)|\right]
        \\ &\le
            \frac{C}{\epsilon^2 + t}
                p_{2(A + 1)(\epsilon^2 + t)}(x, y)
            \IE\left[\tau(t)^2\right]^{1/2}
        \\ &=
            \left( \frac{Ct}{\theta(\epsilon^2 + t)^2} \right)^{1/2}
                p_{2(A + 1)(\epsilon^2 + t)}(x, y)
        \\ &\le
            \frac{C}{\sqrt{\theta \epsilon^2}}
                p_{2(A + 1)(\epsilon^2 + t)}(x, y) ,
    \end{align*}
    where the last inequality follows from $2 \epsilon^2 t \le (\epsilon^2 + t)^2$.

    Finally, on the event $A_2 = \{\tau(t) > A(\epsilon^2 + t)\}$,
    we simply use
    \begin{align*}
        \left|
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \le
        \frac{C}{(\epsilon^2 + t)^{d/2}} ,
    \end{align*}
    so that
    \begin{align*}
        \IE\left[
        1_{A_2}
        \left|
            p_{\epsilon^2 + t + \tau(t)}(x, y)
            -
            p_{\epsilon^2 + t}(x, y)
        \right|
        \right]
        \le
        \frac{C}{(\epsilon^2 + t)^{d/2}}
        \IP\left\{
            \tau(t) > A (\epsilon^2 + t)
        \right\} .
    \end{align*}
\end{proof}

The last result will be useful when combined with the next bound for the heat kernel.

\begin{lemma}
    \label{Lemma:ContinuityHS}
Let $s>0$, and $x, y, z\in \IR^d$. The following estimate holds:
\[
    |p_s(x,z) - p_s(y,z)|
    \leq
    \frac{C\|x-y\|}{\sqrt{s}} \left(p_{2s}(x,z) + p_{2s}(y,z)\right) ,
\]
where the constant $C$ does not depend on $x,y,z$ or $s$.
\end{lemma}

\begin{proof}

    Expanding the difference of two squares,
    \begin{align*}
        e^{- \frac{\|y-z\|^2}{2 s}}
        -
        e^{- \frac{\|x-z\|^2}{2 s}}
        &=
        \left(
            e^{-\frac{\|y-z\|^2}{4 s}}
            -
            e^{-\frac{\|x-z\|^2}{4 s}}
        \right)
        \left(
            e^{-\frac{\|y-z\|^2}{4 s}}
            +
            e^{-\frac{\|x-z\|^2}{4 s}}
        \right) .
    \end{align*}
Now, thinking of the first term in brackets as a function of a single variable on
the line segment $[y,z]$ connecting $y$ to $z$, we can apply the
Intermediate Value Theorem and take the modulus to bound this expression by 
    \[
        \|y-x\|
        \left( \frac{2\|w-z\|}{4s} \exp\left(-\frac{\|w-z\|^2}{4s}\right)\right)
        (4 \pi s)^{d/2}
        \left( p_{2s}(y,z)+p_{2s}(x,z) \right)
    \]
    for some $w\in [y,z]$.
    Using the fact that $x e^{-x^2}$ is uniformly bounded,
    we can bound the first bracket in the last equation by $C/\sqrt{s}$, and the result follows.
\end{proof}

\subsection{Moment bounds for $\rho_\epsilon*\eta$}
\label{bounds on rhoepsilon}

\begin{lemma}
\label{bounds on moments}
Let $\{{\cal F}_t\}_{t\geq 0}$ denote the natural filtration. 
Under the assumptions of Theorem~\ref{thm:local_convergence},
for each $T\in [0,\infty)$, and $k\in\IN$, there exist constants $C=C(k,T)$ and
$\widetilde{C}=\widetilde{C}(k,T)$, independent of $\epsilon$, such that for all 
$x\in\IR^d$ and all $u,t\in [0,T]$ with $u<t$,
\begin{equation}
\label{moment bound rhoepsilon}
\IE\Big[\left.\big(\rho_\epsilon*\eta_t(x)\big)^k\right| {\cal F}_u\Big]
\leq C\langle\psi_{t-u}^{\epsilon,x}(z), \eta_u(dz)\rangle^k
+C\frac{\theta}{N\epsilon^d}
\langle\psi_{t-u}^{\epsilon,x}(z), \eta_u(dz)\rangle;
\end{equation} 
and
\begin{multline}
\label{flandoli trick for rhoepsilon}
\IE\Big[\left. \int_u^t\langle\psi_{t-s}^{\epsilon,x}(z),\eta_s(dz)\rangle^{k-1}
\big\langle\psi_{t-s}^{\epsilon,x}(z)|F(\rho_\epsilon*\eta_s(z))|,
\eta_s(dz)\big\rangle ds\right| {\cal F}_u\Big]
\\
\leq\widetilde{C}\langle\psi_{t-u}^{\epsilon,x}(z), \eta_u(dz)\rangle^k
+\widetilde{C}\frac{\theta}{N\epsilon^d}
\langle\psi_{t-u}^{\epsilon,x}(z), \eta_u(dz)\rangle;
\end{multline}
where the function $\psi^{\epsilon,x}_t(\cdot)$ was
defined in~(\ref{AlmostHeatEquation}).
In particular, 
under the assumptions of Theorem~\ref{thm:local_convergence}, the
expected values of the
quantities on the right hand side of~(\ref{moment bound rhoepsilon})
and~(\ref{flandoli trick for rhoepsilon}) are both integrable 
with respect to Lebesgue measure.  
\end{lemma}
\begin{proof}
To simplify our expressions, we shall consider the case $u=0$, but
the proof goes through unchanged for other values of $u$.

We proceed by induction.

For any time-dependent function $\phi_t(x)$
with time derivative $\dot \phi_t(x) = \partial_t \phi_t(x)$,
\begin{multline}
\label{time dep mg prob}
    \langle \phi_0(x), \eta_t(dx) \rangle
    = 
    \langle \phi_t(x), \eta_0(dx) \rangle
    +
    M_t(\phi_\cdot(\cdot))
    +
    \int_0^t \big\langle \mathcal{L}^\theta \phi_{t-s}(x) 
- \dot \phi_{t-s}(x), \eta_s(dx) \big\rangle ds
    \\  {}
    +
 \int_{0}^t \big\langle \phi_{t-s}(x) F(x, \eta_s) , \eta_s(dx) \big\rangle ds ,
\end{multline}
where $M_t(\phi_\cdot(\cdot))$ is a martingale. 
Taking $\phi_t(\cdot)=\psi_t^{\epsilon,x}(\cdot)$,
\begin{eqnarray}
\nonumber
\rho_\epsilon*\eta_t(x)&=&\langle\psi^{\epsilon,x}_0(y),\eta_t(dy)\rangle
\\
&=&\langle\psi^{\epsilon,x}_t(y),\eta_0(dy)\rangle
+\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
F\big(\rho_\epsilon*\eta_s(y)\big), \eta_s(dy)\big\rangle ds
+M_t(x),
\label{expn rhoepsilon}
\end{eqnarray}
where $M_t(x)$ is a martingale. 
Taking expectations in~(\ref{expn rhoepsilon}), using that $F$ is bounded above, and 
applying Gronwall's inequality to 
$\langle\psi^{\epsilon,x}_{t-s}, \eta_s\rangle$ 
we obtain~(\ref{moment bound rhoepsilon})
in the case $k=1$. 
Moreover, rearranging~(\ref{expn rhoepsilon}) we find 
\begin{equation}
-\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
F(\rho_\epsilon*\eta_s(y), \eta_s(dy)\big\rangle ds
=\langle\psi^{\epsilon,x}_t(y),\eta_0(dy)\rangle
-\langle \psi^{\epsilon,x}_0(y),\eta_t(dy)\rangle +M_t(x),
\end{equation} 
and taking expectations again, 
since $\langle\psi^{\epsilon,x}_0,\eta_t\rangle>0$, and $M_0(x)=0$,
this yields 
$$\IE\Big[-\int_0^t\langle\psi^{\epsilon,x}_{t-s}(y)F(\rho_\epsilon*\eta_s(y), 
\eta_s(dy)\rangle ds\Big|{\cal F}_0\Big]
\leq \widehat{C}\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle.$$
Since $F$ is bounded above, there exists a constant $K$ such that $|F|\leq K-F$ and so 
combined with the bound on $\IE[\langle\psi^{\epsilon, x}_0(y),\eta_t(dy)\rangle]$ just
obtained, this in turn yields 
$$\IE\Big[\int_0^t\big\langle\psi^{\epsilon,x}_{t-s}(y)
|F(\rho_\epsilon*\eta_s(y)|, 
\eta_s(dy)\big\rangle ds\Big|{\cal F}_0\Big]\leq 
\widetilde{C}\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle,$$
which is~(\ref{flandoli trick for rhoepsilon})
in the case $k=1$.

Now suppose that we have established~(\ref{moment bound rhoepsilon}) 
and~(\ref{flandoli trick for rhoepsilon}) for all exponents $j<k$.
First we calculate the generator $\Pgen^N$ of our scaled population process applied
to functions of the form $\langle f,\eta\rangle^k$.
Recalling that each jump of the process involves the birth or death or a single 
individual, and so increments $\langle f,\eta\rangle$ by $\pm f/N$ at the location of
that individual, we find 
\begin{multline}
\label{pgen applied to kth moment}
\Pgen^N\Big(\langle f,\eta\rangle^k\Big)
=
\Big\langle
\int\theta N\Big\{k\frac{f(y)}{N}\langle f,\eta\rangle^{k-1}+
\binom{k}{2}\frac{f(y)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots
+\frac{f(y)^{k}}{N^{k}}\Big\}q_\theta(x,dy),\eta(dx)\Big\rangle
\\
+
\Big\langle\theta N\Big(1-\frac{F(\rho_\epsilon*\eta(x))}{\theta}\Big)
\Big\{-k\frac{f(x)}{N}\langle f,\eta\rangle^{k-1}+
\binom{k}{2}\frac{f(x)^2}{N^2}\langle f,\eta\rangle^{k-2}+\cdots
+\frac{(-1)^{k}f(x)^{k}}{N^{k}}\Big\},\eta(dx)\Big\rangle.
\end{multline}
Mimicking what we did above, we set $f(\cdot)=\psi^{\epsilon,x}_t(\cdot)$ and 
write
\begin{multline}
\label{kth moment for time varying function}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
=\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\IE\Big[\int_0^t\Pgen^N\big(\langle\psi^{\epsilon,x}_{t-s}(y),\eta_s(dy)\rangle^k\big) ds
\\
-\int_0^t\langle k\dot{\psi}^{\epsilon,x}_{t-s}(y), \eta_s(dy)\rangle
\big\langle \psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\big\rangle^{k-1}ds
\Big|{\cal F}_0\Big].
\end{multline}
Using that $\sup_s\|\psi^{\epsilon,x}_s(\cdot)\|_\infty=C/\epsilon^d$,
$N\epsilon^d>1$,
and our inductive
hypothesis, we find
\begin{multline*}
\IE\Big[
\Big\langle
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\int
\frac{\psi_{t-s}^{\epsilon,x}(z)^j}{N^j}
\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
q_\theta(y,dz),
\eta_s(dy)\Big\rangle ds 
\\
+
\int_0^t\theta N\sum_{j=2}^{k}\binom{k}{j}\frac{\psi_{t-s}^{\epsilon,x}(y)^j}{N^j}\langle \psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}
(-1)^j\Big(1-\frac{F(\rho_\epsilon*\eta_s(y))}{\theta} 
\Big),
\eta_s(dy)\Big\rangle ds\Big|{\cal F}_0\Big]
\\
\leq 
C\IE\Big[\Big\langle\int_0^t\sum_{j=2}^k\frac{\theta}{N\epsilon^d}
\Big(\frac{1}{(N\epsilon^d)^{j-2}}\Big)
	\Big\langle \psi^{\epsilon,x}_{t-s}(y)
\Big(2+\frac{|F(\rho_\epsilon*\eta_s(y))|}{\theta}\Big),\eta_s(dy)\Big\rangle
\langle\psi^{\epsilon,x}_{t-s},\eta_s\rangle^{k-j}ds\big|{\cal F}_0\Big]
\\
\leq
C'\frac{\theta}{N\epsilon^d}\sum_{j=1}^{k-1}
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^j
\leq
C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big).
\end{multline*}
Combining this with~(\ref{pgen applied to kth moment})
and~(\ref{kth moment for time varying function}), using once again the fact that $F$
is bounded above, we find 
\begin{multline*}
\label{kth step of induction}
\IE\Big[\langle \psi^{\epsilon,x}_0,\eta_t\rangle^k\Big|{\cal F}_0\Big]
\leq
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k
+\tilde{C}\IE\Big[\int_0^t
\langle\psi^{\epsilon, x}_{t-s}(y),\eta_s(dy)\rangle^k ds
\Big|{\cal F}_0\Big]
\\
+C''\frac{\theta}{N\epsilon^d}
\Big(\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle^k+
\langle\psi^{\epsilon, x}_t(y),\eta_0(dy)\rangle\Big)
,
\end{multline*}
and~(\ref{moment bound rhoepsilon}) 
follows from Gronwall's inequality. Rearranging exactly as in the case $k=1$, we 
recover~(\ref{flandoli trick for rhoepsilon}) 
and the inductive step is complete.
\end{proof}

We shall also need the following consequence of the bounds that we obtained in 
Lemma~\ref{bounds on moments}:
\begin{corollary}
\label{alternative form moment bounds}
Under the assumptions of Theorem~\ref{thm:local_convergence},
for each $k\geq 1$, $T>0$,
\begin{equation}
\label{double integrals with respect to eta}
\IE\Big[\big\langle (\rho_\epsilon*\eta_t)^{k},\eta_t\big\rangle \Big]
<C(k,T)<\infty, \qquad\mbox{ for all }t\in [0,T].
\end{equation}
\end{corollary}
\begin{proof}[Sketch]
First observe that if $A\in (0,1)$, then 
\begin{equation}
\label{heat equation at smaller time}
p_{A\epsilon^2}(x,y)
=\frac{1}{A^{d/2}}p_{\epsilon^2}(x,y)\exp\Big(-\frac{\|x-y\|^2}{2\epsilon^2}
\big(\frac{1}{A}-1\big)\Big)\leq \frac{1}{A^{d/2}}p_{\epsilon^2}(x,y).
\end{equation}
Now consider
    \begin{align*}
\IE\big[\langle \rho_\epsilon*\eta_t(x),\eta_t(dx)\rangle\big]&=
\IE\Big[\int\int p_{\epsilon^2}(x,z)\eta_t(dz)\eta_t(dx)\Big]
\\        
&=
\IE\Big[\int\int\int p_{\epsilon^2/2}(x,y) 
p_{\epsilon^2/2}(y,z)dy\eta_t(dz)\eta_t(dx)\Big]
        \\ &=
       \IE\Big[ \int \left( p_{\epsilon^2/2} * \eta_t(y) \right)^2 dy \Big]
\\&\leq C\int \IE\big[\big(\rho_\epsilon*\eta_t(x)\big)^2\big] dx,
    \end{align*}
where we used~(\ref{heat equation at smaller time}) in the last line.
Using Lemma~\ref{bounds on moments}
and our assumptions on $\eta_0$, this quantity is finite.

To illustrate the inductive step, now consider
\begin{multline}
\label{rewrite for inductive step}
\IE\big[\langle \rho_\epsilon*\eta_t(x)^2,\eta_t(dx)\rangle\big] 
=
\IE\Big[\int\int\int p_{\epsilon^2}(x,z_1) p_{\epsilon^2}(x,z_2)
\eta_t(dz_1)\eta_t(dz_2)\eta_t(dx)\Big]
\\
=
\IE\Big[\int\cdots\int p_{\epsilon^2/2}(x,y_1) 
p_{\epsilon^2/2}(x,y_2)p_{\epsilon^2/2}(y_1,z_1)
p_{\epsilon^2/2}(y_2,z_2)
\eta_t(dz_1)\eta_t(dz_2)dy_1dy_2\eta_t(dx)\Big].
\end{multline}
We use the identity
\[
p_{\epsilon^2/2}(x,y_1) p_{\epsilon^2/2}(x,y_2)
=p_{\epsilon^2}(y_1,y_2)p_{\epsilon^2/4}\Big(x, \frac{y_1+y_2}{2}\Big)
\]
to rewrite~(\ref{rewrite for inductive step}) as
\begin{align*}
&\IE\Big[\int p_{\epsilon^2/2}*\eta_t(y_1)\, p_{\epsilon^2/2}*\eta_t(y_2)\,
p_{\epsilon^2/4}*\eta_y\big(\frac{y_1+y_2}{2}\big)\, 
p_{\epsilon^2}(y_1,y_2)dy_1dy_2\Big]
\\
&\leq 
\IE\Big[\int\Big\{
\big(p_{\epsilon^2/2}*\eta_t(y_1)\big)^3+ \big(p_{\epsilon^2/2}*\eta_t(y_2)\big)^3+
\big(p_{\epsilon^2/4}*\eta_t\big(\frac{y_1+y_2}{2}\big)\big)^3
\Big\}p_{\epsilon^2}(y_1,y_2)dy_1dy_2\Big],
\end{align*}
where we have used that for any non-negative real numbers
$\beta_1$, $\beta_2$, $\beta_3$, 
$\beta_1\beta_2\beta_3\leq \beta_1^3+\beta_2^3+\beta_3^3$.
For the first two terms in the sum we integrate with respect to $y_2$ and $y_1$
respectively to reduce to an expression of the form considered in 
Lemma~\ref{bounds on moments}. For the final term, the change of variables
$z_1=y_1+y_2$, $z_2=y_1-y_2$ in the integral similarly 
allows us to integrate out the heat kernel, and we conclude that the result
holds for $k=2$.

We can proceed in the same way for larger values of $k$, using repeatedly that
\[
p_{t_1}(x,y_1)p_{t_2}(x,y_2)=p_{\frac{t_1t_2}{t_1+t_2}}
\Big(x,\frac{t_2y_1+t_1y_2}{t_1+t_2}\Big)p_{t_1+t_2}(y_1,y_2)
\]
to write 
\[
\prod_{j=1}^k p_{\tau}(x,x_j)
=\prod_{j=2}^{k}p_{\frac{j\tau}{j-1}}\big(x_j,X_{j-1}\big)
p_{\frac{\tau}{k}}(x, X_k)
\]
where
\[
X_1=x_1, \qquad X_j=\frac{j-1}{j}x_j+\frac{1}{j}X_{j-1}, \mbox{for }j\geq 2.
\]
This yields
\begin{multline*}
\big\langle\big(\rho_\epsilon*\eta_t(x)\big)^k,\eta_t(dx)\big\rangle
=\int\cdots\int\prod_{i=2}^kp_{\epsilon^2 i/(i-1)}(y_,X_i)
\prod_{i=1}^kp_{\epsilon^2/2}*\eta_t(y_i)p_{\epsilon^2/k}*\eta_t(X_k)
dy_1\ldots dy_k
\\
\leq
\int\cdots\int 
\prod_{i=2}^kp_{\epsilon^2 i/(i-1)}(y_i,X_i)
\Big\{\sum_{i=1}^k\big(p_{\epsilon^2/2}*\eta_t(y_i)\big)^{k+1}
+\big(p_{\epsilon^2/k}*\eta_t(X_k)\big)^{k+1}\Big\}dy_1\ldots dy_k,
\end{multline*}
and once again we can change variables in the integrals and 
use~(\ref{heat equation at smaller time}) to bound this by a constant 
multiple of $\int\IE\big[\big(\rho_\epsilon*\eta_t(x)\big)^{k+1}\big]dx$, and
the inductive step is complete.
\end{proof}

\begin{corollary}[Tightness of $\{(\rho_\epsilon*\eta_t^N(x) dx)_{t\geq 0}\}$]
\label{lem:density_tightness} 
Under the assumptions of Theorem~\ref{thm:local_convergence},
the sequence of measure valued processes $\{ \rho*\eta_t^N(x) dx \}_{t \geq 0}$
(taking values in $D([0,T], \measures)$ is tight.
\end{corollary}
\begin{proof}

First observe that the proof, from Lemma~\ref{lem:eta_compact_containment}, 
that 
$\IE[\sup_{0\leq t\leq T} \langle 1,\eta_t^N\rangle]$ is 
bounded goes through unchanged, and since 
$\langle 1, \rho_\epsilon*\eta_t^N(x)dx\rangle=\langle 1, \eta_t^N\rangle$,  
compact containment follows.

As in the nonlocal case, it suffices to prove that for 
for $T>0$, and any
$f\in C_b^\infty (\IR^d)$ with bounded second derivatives, the
sequence of real-valued processes 
$\big\{\big(\int f(x)\rho_\epsilon*\eta_t^N(x)dx\big)_{t\geq 0}\big\}_{N\geq 1}$ is tight.  
Let us temporarily write $X_f^N(t)$ for $\int f(x)\rho_\epsilon*\eta_t^N(x)dx$
and set
\[
w'\big(X^N,\delta,T\big)= \inf_{\{t_i\}}\max_i\sup_{s,t\in [t_{i-1},t_i)}
\big| X_f^N(t)-X_f^N(s)\big|,
\]
where $\{t_i\}$ ranges over all partitions of the form 
$0=t_0<t_1<\cdots <t_{n-1}<T\leq t_n$ with 
$\min_{1\leq i\leq n}(t_i-t_{i-1})>\delta$ and $n\geq 1$.
Using Corollary~3.7.4 of~\cite{ethier/kurtz:1986}, 
to prove tightness
of the sequence of real-valued processes $X_f^N$
it suffices to check compact containment of 
the sequence $\{\rho_\epsilon*\eta_t^N(x)dx\}_{N\geq 1}$ at any rational time $t$ and
that for every $\nu>0$ and $T>0$, there exists $\delta>0$ such that
\[
\limsup_{N\to\infty}\IP\big[w'\big(X_f^N,\delta,T\big)>\nu\big]<\nu.
\]
Evidently this will follow 
if we can show that this condition is satisfied when we replace the 
minimum over all partitions with mesh at least $\delta$ in the 
definition of $w'$, by the partition into
intervals of length exactly $\delta$.

We have
\begin{multline}
\label{tightness estimate}
\left|
\langle \rho_{\epsilon} * f, \eta_t^N \rangle 
- \langle \rho_{\epsilon} * f, \eta_s^N \rangle            
\right|
\leq
\left|
\int_s^t\Big\langle\theta\int\big(\rho_\epsilon*f(y)-\rho_\epsilon*f(x)\big)
q_{\theta}(x,dy),\eta_u^N(dx)\Big\rangle du
\right|
\\
+\int_s^t\Big\langle |F\big(\rho_\epsilon*\eta_u^N(x)\big)|\rho_\epsilon*f(x), 
\eta_u^N(dx)\Big\rangle du +2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|,
\end{multline}
where $\widehat{M}^N(f)$ is the martingale of~(\ref{eqn:eta_martingale})
with the test function $f$ replaced by $\rho_\epsilon*f$. 
We control each of the three terms on the right hand side separately.

By the Intermediate Value Theorem, using $T_t$ to denote the heat semigroup,
there exists $s\in (0,1/\theta)$ such that
\begin{multline*}
\left|
\theta\int\big(\rho_\epsilon*f(y)-\rho_\epsilon*f(x)\big)
q_{\theta}(x,dy) \right|
=
\left|\theta\Big(T_{\epsilon^2+1/\theta}f(x)-T_{\epsilon^2}f(x)\Big)\right|
\\
=
\left|\partial_sT_{\epsilon^2+s}f(x)\right|=\left|T_{\epsilon^2+s}\Delta f(x)\right|
\leq\|\Delta f\|_\infty.
\end{multline*}
We therefore have
\begin{multline*}
        \left|
        \langle \rho_{\epsilon} * f, \eta_t^N \rangle
        -
        \langle \rho_{\epsilon} * f, \eta_s^N \rangle
        \right|
        \leq
        \| \Delta f \|_\infty |t-s|
        \sup_{0 \leq u \leq T } \langle 1, \eta_u^N \rangle
\\
+
	\int_s^t
	\big\langle \rho_\epsilon*|f|(x)\, \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du +2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|.
    \end{multline*}
Using the Burkholder-Davis-Gundy inequality, our
moment bounds, and Markov's inequality
\begin{equation}
\label{martingale term to zero}
\limsup_{N\to\infty}\IP\Big[
2\sup_{0\leq u\leq T}|\widehat{M}^N(f)_u|
>\frac{\nu}{3}
\Big]\leq \limsup_{N\to\infty}\frac{36}{\nu^2}\IE\big[[\widehat{M}^N(f)]_T\big]
\leq \limsup_{N\to\infty}\frac{36}{\nu^2}\frac{C\theta}{N\epsilon^d} 
=0. 
\end{equation}
Now consider
\begin{multline}
	\label{square of F-integral}
\IE\Big[
	\Big(\int_s^t\big\langle\rho_\epsilon*|f|(x) \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du\Big)^2\Big]
\\
=
2\IE\Big[
	\int_s^t\big\langle\rho_\epsilon*|f|(x) \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle 
	\int_u^t\big\langle \rho_\epsilon*|f|(x)\big|F\big(\rho_\epsilon*\eta_r^N(x)\big)\big|,
\eta_r^N(dx)\big\rangle dr
du\Big].
\end{multline}
Since $F$ is polynomial, we use the approach
of Corollary~\ref{alternative form moment bounds}, the tower 
property, and 
Lemma~\ref{bounds on moments}, to bound this
in terms of sums of terms of the form
\[
\IE\Big[\int_s^t(t-u)\int\rho_\epsilon*|f|(x)\rho_\epsilon*\eta_u^N(x)^jdx
\int\rho_\epsilon*|f|(y)\rho_\epsilon*\eta_u^N(y)^kdydu\Big].
\]
Now observe that, again using Lemma~\ref{bounds on moments}, 
\begin{multline*}
	\int\int\rho_\epsilon*|f|(x)\rho_\epsilon*\eta_u^N(x)^j
	\rho_\epsilon*|f|(y)\rho_\epsilon*\eta_u^N(y)^k dx dy
	\\
	\leq
	\int\int\|f\|_\infty\rho_\epsilon*\eta_u^N(x)^{j+k}
	\rho_\epsilon*|f|(y)
	dx dy
+
	\int\int
	\rho_\epsilon*|f|(x)
	\|f\|_\infty
	\rho_\epsilon*\eta_u^N(y)^{j+k}
dx dy
	\\ \leq C\int |f|(x)dx.
\end{multline*}
Thus the quantity~(\ref{square of F-integral}) 
is bounded by $C(t-s)^2$ for some constant $C$ which we can 
take to be independent of $s$, $t$ and $\epsilon$. Markov's inequality 
then gives
\[
\IP\Big[\|f\|_\infty\int_s^t
\big\langle \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du \geq \frac{\nu}{3}\Big]
\leq C\frac{(t-s)^2}{\nu^2}.
\]
A union bound gives that
\begin{equation}
\label{union bound}
\IP\Big[\max_i\|f\|_\infty\int_{t_{i-1}}^{t_i}
\big\langle \big|F\big(\rho_\epsilon*\eta_u^N(x)\big)\big|,
\eta_u^N(dx)\big\rangle du \geq \frac{\nu}{3}\Big]
\leq C\frac{T\delta}{\nu^2}.
\end{equation}
Now using Markov's inequality, we can choose $K$ so that
\[
\IP\Big[\|\Delta f\|_\infty\, \sup_{0\leq t\leq T} \langle 1,\eta_t^N\rangle 
>K\Big]<\frac{\nu}{3}, 
\]
and so choosing $\delta$ so that $K\delta<\nu/3$ in this 
expression and $C\delta<\nu^3/3T$ in~(\ref{union bound}),
combining with~(\ref{martingale term to zero}), the result follows.
\end{proof}

\subsection{Identification of the limit}
\label{limit in onestep case}

We now turn to the identification of the limit points of the sequence 
$\big\{\big(\rho_\epsilon*\eta_t^N(x)dx\big)_{t\geq 0}\big\}_{N\geq 1}$.
We consider the terms in the semimartingale 
decomposition~(\ref{eqn:eta_martingale})
of $\langle f, \rho_\epsilon*\eta_t^N(x)dx\rangle=
\langle \rho_\epsilon*f, \eta_t^N(dx)\rangle$.
By a Taylor expansion, writing $\{T_t\}_{t\geq 0}$ for the heat semigroup,
\begin{align*}
\int_0^t \big\langle \mathcal{L}^\theta T_{\epsilon^2} f, 
	\eta_s^N(dx)\big\rangle ds 
&= \int_0^t \big\langle \frac{1}{2} \Delta T_{\epsilon^2} f(x), 
	\eta_s^N(dx) \big\rangle 
	+ \mathcal{O}\Big(\frac{1}{\theta}\Big) 
\\ &= \int_0^t \big\langle \frac{1}{2} \Delta f(x), \rho_\epsilon* \eta_s^N(x) dx 
	\big\rangle ds + \mathcal{O}\Big(\frac{1}{\theta}\Big).
\end{align*}
Thus, from weak convergence we can deduce that under our scaling,
for any (weakly) convergent subsequence $\{\rho_\epsilon*\eta^N(x)dx\}_{N\geq 1}$,
\[ 
\int_0^t \big\langle \mathcal{L}^\theta T_{\epsilon^2} f, 
\eta_s^N(dx)\big\rangle ds 
\rightarrow \int_0^t \big\langle \frac{1}{2} \Delta f(x), 
\varphi (s,x) dx \big\rangle ds. 
\]

Since we only know that for every sequence of measures satisfying the
conditions of Theorem~\ref{thm:local_convergence}
there is a {\em weakly} convergent subsequence,
the difficulty is the non-linear term in the martingale problem. 
The first step is to prove an approximation of the form:
\begin{align*}
\IE\Big[
\int_0^t\big\langle T_{\epsilon^2} f(y) F\big(\rho_\epsilon *\eta_s(y)\big), 
\eta_s(dy) \big\rangle ds\Big]
\sim 
\IE\Big[
\int_0^t\big\langle T_{\epsilon^2} f(y) F\big(\rho_\epsilon * \eta_s(y)\big), 
\rho_\epsilon * \eta_s(y) dy \big\rangle ds\Big].
\end{align*}
This is the content of Proposition~\ref{CuadraticApprox}, 
but first we need a continuity 
estimate on $\rho_\epsilon*\eta$.

We fix $\delta>0$ and 
suppose that $s>\delta$.
Let us write
\[
\widehat{\epsilon}(\delta,\epsilon,\theta):=
e^{-\epsilon^2\theta}+\frac{1}{(\epsilon^2+\delta)^{d/2}}
\frac{1}{(A+1)^{\epsilon^2\theta}}e^{-\epsilon^2\theta},
\]
and note that under the assumption that $\epsilon^2\theta\to \infty$, for 
each fixed $\delta>0$,  
$\lim_{\epsilon\to 0, \theta\to\infty}
\widehat{\epsilon}(\delta,\epsilon,\theta)=0$. 
Using the semimartingale decomposition~(\ref{expn rhoepsilon}),
and Lemmas~\ref{lem:poisson_ld} and~\ref{Lemma:BoundPsiHS2},
we have
\begin{align}
& |\rho_\epsilon*\eta_s(y)-\rho_\epsilon*\eta_s(w)| \nonumber 
= |\langle p_{\epsilon^2}(y,z)-p_{\epsilon^2}(w,z),\eta_s(dz) \rangle | 
\nonumber
\\ 
&\leq 
\big\langle |p_{\epsilon^2+s}(y,z)-p_{\epsilon^2+s}(w,z)|, 
\eta_0(dz)\big\rangle \nonumber 
+\frac{C}{(\theta\epsilon^2)^{d/2}}\big\langle 
p_{2(A+1)(\epsilon^2+s)}(y,z)+p_{2(A+1)(\epsilon^2+s)}(w,z), \eta_0(dz)
\big\rangle
\nonumber
\\ 
&+ \int_{s-\delta}^s \big\langle 
|\psi_{s-r}^{\epsilon,y}(z) + \psi_{s-r}^{\epsilon, w}(z) |
\cdot |F(\rho_\epsilon*\eta_r(z)|, 
\eta_r(dz) \big\rangle  dr
\nonumber
\\
&+
\int_0^{s-\delta}
\big\langle |p_{s-r+\epsilon^2}(y,z)-p_{s-r+\epsilon^2}(w,z)| 
\cdot |F(\rho_\epsilon*\eta_r(z)|, 
\eta_r(dz) \big\rangle  dr
\nonumber
\\
&+
\frac{C}{(\epsilon^2\theta)^{d/2}}
\int_0^{s-\delta}
\big\langle |p_{2(A+1)(s-r+\epsilon^2)}(y,z)+
p_{2(A+1)(s-r+\epsilon^2)}(w,z)|
\cdot |F(\rho_\epsilon*\eta_r(z)|, 
\eta_r(dz) \big\rangle  dr
\nonumber
\\
&
+|M_s(y)|+|M_s(w)| 
+\mathcal{O}\big(\widehat{\epsilon}(\delta,\epsilon,\theta)
\big)  
\nonumber
\\
&
\leq \Big\langle \frac{\|y-w\|}{\sqrt{s+\epsilon^2}}
\big(p_{2(s+\epsilon^2}(y,z)+p_{2(s+\epsilon^2}(w,z)\big),\eta_0(dz)\Big\rangle
\nonumber
\\
&+
\int_0^{s-\delta}\Big\langle\frac{\|y-w\|}{\sqrt{s-r+\epsilon^2}}
\big(p_{2(s-t+\epsilon^2}(y,z)+p_{2(s-r+\epsilon^2}(w,z)\big)
\cdot |F(\rho_\epsilon*\eta_r(z)|, 
\eta_r(dz)\Big\rangle dr
\nonumber
\\
&
+\frac{C}{(\theta\epsilon^2)^{d/2}}\big\langle 
p_{2(A+1)(\epsilon^2+s)}(y,z)+p_{2(A+1)(\epsilon^2+s)}(w,z), \eta_0(dz)
\big\rangle
\nonumber
\\
&+
\frac{C}{(\epsilon^2\theta)^{d/2}}
\int_0^{s-\delta}
\big\langle |p_{2(A+1)(s-r+\epsilon^2)}(y,z)+
p_{2(A+1)(s-r+\epsilon^2)}(w,z)| 
\cdot |F(\rho_\epsilon*\eta_r(z)|, 
\eta_r(dz) \big\rangle  dr
\nonumber
\\
&
+|M_s(y)|+|M_s(w)| 
+\mathcal{O}\big(\widehat{\epsilon}(\delta,\epsilon,\theta)
\big) . 
\label{difference in rhos}
\end{align} 
Although this expression is lengthy, when we need to control its 
contribution to expressions below, it will really just be a matter of
applying the tower property and our moment bounds from 
Lemma~\ref{bounds on moments} and 
Corollary~\ref{alternative form moment bounds}.

\comment{perhaps explain: the next proposition is useful because we'll want to use weak convergence of $\smooth{\epsilon}\eta_s$; but in the martingale we have the term on the left in the proposition; probably there are other ways to do this.}
\comment{ACTUALLY BETTER we want to transfer $T_\epsilon^2$ over onto $\eta$; this happens in two steps, first this proposition; next using that $T_{\epsilon^2}f - f$ is small}

\begin{proposition} 
\label{CuadraticApprox}
Under the conditions of Theorem~\ref{thm:local_convergence},
\begin{equation}
\label{replacing nonlinear term}
\lim_{\epsilon\to 0}\Big|\int_0^t \IE\Big[\big\langle T_{\epsilon^2} f(y) 
F(\rho_\epsilon *\eta_s(y)), \eta_s(y) \big\rangle 
- \big\langle T_{\epsilon^2} f(y) 
F(\rho_\epsilon * \eta_s(y)), \rho_\epsilon * \eta_s(y) dy 
\big\rangle\Big]ds \Big|
=0.
\end{equation}
\end{proposition}

\begin{proof}
In fact we are going to fix $\delta>0$, and show that
the expression on the left hand side of~\eqref{replacing nonlinear term} is
less than a constant times $\delta$, with a constant independent of $\delta$,
$N$, and $\epsilon$. Since $\delta$ is arbitrary, the result will follow.

We first note that,
\begin{align*}
    &
\langle T_{\epsilon^2} f(y) F(\rho_\epsilon * \eta_s(y)), 
\rho_\epsilon * \eta_s(y) dy \rangle 
-\langle T_{\epsilon^2} f(y) F(\rho_\epsilon *\eta_s(dy)), \eta_s(dy) \rangle 
\\ 
& = 
\langle \int T_{\epsilon^2} f(y) 
F(\rho_\epsilon * \eta_s(y))\rho_\epsilon(y-w) dy, \eta_s(dw)\rangle - \langle 
T_{\epsilon^2} f(w) F(\rho_\epsilon*\eta_s(w)), \eta_s(dw) \rangle 
\\ 
&= 
\langle \int \big\{ T_{\epsilon^2} f(y) F(\rho_\epsilon*\eta_s(y))-T_{\epsilon^2}f(w) 
F(\rho_\epsilon*\eta_s(w))\big\} 
\rho_\epsilon(w-y)dy, \eta_s(dw) \rangle.
\end{align*}
Let us denote the integral against $dy$ in the last expression by $I$, that is 
\[ I :=  \int \{ T_{\epsilon^2} f(y) 
F(\rho_\epsilon*\eta_s(y))-T_{\epsilon^2}f(w) 
F(\rho_\epsilon*\eta_s(w))\} \rho_\epsilon(w-y)dy, \]
and note that $|I|$ is bounded by
\begin{align}
& \int 
\Big\{ |F(\rho_\epsilon*\eta_s(y))-F(\rho_\epsilon*\eta_s(w))|T_{\epsilon^2}f(y)+
F(\rho_\epsilon*\eta_s(w))|T_{\epsilon^2}f(y)-T_{\epsilon^2}f(w))|\Big\} 
\rho_{\epsilon}(w-y)dy
\nonumber \\ &\leq \int \| f\|_{\infty} 
\big|F(\rho_\epsilon*\eta_s(y))-F(\rho_\epsilon*\eta_s(w))\big| \rho_\epsilon(w-y)dy + 
C \epsilon\|f' \|_\infty |F(\rho_\epsilon*\eta_s(w))|, \label{FirstBoundCT}
\end{align}
where we have used that 
\[
\int |T_{\epsilon^2}f(y)-T_{\epsilon^2}f(w)| p_{\epsilon^2}(w,y)dy
\leq \|f'\|_\infty\int|y-w|p_{\epsilon^2}(w,y)dy.
\]
Now recall that $F$ is a polynomial of degree $n$, and so 
there exist real numbers $b_k$ such
that $F(a)-F(b)= (a-b)\sum_{k=1}^{n-1}b_kx^ky^{n-1-k}$ and so 
\[
|F(\rho_\epsilon*\eta_s(y))-F(\rho_\epsilon*\eta_s(w))|
\leq 
|\rho_\epsilon*\eta_s(y)-\rho_\epsilon*\eta_s(w)|
\sum_{k=1}^{n-1}|b_k|\left(\rho_\epsilon*\eta_s(y)^{n-1}+\rho_\epsilon*\eta_s(w)^{n-1}\right).
\]

Combining the above, we have reduced the problem to showing that 
\begin{equation}
\label{target expression}
	\lim_{\epsilon\to 0}
	\int_0^t\big\langle \int |\rho_\epsilon*\eta_s(y)-\rho_\epsilon*\eta_s(w)|
\left(\rho_\epsilon*\eta_s(y))^k+\rho_\epsilon*\eta_s(w)^k\right)
p_{\epsilon^2}(w,y)dy, \eta_s(dw)\big\rangle ds
=0,
\end{equation}
\comment{which we prove by showing the expectation of this positive quantity goes to zero; e.g. see Lemma 7.5(?)}.

We are going to use the estimate~(\ref{difference in rhos}).
First note that by Lemma~\ref{bounds on moments}
the contribution to~(\ref{target expression}) from the integral over
the time interval $[0,\delta]$ is ${\mathcal O}(\delta)$. We focus instead
on the interval $(\delta, t]$.

The first term in~(\ref{difference in rhos}) gives
\begin{multline*}
\int_\delta^t\IE\Big[\Big\langle\int
	\big\langle\frac{\|y-w\|}{\sqrt{s+\epsilon^2}}
\big(p_{2(s+\epsilon^2)}(y,z)+p_{2(s+\epsilon^2)}(w,z)\big),\eta_o(dz)\big\rangle
\\
\big(\rho_\epsilon*\eta_s(y)^k+\rho_\epsilon*\eta_s(w)^k\big)
p_{\epsilon^2}(w,y)dy, \eta_s(dw)\Big\rangle\Big] ds.
\end{multline*}
We ``borrow'' from the exponential term in $p_\epsilon^2(w,y)$ to bound
this by
\begin{multline}
C\int_\delta^t\frac{\epsilon}{\sqrt{s+\epsilon^2}}
\IE\Big[\Big\langle\int\big\langle
\big(p_{2(s+\epsilon^2)}(y,z)+p_{2(s+\epsilon^2)}(w,z)\big),\eta_0(dz)
\big\rangle
\\
\big(\rho_\epsilon*\eta_s(y)^k+\rho_\epsilon*\eta_s(w)^k\big)
p_{2\epsilon^2}(w,y)dy, \eta_s(dw)\Big\rangle ds\Big] .
\label{contribution from initial condition terms}
\end{multline}
The four terms in the product are taken separately, according to the
combinations of $w$ and $y$ appearing. First,
\[
\int_\delta^t\frac{\epsilon}{\sqrt{s+\epsilon^2}}
\IE\Big[\Big\langle\int\big\langle
p_{2(s+\epsilon^2)}(y,z)   
,\eta_0(dz)
\big\rangle
\rho_\epsilon*\eta_s(y)^k 
p_{2\epsilon^2}(w,y)dy, \eta_s(dw)\Big\rangle  \Big] ds
\]
can be rewritten as
\begin{multline*}
\int_\delta^t\frac{\epsilon}{\sqrt{s+\epsilon^2}}
\IE\Big[\int\int\big\langle
p_{2(s+\epsilon^2)}(y,z)  
,\eta_0(dz)
\big\rangle
\rho_\epsilon*\eta_s(y)^k 
p_{\epsilon^2}(x,y) \rho_\epsilon*\eta_s(x) dy dx ds \Big]
\\
\leq
\int_\delta^t\frac{\epsilon}{\sqrt{s+\epsilon^2}}
\IE\Big[\int\int\big\langle
p_{2(s+\epsilon^2)}(y,z)  
,\eta_0(dz)
\big\rangle
\big(\rho_\epsilon*\eta_s(y)^{k+1} +\rho_\epsilon*\eta_s(x)^{k+1}\big)\Big]
p_{\epsilon^2}(x,y) dy dx ds,
\end{multline*}
and using Lemma~\ref{bounds on moments} and the tower property, and integrating
with respect to $s$, under our assumptions on $\eta_0$,
this is 
bounded by
\begin{multline*}
C\epsilon \int_\delta^t\frac{1}{\sqrt{s+\epsilon^2}}
\IE\Big[\int\int\big\langle
p_{2(s+\epsilon^2)}(y,z)  
,\eta_0(dz)
\big\rangle
\\
\big(\rho_\epsilon*\eta_0(y)+\rho_\epsilon*\eta_0(y)^{k+1} 
+\rho_\epsilon*\eta_0(x)
+\rho_\epsilon*\eta_0(x)^{k+1}\big)\Big]
p_{\epsilon^2}(x,y) dy dx ds 
={\mathcal O}(\epsilon).
\end{multline*}
The term involving 
$ 
\big\langle
p_{2(s+\epsilon^2)}(w,z)   
,\eta_0(dz)
\big\rangle
\rho_\epsilon*\eta_s(w)^k
$ 
is handled similarly.

On the other hand
\begin{multline*}
\big\langle\langle p_{2(s+\epsilon^2)}(y,z), \eta_0(dz)\rangle
\rho_\epsilon*\eta_s(w)^kp_{2\epsilon^2}(w,y)dy, \eta_s(dw)\big\rangle
\\
\leq \frac{C}{(s+\epsilon^2)^{d/2}}\langle 1,\eta_0\rangle\big\langle
\rho_\epsilon*\eta_s(w)^k,\eta_s(dw)\big\rangle,
\end{multline*}
and since $\langle 1,\eta_0\rangle$ is uniformly bounded we apply 
Corollary~\ref{alternative form moment bounds} to obtain a bound on
the contribution to~(\ref{contribution from initial condition terms})
from this term of the form
\[C\epsilon\int_\delta^t\frac{1}{(s+\epsilon^2)^{d/2}}ds.\]
Note that for fixed $\delta$, this bound tends to zero as $\epsilon\to 0$.

Now consider the contribution to he left hand side of~(\ref{target expression})
from the second term in~(\ref{difference in rhos}). Since $F$ is a polynomial,
it is bounded by a sum of terms of the form
\begin{multline*}
\int_\delta^t\int_0^{s-\delta}\Big\langle\int\frac{\|y-w\|}{\sqrt{s-r+\epsilon^2}}
\big\langle\big(p_{2(s-r+\epsilon^2)}(y,z)+ p_{2(s-r+\epsilon^2)}(w,z)\big)
\rho_\epsilon*\eta_r(z)^j,\eta_r(dz)\big\rangle dr
\\
\rho_\epsilon*\eta_s(y)^kp_{\epsilon^2}(y,w)dy,\eta_s(dw)
\Big\rangle ds
\\
\leq C\epsilon
\int_\delta^t\int_0^{s-\delta}
	\frac{1}{\sqrt{s-r+\epsilon^2}}
	\Big\langle\int
\big\langle
\big(p_{2(s-r+\epsilon^2)}(y,z)+ p_{2(s-r+\epsilon^2)}(w,z)\big)
\rho_\epsilon*\eta_r(z)^j,\eta_r(dz)\big\rangle dr
\\
\rho_\epsilon*\eta_s(y)^kp_{2\epsilon^2}(y,w)dy,\eta_s(dw)
\Big\rangle ds,
\end{multline*}
where as usual we have ``borrowed'' from the exponential term in
$p_\epsilon^2(y,w)$ to replace $\|y-w\|$ by a constant times $\epsilon$.

Once again, our approach is to rearrange terms so that we can apply
Lemma~\ref{bounds on moments}
or 
Corollary~\ref{alternative form moment bounds} to obtain a bound on
the contribution to~(\ref{target expression})
from these terms of the form $C\epsilon$ (where $C$ may depend on $\delta$
but not $\epsilon$).

For example, using the Chapman-Kolmogorov equation to rewrite
\[
\int\Big\langle 
\big\langle p_{2(s-r+\epsilon^2)}(y,z)\rho_\epsilon*\eta_r(z)^j, \eta_r(dz)
\big\rangle
\rho_\epsilon*\eta_s(y)^k p_{2\epsilon^2}(y,w), \eta_s(dw)\Big\rangle dy
\]
as 
\[
\int\int
\big\langle p_{2(s-r+\epsilon^2)}(y,z)\rho_\epsilon*\eta_r(z)^j, \eta_r(dz)
\big\rangle
\rho_\epsilon*\eta_s(y)^k p_{\epsilon^2}(y,x)\rho_\epsilon*\eta_s(x) dx dy,
\]
and using Lemma~\ref{bounds on moments} and the tower property, we are led to 
control terms of the form
\[
\IE\Big[\int\big\langle p_{2(s-r+\epsilon^2)}(y,z)\rho_\epsilon*\eta_r(z)^j, 
\eta_r(dz)\big\rangle \rho_\epsilon*\eta_r(y)^{k+1} dy\Big].
\]
This, in turn, is at most
\begin{multline*}
\IE\Big[\big\langle \rho_\epsilon*\eta_r(z)^{j+k+1}, \eta_r(dz)\rangle\Big]
+\IE\Big[\int\int p_{2(s-r)}(y,x)\rho_\epsilon*\eta_r(x)\rho_\epsilon*\eta_r(y)^{k+1} dy dx\Big]
\\
\leq
\IE\Big[\big\langle \rho_\epsilon*\eta_r(z)^{j+k+1}, \eta_r(dz)\rangle\Big]
+2\int\IE\Big[\rho_\epsilon*\eta_r(x)^{k+2}\Big]dx,
\end{multline*} 
which is bounded by Lemma~\ref{bounds on moments}.

We now turn to the contribution arising from the martingale terms 
in~(\ref{difference in rhos}):
\[
\IE\Big[
\int_\delta^t\Big\langle\int 
\big(|M_s(y)|+|M_s(w)|\big)\big(\rho_\epsilon*\eta_s(y)^k+
\rho_\epsilon*\eta_s(w)^k\big)p_{\epsilon^2}(w,y)dy, \eta_s(dw)\Big\rangle ds
\Big] .
\]
To deal with the two terms in the produce involving $|M_s(y)|$, we write 
the first as $\int |M_s(y)|\rho_\epsilon*\eta_s(y)^{k+1}dy$ and then use
H\"older's inequality, Lemma~\ref{bounds on moments}, and the fact
that $\IE\big[|M_s(y)|^2\big]$ is ${\mathcal O}\big(\theta/(N\epsilon^d)\big)$
to see that the contribution from this term tends to zero in the limit.
For the second, we use the idea of the proof of 
Corollary~\ref{alternative form moment bounds} to reduce to a form to which
we can apply H\"older's inequality.

\comment{Probably should put a bit more detail in here?}
\comment{key bit is that we can always write $M_s(w)$ as a convolution of $p_{\epsilon^2/}$ with something}

Control of the terms arising from approximating $\psi^{\epsilon,x}$
by the heat kernel follows in an entirely analogous way.

Combining the above, we see that given $\delta>0$, 
\[
\lim_{\epsilon\to 0}\Big|\int_0^t \IE\Big[\big\langle T_{\epsilon^2} f(y) 
F(\rho_\epsilon *\eta_s(y)), \eta_s(y) \big\rangle 
- \big\langle T_{\epsilon^2} f(y) 
F(\rho_\epsilon * \eta_s(y)), \rho_\epsilon * \eta_s(y) dy 
\big\rangle\Big]ds\Big|
<C\delta,
\]
where the constant $C$ is independent of $\delta$.
Since $\delta$ was arbitrary, the proof is complete.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To complete the characterisation of the limit, it remains to show that if we
take a convergent subsequence 
$\big\{\big(\rho_\epsilon*\eta_t^N(dx)\big)_{t\geq 0}\big\}$ 
converging to a limit point $\big(\varphi(t,x)dx\big)_{t\geq 0}$, then
\[ 
\int_0^t
\int f(x) \rho_\epsilon * \eta_s(x) F(\rho_\epsilon * \eta_s^N(x)) dx ds 
\rightarrow \int_0^t\int f(x) \varphi(s,x) F(\varphi(s,x)) dx ds. 
\]
Since $F$ is a polynomial, we consider powers of $\rho_\epsilon*\eta$.
To illustrate the approach, we first prove that
\begin{equation} 
\label{eq:QuadConvergence} 
\int_0^t\int f(x) 
\rho_\epsilon * \eta_s^N(x)^2 dx ds
\rightarrow \int_0^t \int f(x) 
\varphi(s,x)^2 dx ds. 
\end{equation}
The converge of higher powers will follow in an entirely analogous manner, but
with more complex expressions.

The approach is standard. We fix $\tau>0$ and, in keeping with
our notation $\rho_\epsilon$, in this subsection, use $\rho_\tau$ to 
denote the symmetric Gaussian kernel with variance parameter $\tau^2$.
Our strategy is to show that, up to an error that tends
to zero as $\tau \to 0$, 
\begin{equation}
\label{approx2}
\int_0^t\int f(z) \rho_\epsilon*\eta_s(z)^2 dz ds
\sim 
\int_0^t\int \int f(z) (\rho_\epsilon * \eta_s)(z) 
\rho_\tau (z-y) (\rho_\epsilon * \eta_s)(y) dz dy ds.
\end{equation}
Analogously, also up to an error that vanishes as $\tau \to 0$,
\begin{equation}
\label{approx3}
\int_0^t\int f(z) \varphi(s,z)^2 dz ds
\sim 
\int_0^t\int \int f(z) \varphi(s,z) 
\rho_\tau (z-y) \varphi(s,y) dz dy ds.
\end{equation}
On the other hand,
weak convergence of $\rho_\epsilon * \eta$ 
(plus continuity of the mapping $(z,y) \rightarrow f(z) \rho_\tau(z-y)$) 
gives that 
\begin{equation}
\int_0^t\int \int f(z) (\rho_\epsilon * \eta_s)(z) 
\rho_\tau (z-y) (\rho_\epsilon * \eta_s)(y) dz dy ds
\to \int_0^t \int \int f(z) \varphi(s,z) 
\rho_\tau (z-y) \varphi(s,y) dz dy ds.
\end{equation}
Since $\tau$ is arbitrary, the 
convergence~(\ref{eq:QuadConvergence})
will follow.

\begin{proposition}
	\label{quadratic convergence}
Under the conditions of Theorem~\ref{thm:local_convergence}, we have that
along any convergent subsequence,
\begin{multline}
\label{approx4}
\limsup_{\epsilon \to 0}
\IE\Big[\Big|
\int_0^t\int f(y) \rho_\epsilon*\eta_s(y)^2 dy ds
\\
-
\int_0^t\int \int f(z) (\rho_\epsilon * \eta_s)(z) 
\rho_\tau (z-y) (\rho_\epsilon * \eta_s)(y) dz dy ds\Big|\Big]\leq C\tau,
\end{multline}
where $C$ is independent of $\tau$.
\end{proposition}
\begin{proof}
First note,
\begin{align}
&\int_0^t 
\mathbb{E}\left[\big| \langle f(y), (\eta_s * \rho_\epsilon(y))^2) dy \rangle 
- \int \int f(y) (\rho_\epsilon * \eta_s)(z) \rho_\tau (z-y) 
(\rho_\epsilon * \eta_s)(y) dz dy \big|\right] ds 
\nonumber 
\\ 
&\leq  
\Vert f \Vert_\infty \int  \int_0^t  
\mathbb{E}\left[ \int \left\{ \left|(\rho_\epsilon * \eta_s)(y)
-(\rho_\epsilon * \eta_s)(z) \right|   \rho_\tau(z-y) dz \right\} 
(\rho_\epsilon *\eta_s)(y) \right] ds  dy 
\label{b1}
\end{align}
Now proceed exactly as in the proof of 
Proposition~\ref{CuadraticApprox}.
The only distinction is that $|p_{\epsilon^2}(y,z)-p_{\epsilon^2}(w,z)|$
is replaced by $|p_{\tau}(y,z)-p_{\tau}(w,z)|$ and 
the estimate $\|y-w\|p_{\tau^2}(y,w)\leq C\tau p_{2\tau^2}(y,w)$
replaces the corresponding statement with $\epsilon^2$ replacing 
$\tau^2$ in our previous argument.
\end{proof}

The extension of Proposition~\ref{quadratic convergence}
to higher moments is straightforward, if notationally messy. For 
fixed (but arbitrary) $\tau$, one shows that
\begin{multline*}
	\limsup_{\epsilon\to 0}
	\IE\Big[\Big|
	\int_0^t\int f(y)\rho_\epsilon*\eta_s^N(y)^k dy ds
\\
	- \int_0^t\int\cdots\int f(y_1)
	\rho_\epsilon*\eta_s^N(y_1) 
\prod_{i=2}^{k}\rho_\tau (y_i-y_{i-1})\rho_\epsilon*\eta_s^N(y_i)
dy_{k}\ldots dy_1 ds\Big| \Big]
	\\
	\leq C\tau,
\end{multline*}
with a corresponding statement when $\rho_\epsilon*\eta_s^N(x)$ is replaced
by $\varphi(s,x)$ and then use weak convergence to see that, up to an error 
of order $\tau$, any limit point of
the sequence $\{\rho_\epsilon*\eta^N(x)dx\}$ solves (the weak form 
of) equation~(\ref{target equation in local convergence}).
Since $\tau$ was arbitrary, the proof of Theorem~\ref{thm:local_convergence}
is complete.

% % % % % % % % % % % % %
\section{Lookdown Proofs}
    \label{sec:lookdown_proofs}

Now we turn to results about the lookdown process,
first establishing the basic connection between the processes, Proposition~\ref{thm:mmt_application},
and then in the next section, convergence of the lookdown process itself.

\begin{proof}[of Proposition~\ref{thm:mmt_application}]
    This proposition is the content of the Markov Mapping Theorem,
    reproduced from \citet{etheridge/kurtz:2019} as Theorem~\ref{thm:mmt},
    applied to our situation.
    In the notation of that theorem, $\gamma$ is what we have called $\kappa$ above,
    $\alpha$ is the transition function that assigns levels uniformly on $[0, N]$ (in the first case)
    or as a Poisson process with Lesbegue intensity (in the limiting case),
    and we need a continuous $\psi^N(\lp)$ such that $|A^N f(\lp)| \le c_f \psi^N(\lp)$
    for all $f$ in the domain of $A^N$ (and similarly a function $\psi$ for $A$).
    We also need that applying the lookdown generator to a function and averaging over levels
    is equivalent to applying the population process generator to the function
    whose dependence on levels has been averaged out,
    a condition which we precisely state, and verify,
    in Lemmas~\ref{lem:averaged_generators} and~\ref{lem:averaged_generators_limit}
    of the Appendix.

    For finite $N$,
    taking $f(\lp)$ of the form \eqref{eqn:test_functions},
    we can use $\psi^N(\lp) = \langle C (1 + u |F(x, \eta)|), \lp(dx, du) \rangle$
    for an appropriate constant $C$.
    For the scaling limit, recall that the test functions $f$ are of the form
    $f(\lp) = \prod_{(x, u) \in \lp} g(x, u)$
    with $g(x, u) = 1$ for $u \ge u_0$,
    and so taking for example the most troublesome term in $Af(\lp)$,
    we see that since $0 \le f(\lp) / g(x, u) \le 1$ for any $(x, u) \in \lp$,
    \begin{align*} % \label{eqn:part_of_psi}
        \big| f(\lp) \sum_{(x, u) \in \lp} F(x, \eta) u \frac{\partial_u g(x, u)}{g(x, u)} \big|
        & \le
        \| \partial_u g \|_\infty \sum_{(x, u) \in \lp} | F(x, \eta) u \ind_{u \le u_0} | \\
        & \le
        \| \partial_u g \|_\infty e^{u_0} \sum_{(x, u) \in \lp} | F(x, \eta) u e^{-u} | .
    \end{align*}
    The first line would be just what we want,
    except that $\psi(\lp)$ cannot depend on $f$, and hence neither on $u_0$.
    So, the second line provides us with the required bound:
    we absorb $\| \partial_u g \|_\infty e^{u_0}$ into $c_f$
    and take $\psi(\lp) = \langle F(x, \eta) u e^{-u}, \lp(dx, du) \rangle$.
    % Furthermore, since $F(x, \eta)$ only depends on spatial location, not level,
    % for a given $\lp$ (and hence $\eta = \kappa(\lp)$),
    % $\sup_x |F(x, \eta)|$ is finite.
    % Therefore, we may bound \eqref{eqn:part_of_psi} by
    % $\sup_x |F(x, \eta)| \langle u e^{-u}, \lp \rangle$.
    % To do this we must restrict $\lp$ to lie in the space of measures
    % for which this is finite;
    % however, this is a completely metrizable space
    % \citep[see Condition 2.1 in][]{etheridge/kurtz:2019},
    % and our processes fall in this space (by conditional uniformity).
\end{proof}

% % % % % % % % % % % % %
\subsection{Tightness of the Lookdown Process}
    \label{sec:lookdown_convergence}

Now we turn to the main theorem on convergence of the lookdown process, 
Theorem~\ref{thm:lookdown_convergence}, whose proof
follows a similar pattern to those of convergence for the population processes
in Section~\ref{sec:population_density_proof}.

\begin{remark}
The process we consider is similar to the state-dependent branching processes
of \citet{KR}, so one might expect that the proofs there would carry over with little change.
However, there is an important difference:
Recall that the level $U_a(t)$ of a line of descent evolves as
$$
\dot u = c_\theta(x, \eta) u^2 - b_\theta(x, \eta) u,
$$
where $b_\theta(x, \eta)$ and $c_\theta(x, \eta)$ are defined in \eqref{b_defn}
and \eqref{c_defn} respectively.
Note that $c_\theta(x, \eta) \ge 0$ and $b_\theta(x, \eta)$ may take either sign.
Assumptions~\ref{def:model_setup} imply that $c^N(x, \eta)$ is bounded,
while $b^N(x, \eta)$, because of $F(x, \eta)$, is bounded above but not necessarily below.
In \citet{kurtz/rodrigues:2011}, $b^N$ was bounded above and $c^N$ was bounded away from zero,
so they noted that if $U_a(t) \ge b^N/c^N$ for some label $a$,
that line of descent would only move upwards from that time onwards.
Furthermore, coefficients did not depend on the state of the process (i.e., on $\eta$),
thus allowing the processes to be jointly and simultaneously constructed
for all values of $N$, with a pointwise embedding
of $(\lp^N_t)_{t \ge 0}$ within $(\lp^{M})_{t \ge 0}$ for $b^N/c^N < N < M$.
In other words, individuals with levels above $N > b^N/c^N$ at time $t_0$
do not affect $(\lp^N_t)_{t \ge t_0}$,
thus allowing a comparison of the number of lines of descent below level $u_0$
to a branching process.
Although we have provided a
joint construction of $\lp^N$ for all $N$ in Section~\ref{sec: individual lines of descent},
it does not have this monotonicity:
for one thing, $b^N$ and $c^N$ depend on the population process $\eta$
and so all individuals can affect all other ones (even those with lower levels).
Furthermore, in the deterministic case $\theta/N$, and hence $c$, converges to zero,
and so lines of descent with arbitrarily high level may drift back downwards.
Indeed, this must be the case if the population persists,
since in the deterministic case there is no branching.
\end{remark}

To complete the description of the process,
each line of descent
gives birth to lines at higher levels at rate $2 (N - u) c_\theta(x, \eta)$,
and each such new line chooses a level uniformly from $[u, N]$,
a spatial location $y$ from the kernel
\begin{equation} \label{eqn:qm_defn}
    q^m(x, dy, \eta)
    =
    r(y, \eta) q(x, dy) / \int_{\IR^d} r(y, \eta) q(x, dy),
\end{equation}
and the two lines swap spatial locations with probability 1/2.

It is evident from the description of the process
(or, by differentiating in Definition~\ref{defn:lookdown_mgale})
that
\begin{align} \label{eqn:f_xi_mgale}
\begin{split}
    \langle f, \lp_t^N \rangle
    &=
    \langle f, \lp_0^N \rangle
    + M^f_t
\\ & \qquad {}
    +
    \int_0^t
    \big\langle
        c_\theta(x, \eta^N_s)
        \int_u^N \int_{\IR^d}
        \left(
            f(y, u_1) + f(x, u_1) + f(y, u) - f(x, u)
        \right)
        q^m(x, dy, \eta^N_s) du_1 
\\ & \qquad \qquad \qquad \qquad {}
        +
        \left(
            c_\theta(x, \eta^N_s) u^2
            - b_\theta(x, \eta^N_s) u
        \right)
        \frac{d}{du} f(x, u)
    ,
    \lp_s^N(dx, du) \bigg \rangle
    ds ,
\end{split}
\end{align}
where $M^f$ is a martingale
with quadratic variation
\begin{align} \label{eqn:f_xi_qv}
    \begin{split}
    [M^f]_t
    &=
    \int_0^t
    \bigg \langle c_\theta(x, \eta_s^N) \int_u^N \int_{\IR^d}
        \big[ f(y, u_1)^2
\\&\qquad \qquad {}
        + \left( f(x, u_1) + f(y, u) - f(x, u) \right)^2 \big]
    du_1 q^m(x, dy, \eta^N_s)
    ,
    \lp_s^N(dx, du) \bigg \rangle
    ds .
\end{split}\end{align}

In addition to tightness of the measure-values processes $\lp^N$,
the bounds used in the proofs below also imply
tightness of the number of lines of descent and the number of births below a fixed level,
and of the motion of individual lines of descent.
In other words, the limiting ``line of descent'' construction holds.

\begin{proof}[Proof of Theorem~\ref{thm:lookdown_convergence}]
    As in Section~\ref{sec:population_density_proof},
    the theorem will follow from tightness and characterization of the limit.
    This time, the processes $\lp^N$ take values in $\lpmeasures$,
    the space of locally finite measures on space $\times$ levels.
    (They will in fact be point measures, including the limit,
    but that is a consequence of this theorem.)
    Again, tightness follows from a compact containment condition,
    tightness of one-dimensional distributions,
    and an application of \citet{EK} Theorem~3.9.1.

    Lines of descent can escape to infinite level in finite time, and so
    we endow $\lpmeasures$ with the vague topology ``in the level coordinate'',
    induced by test functions on $\overline{\IR^d} \times [0,\infty)$
    of the form $g(x) h(u)$, where $g \in C_b(\overline{\IR^d})$ is bounded and continuous
    and $h \in C_c([0, \infty))$ is compactly supported.
    This space is completely metrizable,
    and we may choose a countable set of nonnegative, compactly supported $f_k$
    such that a subset $K \subset \lpmeasures$ is relatively compact if and only if
    $\sup_{\lp \in K} \langle f_k, \lp \rangle < \infty$ for each $k$;
    below, Lemma~\ref{lem:compact containment lookdown} proves exatly this, and therefore compact containment.

    We require that $\{(F(\lp^N_t))_{t \ge 0}\}_N$
    is tight as a sequence of real-valued c\`adl\`ag processes, for all $F$ in a subset
    of $\bar C(\lpmeasures)$, the bounded, continuous functions on $\lpmeasures$,
    that is dense with respect to uniform convergence on compact subsets.
    Lemma~\ref{lem:test_fn_tightness} shows that $\{\langle f, \lp^N_t \rangle\}_N$ is a tight sequence
    for any $f: \overline{\IR^d} \times [0,\infty) \to \IR$ with compact support
    in the level direction.
    The functions $\lp \mapsto \langle f, \lp \rangle$
    do not form not a dense subset of $\bar C(\lpmeasures)$,
    but they do separate points and vanish nowhere, so
    by the Stone-Weierstrass theorem, the algebra they generate is dense
    % in $C_0(\lpmeasures)$, the continuous functions that vanish at infinity
    % (i.e., are less than $\epsilon$ outside of a compact set, for any $\epsilon$).
    % Since $C_0(\lpmeasures)$ is dense
    in $\bar C(\lpmeasures)$ (with respect to uniform convergence on compact subsets).
    So, it suffices to show that tightness for the processes $\langle f, \lp^N_t\rangle$
    extends to finite sums and products of these processes,
    which is shown in Lemma~\ref{lem:product_tightness}.

    The analogue of Lemma~\ref{lem:limit_mgale},
    in particular that the martingale properties are preserved under passage to the limit,
    are proved in Section~\ref{sec:lookdown_generator_proofs}.
    \comment{TODO: refer to correct thing}
    Finally, we must show that the limiting lookdown process $\lp$
    projects to the limiting process $\eta$, i.e., a solution of the martingale
    problem in Theorem~\ref{thm:nonlocal_convergence}.
    Let $N_k \to \infty$ be a sequence along which $\lp^{N_k}$ converges.
    By Theorem~\ref{thm:nonlocal_convergence},
    there is a subsequence $N_{k(j)}$ along which the projected population processes
    $\eta^{N_{k(j)}}$ converge, and the limit solves the martingale problem.
    Thus any limit point of $\lp^N$ projects to a population process $\eta$
    solving the martingale problem of Theorem~\ref{thm:nonlocal_convergence}.
\end{proof}

What we need for compact containment will come from the following Lemma.
The generality is unimportant --
for concreteness one may take $h(u) = e^{-u}$.

\begin{lemma}
        \label{lem:h_containment}
    Let $h$ be a positive, continuous, nonincreasing, differentiable function on $[0, \infty)$
    such that
    $\int_0^\infty \int_u^\infty h(v) dv du$,
    $\int_0^\infty u^2 |h'(u)| du$,
    and $\int_0^\infty h(u)^2 du$
    are all finite,
    Suppose that Assumptions~\ref{def:model_setup} hold,
    and that $\theta/N \to \alpha$ and $\lp_0^N \to \lp_0$ weakly as $N \to \infty$,
    where each $\lp_0^N$ is conditionally uniform given $\eta_0^N$
    in the sense of \eqref{eqn:conditionally_uniform}
    and $\lp_0$ is conditionally uniform given $\eta_0$
    in the sense of \eqref{eqn:conditionally_poisson}.
    Then for any $T$ there exists a constant $K(T)$ such that for all $M > 0$,
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle h, \lp^N_t \rangle > M \right\}
    <
    \frac{K(T)}{M} .
    $$
\end{lemma}

We postpone the proof of this Lemma to show how it gets us to compact containment.
First, we show that this implies compact containment for arbitrary
compactly supported functions.

\begin{lemma}
    \label{lem:compact containment hn}
    Suppose $f \in C(\bar \IR^d \times [0, \infty))$ and there is a $u_f$ such that
    $f(x, u) = 0$ for all $x$ and $u \ge u_f$.
    Under the assumptions of Lemma~\ref{lem:h_containment},
    for any $T$ there exists a constant $K(f,T)$ such that for all $M > 0$,
    $$ 
    \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle f, \lp^N_t \rangle > M \right\}
    <
    \frac{K(f,T)}{M} .
    $$
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:compact containment hn}]
    Let $h$ be as in Lemma~\ref{lem:h_containment},
    so there is a $c_f < \infty$ such that $f(x, u) \le c_f h(u)$ for all $x$ and $u$.
    Therefore, $\langle f, \xi \rangle \le \langle h, \xi \rangle$,
    and so by Lemma~\ref{lem:h_containment},
    \begin{align*}
        \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle f, \lp^N_t \rangle > M \right\}
        \le
        \limsup_{N \to \infty} \IP\left\{ \sup_{0 \le t \le T} \langle h, \lp^N_t \rangle > M/c_f \right\}
        <
        \frac{K(T) c_f}{M} .
    \end{align*}
\end{proof}

\begin{lemma}[Compact containment for $\lp$] \label{lem:compact containment lookdown}
    Let $f_1, f_2, \ldots$ be a sequence of functions each satisfying the conditions
    of Lemma~\ref{lem:compact containment hn}.
    Under the assumptions of Lemma~\ref{lem:h_containment},
    for any $T$ and $\delta > 0$ there exists a sequence
    $(C_1, C_2, \ldots)$ of finite constants such that
    \begin{align} \label{eqn:compact eqn lookdown}
    \limsup_{N \to \infty}
    \IP\left\{
        \sup_{0 \le t \le T} \langle f_k, \lp^N_t \rangle > C_k
        \text{ for some } k \ge 1
    \right\}
    < \delta .
    \end{align}
\end{lemma}

In other words, the processes $\lp^N$ stay in 
the set
$$
    \left\{
        \lp \in \lpmeasures \;:\;
        \langle h_k, \lp \rangle \le C_k
        \text{ for all } k \ge 1
    \right\} ,
$$
for all $0 \le t \le T$
with uniformly high probability,
a set which is compact if we choose $\{f_k\}$ appropriately.
% \comment{
%     Note that we could instead of the vague topology on locally finite measures,
%     take the space of measures with $\langle h, \xi \rangle < \infty$
%     for which $\xi_n \to \xi$ if $\langle f, \xi \rangle < \infty$ for all $f$ for which $f \le c_f h$,
%     which is completely metrizable
%     as in Condition 2.1 of \citet{etheridge/kurtz:2019}.
% }

\begin{proof}[of Lemma~\ref{lem:compact containment lookdown}]
    By a union bound,
    $$ 
    \IP\left\{
        \sup_{0 \le t \le T} \langle f_k, \lp^N_t \rangle > C_k
        \text{ for some } k \ge 1
    \right\}
    \le
    \sum_{k \ge 1}
    \IP\left\{
        \sup_{0 \le t \le T} \langle f_k, \lp^N_t \rangle > C_k
    \right\} ,
    $$
    so \eqref{eqn:compact eqn lookdown} follows
    by taking $C_k = 2^{k-1} K(f_k, T) / \delta$ and using Lemma~\ref{lem:compact containment hn}.
\end{proof}

Finally, we prove the key lemma.

\begin{proof}[of Lemma~\ref{lem:h_containment}]
    Applied to $f(x, u) = h(u)$, the martinagle representation~\eqref{eqn:f_xi_mgale} is
    \begin{align*}
    \begin{split}
        \langle h, \lp_t^N \rangle
        &=
        \langle h, \lp_0^N \rangle
        + M^{h}_t 
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 c^N(x, \eta^N_s) \int_u^N h(v) dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds
        \\ & \qquad \qquad {}
        +
        \int_0^t
        \big\langle
            \left(
                c^N(x, \eta^N_s) u^2
                - b^N(x, \eta^N_s) u
            \right)
            h'(u)
        ,
        \lp_s^N(dx, du) \big \rangle
        ds ,
    \end{split}
    \end{align*}
    where $M^{h}_t$ is a martingale
    with quadratic variation
    \begin{align*}
        [M^{h}]_t
        &=
        \int_0^t
        \langle 2 c^N(x, \eta_s^N) \int_u^N h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}
    Now, note that $0 \le c^N(x, \eta_x^N) \le C_a < \infty$
    and $b^N(x, \eta_s^N) \le C_b < \infty$,
    and we have assumed that $h'(u) \le 0$ (since $h$ is nonincreasing),
    so we may bound
    \begin{align} \label{eqn:hn_bound},
    \begin{split} 
        \langle h, \lp_t^N \rangle
        &\le
        \langle h, \lp_0^N \rangle
        + M^{h}_t
        \\ & \qquad {}
        +
        \int_0^t
        \big\langle
            2 C_a \int_u^\infty h(v) dv
            + \left( C_a u^2 + C_b u \right) |h'(u)|
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{split}
    \end{align}
    Now, since $\lp_t^N$ is conditionally uniform given $\eta_t^N$,
    we know that for appropriate $f$,
    $  \IE[\langle f, \lp^N_t \rangle] = \IE[\langle \widetilde{f}, \eta^N_t \rangle] , $
    where $\widetilde{f}(x) = \int_0^\infty f(x, u) du$.
    By our assumptions on $h$,
    we know that
    $$
        \int_0^\infty \left( 2 C_a \int_u^\infty h(v) dv
        + \left( C_a u^2 + C_b u \right) |h'(u)| \right) du
        < C
    $$
    for some $C < \infty$, and so
    \begin{align*}
    \begin{split}
        \IE\left[ \langle h, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h, \lp_0^N \rangle \right]
        +
        C
        \int_0^t
        \IE\bigg[
            \big\langle
            1
            ,
            \eta_s^N \big \rangle
        \bigg]
        ds ,
    \end{split}
    \end{align*}
    which we know by Lemma~\ref{lem:eta_f_bound}
    is bounded by $C_0 e^{C_1 t}$ for some other constants $C_0$ and $C_1$.

    Now consider the maximum. By \eqref{eqn:hn_bound},
    using that the integrand is nonnegative,
    \begin{align*}
        \sup_{0 \le t \le T} \langle h, \lp_t^N \rangle
        &\le
        \langle h, \lp_0^N \rangle
        + \sup_{0 \le t \le T} M^{h}_t 
        \\ & \qquad {}
        + \int_0^T 
        \big \langle
            2 C_a \int_u^\infty h(v) dv
            + \left( C_a u^2 + C_b u \right) |h'(u)|
        ,
        \lp_s^N(dx, du) \big \rangle
        ds .
    \end{align*}

    Again since $\sqrt{x} \le 1 + x$ for $x \ge 0$,
    the Burkholder-Davis-Gundy inequality
    tells us that there is a $C'$ such that
    \begin{align*}
    \IE\left[ \sup_{0 \le t \le T} M^{h}_t \right]
    & \le
        C'\left( 1 + \IE\left[ [M^{h}]_T \right] \right)
    \\ & \le
        C' \left( 1 + \int_0^T \IE\left[
        \langle 2 c^N(x, \eta_s^N) \int_u^\infty h(v)^2 dv
        ,
        \lp_s^N(dx, du) \big \rangle
        \right] ds \right) 
    \\ & \le
        C' \left( 1 + 2 C_a \int_0^\infty h(v)^2 dv
        \int_0^T \IE\left[
            \langle 1 , \lp_s^N(dx, du) \big \rangle
        \right] ds \right)  
    \\ & \le
        C_2 e^{C_1 T} ,
    \end{align*}
    for a constant $C_2$ which is finite
    by our assumption that $\int_0^\infty h(v)^2 dv < \infty$.

    Therefore, 
    \begin{align*}
        \IE\left[ \sup_{0 \le t \le T} \langle h, \lp_t^N \rangle \right]
        &\le
        \IE\left[ \langle h, \lp_0^N \rangle \right]
        + (C_2 + C_0 / C_1) e^{C_1 T},
    \end{align*}
    and so
    \begin{align*}
        \IP\left\{ \sup_{0 \le t \le T} \langle h_n, \lp_t^N \rangle > K \right\}
        &\le
        \frac{
            \IE\left[ \langle h, \lp_0^N \rangle \right]
            + (C_2 + C_0 / C_1) e^{C_1 T}
        }{ K } .
    \end{align*}
\end{proof}

\begin{lemma}
    \label{lem:test_fn_tightness}
    Let $f$ be a bounded, continuous real-valued function on $\IR^d \times [0, \infty)$
    with uniformly bounded first and second derivatives
    for which there exists a $u_0$ such that if $u > u_0$ then $f(x, u) = 0$.
    Then, the sequence of real-valued processes $(\langle f, \lp_t^N \rangle)_{t \ge 0}$
    are tight in the space of c\`adl\`ag processes.
\end{lemma}

\begin{proof}[of Lemma \ref{lem:test_fn_tightness}]
    Again, we use the Aldous-Rebolledo criterion.
    Tightness of $\langle f, \lp_t \rangle$ for a fixed $t$
    follows from Lemma~\ref{lem:compact containment hn},
    so we need only prove conditions analogous to
    \eqref{eqn:eta_projections_goal1}
    and \eqref{eqn:eta_projections_goal2}
    applied to the martingale representation
    of equations~\eqref{eqn:f_xi_mgale} and~\eqref{eqn:f_xi_qv}.
    Rewriting~\eqref{eqn:f_xi_mgale} with $c_\theta = c_\theta(x, \eta_s)$,
    \begin{align*}
        \langle f, \lp_t \rangle
        &=
        \langle f, \lp_0 \rangle
        + M^f_t
        + \int_0^t \Big\langle
            c_\theta \int_u^N \int (f(y, u_1) + f(x, u_1)) q^m(x,dy,\eta) du_1
        \\ {} &\qquad
            + c_\theta (N - u) \int (f(y, u) - f(x, u)) q^m(x,dy,\eta)
            + (c_\theta u^2 - b_\theta u) \frac{d}{du} f,
            \lp_s
        \Big\rangle ds
        \\ &\le
        \langle f, \lp_0 \rangle
        + M^f_t
        + \int_0^t \Big\langle
            c_\theta \int_u^N \int (f(y, u_1) + f(x, u_1)) q^m(x,dy,\eta) du_1
        \\ {} &\qquad
            + c_\theta (N - u) \int (f(y, u) - f(x, u)) q^m(x,dy,\eta)
            + (c_\theta u^2 - b_\theta u) \frac{d}{du} f,
            \lp_s
        \Big\rangle ds
    \end{align*}
    The bounds analogous to \eqref{eqn:eta_projections_goal1} and \eqref{eqn:eta_projections_goal2}
    follow as in the proof of Lemma~\ref{lem:eta_projections_tightness}:
    for instance, observe that using that $c_\theta \le C_0$ for some $C_0$,
    the predictable part of this semimartingale decomposition is bounded by
    \begin{align*}
        \Big\langle
            2 C_0 \|f\|_\infty u_f
            + (1 - u/N) \gamma B^\theta_f
            + (C_0 u^2 - b_\theta u) \frac{d}{du} f,
            \lp_s
        \Big\rangle
        &\le
        \langle
            C_0 u_f^2 + \sup_x |b_\theta(x, \eta_s)| u_f \|\frac{d}{du} f\|_\infty
        \rangle ,
    \end{align*}
    which can be bounded as we did for \eqref{eqn:eta_projections_goal1}.
\end{proof}

% % % % % % % % % % % % %
\subsection{Motion of ancestral lineages}
\label{sec:lineages_proof}

In this section we prove Theorem~\ref{thm:lineages}.
The argument follows directly from the discussion in Section \ref{sec:limiting_lines_of_descent}.

\begin{proof}[of Theorem~\ref{thm:lineages}]

Here we have taken the high-density, deterministic limit
(so, $\theta, N \to \infty$ and $\theta/N \to 0$)
and so suppose that the limiting process has density $\varphi_t(x)$
at location $x$ and time $t$.
Let $Y$ denote the spatial motion followed by a single line of descent.
Above equation~\eqref{eqn:limiting_generator},
we showed that $Y$ is a diffusion with generator at time $s$
$$
    \mathcal{L}^Y_s g(x) = \gamma ( \DG(r(\cdot,\eta_s) g(\cdot))(x) - g(x) \DG r(x,\eta_s) ) .
$$
% (This is $\mathcal{L}^Y_s g(x) = \gamma ( \DG(rg)(x) - g \DG r(x) )$ in shorter but more cryptic notation.)
The diffusion is time-inhomogeneous if the density is not constant with time
(since recall that the coefficients $\gamma$ and $r$ depend on the density).
Let $\varphi_t(x)$ be the limiting density, which solves \eqref{general deterministic limit},
$\partial_t \varphi = r \DG^*[ \varphi_t \gamma ] + \varphi F$.
Formally, the intensity of levels at $y$ at time $t$
that are descended from individuals that were at $x$ at time $s$
(with $s < t$) is
\begin{equation} \label{eqn:formal_intensity}
    \varphi_s(x) \IE_{s,x} \left[
        \exp\left(
            \int_s^t (F + \gamma \DG r)(Y_u) du
        \right)
        \ind_{Y_t = y}
    \right]
    dy ,
\end{equation}
where the subscript $s, x$ in the expectation indicates that $Y_s = x$.
To see why this should be true, 
suppose that an ancestor at time $s$ has level $v$. Conditional on its 
spatial motion $\{Y_u\}_{s\leq u\leq t}$, its level at time $t$ will
be $v \exp(-\int_s^t(F+\gamma\Delta r)(Y_u)du)$. This will be less than a given level 
$\lambda$ if $v < \lambda \exp(\int_s^t(F+\gamma\Delta r)(Y_u)du)$. 
The intensity of levels at $y$ that are descended from individuals at
$x$ can therefore be obtained as the limit as $\lambda\to\infty$ of 
$1/\lambda$ times the number of levels at $x$ at time zero with
$u<\lambda \exp(\int_s^t(F+\gamma\Delta r)(Y_u)du)$ and for which
the corresponding individual is at $y$ at time $t$, which is 
precisely the quantity in~\eqref{eqn:formal_intensity}.

By our construction in Section~\ref{sec:limiting_lines_of_descent},
when we integrate~\eqref{eqn:formal_intensity}
with respect to $x$ we recover $\varphi_t(y)dy$. 
Consider an individual sampled at location $y$ at time $t$,
and write $p(t,s,y,x)$ for the probability density
that their ancestor at time $s$ was at $x$.
As a consequence of~\eqref{eqn:formal_intensity},
still formally,
\begin{equation}
\label{eqn:ptsyx}
    p(t,s,y,x)
    =
    \frac{\varphi_s(x)}{\varphi_t(y)}
    \IE_{s,x}\left[
        \exp\left( \int_s^t (F + \gamma\DG r)(Y_u) du \right)
        \ind_{Y_t=y}
    \right]
\qquad \text{ for } s < t.
\end{equation}

To make~(\ref{eqn:ptsyx}) meaningful, we multiply by suitable test functions
$f$ and $g$ and integrate.
\begin{align*}
&\int \int f(y) \varphi_t(y) p(t,s,y,x) g(x) dy dx \\
&\qquad =
\int g(x) \varphi_s(x)
\IE_{x,s}\left[
    \exp\left(
        \int_s^t(F+\gamma\DG r)(Y_u)du
    \right)f(Y_t)
\right] dx .
\end{align*}

Writing $\hat{T}_{t,s}$ for the time-inhomogeneous semigroup
corresponding to the motion of ancestral lineages backwards in time
(that is, $\hat{T}_{t,s} f(y) = \int p(t,s,x,y) f(x) dy$),
we can write this as 
$$
    \int f(y)\varphi_t(y)\hat{T}_{t,s}g(y)dy
    =
    \int g(x) \varphi_s(x)
        \IE_{s,x} \left[
            \exp\left(
                \int_s^t(F+\gamma\DG r)(Y_u)du
            \right)f(Y_t)
        \right]dx.
$$
Differentiating with respect to $t$ at $t=s$, and writing 
$\Lgen_s$ for the generator of $\hat{T}_{t,s}$ at time $s$, we find
\begin{align*}
&
    \int f(y) \left\{
        \varphi_s(y) \Lgen_s g(y)
        + \left(
            r(y) \DG^*(\gamma \varphi_s)(y) + \varphi_s(y) F(y)
        \right) g(y)
    \right\} dy \\
&\qquad =
    \int g(x) \varphi_s(x) \left(
        {\mathcal L}^Y_s f(x) + (F(x)+\gamma(x) \DG r(x)) f(x)
    \right) dx \\
&\qquad =
    \int f(x) \left(
        ({\mathcal L}^Y_s)^* (\varphi_s g)(x) + (F(x) + \gamma(x) \DG r(x)) \varphi_s(x) g(x)
    \right) dx ,
\end{align*}
where $({\mathcal L}^Y_s)^*$ is the adjoint of ${\mathcal L}^Y_s$
and we have used that $\varphi_s$ solves $\dot \varphi = r \DG^*(\gamma \varphi) + F \varphi$.
Since $f$ was arbitrary,
\begin{eqnarray*}
\Lgen_s g
    &=&
    \frac{1}{\varphi_s} \left[
        ({\mathcal L}^Y_s)^* (\varphi_s g)
        + \gamma \varphi_s g \DG(r)
        - r g \DG^*(\gamma \varphi_s)
    \right] .
\end{eqnarray*}
(Note that the $\phi_s F g$ terms have cancelled.)
Since the adjoint of ${\mathcal L}^Y_s$ is
\begin{align*}
    ({\mathcal L}^Y_s)^* f
    &=
    r \DG^* (\gamma f) - \gamma f \DG r ,
\end{align*}
we can rewrite the generator of a lineage as 
\begin{eqnarray*}
\Lgen_s g
    &=&
    \frac{r}{\varphi_s} \left[
        \DG^* (\gamma \varphi_s g)
        - g \DG^*(\gamma \varphi_s)
    \right] .
\end{eqnarray*}
This is equation \eqref{eqn:lineage_generator}.

To simplify to equation \eqref{eqn:lineage_generator2},
first define
$
    \DD f(x) = \sum_{ij} \covq_{ij} \partial_{ij} f(x),
$
% and note that $\DD$ satisfies the following identity:
% \begin{align}
%     \DD(fg)
%     &=
%     f \DD g
%     + 2 (\covq \grad f) \cdot \grad g
%     + g \DD f .
% \end{align}
and so the adjoint of $\DD$ is
$$
    \DD^* f(x)
    =
    \sum_{ij} \partial_{ij} (\covq_{ij} f(x)) .
$$
Note that $\DD^*$ satisfies the following identity:
\begin{align*}
    \DD^*(fg)
%     &=
%     \sum_{ij} \left\{
%         g \partial_{ij} (\covq_{ij} f)
%         + 2 \partial_{i} (\covq_{ij} f) \partial_j(g)
%         + \covq_{ij} f \partial_{ij} g
%     \right\} \\
    &=
    \sum_{ij} \left\{
        g \partial_{ij} (\covq_{ij} f)
        + 2 f \partial_{i} (\covq_{ij}) \partial_j(g)
        + 2 \covq_{ij} \partial_{i} (f) \partial_j(g)
        + \covq_{ij} f \partial_{ij} g
    \right\} \\
    &=
    g \DD^* f
    + 2 f \vec{c} \cdot \grad g
    + 2 (C \grad f) \cdot \grad g
    + f \DD g ,
\end{align*}
where $\vec{c}_j = \sum_i \partial_j C_{ij}$.
So, with $f = \gamma \varphi_s$,
\begin{eqnarray*}
\Lgen_s g
    &=&
    \frac{r}{\varphi_s} \left[
        \DD^*(\gamma \varphi_s g) - \grad \cdot (\gamma \varphi_s g \meanq)
        - g \DD^*(\gamma \varphi_s) + g \grad \cdot (\gamma \varphi_s \meanq)
    \right] \\
    &=&
    \frac{r}{\varphi_s} \left[
        \gamma \varphi_s \DD g
        + 2 \gamma \varphi_s \vec{c} \cdot \grad g
        + 2 (C \grad (\gamma \varphi_s)) \cdot \grad g
        - \gamma \varphi_s \meanq \cdot \grad g
    \right] \\
    &=&
    r \gamma \left[
        \DD g
        + 2 \vec{c} \cdot \grad g
        + 2 (C \grad \log(\gamma \varphi_s)) \cdot \grad g
        - \meanq \cdot \grad g
    \right] ,
\end{eqnarray*}
which is equation \eqref{eqn:lineage_generator2},
\end{proof}

\begin{proof}[of Corollary~\ref{cor:stationary_dist}]
For the moment, we will write $r(x)$ for $r(x, \eta)$
and similarly for $\gamma(x)$.
First, note that in this case
$$
    \Lgen f = \sigma^2 r \gamma \left(
        \Delta f
        + \grad ( 2 \log(\gamma \varphi) - h ) \cdot \grad f
    \right) ,
$$
and that
$$
    \int_{\IR^d} e^{H(x)} f(x) (\Delta + \grad H(x) \cdot \grad) g(x) dx
    =
    - \int_{\IR^d} e^{H(x)} \left\{
        \grad f(x) \cdot \grad g(x)
    \right\} dx .
$$
Therefore, by choosing $H(x) = 2 \log(\gamma(x) \varphi(x)) - h(x)$
and
$$ \pi(x)
    = e^{H(x)} / (\sigma^2 r(x) \gamma(x))
    = \frac{\gamma(x) \varphi(x)^2 e^{-h(x)}}{\sigma^2 r(x)} ,
$$
we have that
$$
    \int_{\IR^d} \pi(x) f(x) \Lgen g(x) dx
    =
    - \int_{\IR^d} e^{H(x)} \grad f(x) \cdot \grad g(x) dx .
$$
Since this is symmetric in $f$ and $g$, the process $Y$ is reversible with respect to $\pi$;
the constant factor of $\sigma^2$ does not affect the result.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix

% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %%
\section{Markov Mapping Theorem}
\label{apx:mmt}

The following appears as Theorem A.2 in \citet{etheridge/kurtz:2019},
specialized slightly to the case that the processes 
are c\'adl\'ag and have no fixed points of discontinuity.
For an $S_0$-valued, measurable process $Y$, $\hat{\mathcal{F}}^Y_t$
denotes the completion of the $\sigma$-algebra generated by
$Y(0)$ and $\{\int_0^r h(Y(s)) ds, r \le t, h \in B(S_0)\}$.
Also, let $D_S[0,\infty)$ denote the space of c\`adl\`ag, $S$-valued functions
with the Skorohod topology, and $M_S[0,\infty)$ the space of Borel measurable functions
from $[0,\infty)$ to $S$,
topologized by convergence in Lesbegue measure.
For other definitions see \citet{etheridge/kurtz:2019}.

\begin{theorem}[Markov Mapping Theorem] \label{thm:mmt}
    Let $(S,d)$ and $(S_0,d_0)$ be complete, separable metric spaces.
    Let $A \subset \bar C(S) \times C(S)$ and $\psi \in C(S)$, $\psi \ge 1$.
    Suppose that for each $f \in \mathcal{D}(A)$ there exists $c_f$
    such that
    $$ |Af(x)| \le c_f \psi(x), \qquad x \in A, $$
    and define $A_0 f(x) = Af(x) / \psi(x)$.

    Suppose that $A_0$ is a countably determined pre-generator,
    and suppose that $\mathcal{D}(A) = \mathcal{D}(A_0)$
    is closed under multiplication and is separating.
    Let $\gamma : S \to S_0$ be Borel measurable,
    and let $\alpha$ be a transition function from $S_0$ into $S$
    ($y \in S_0 \to \alpha(y,\cdot)\in\mathcal{P}(S)$ is Borel measurable)
    satisfying $\int h\circ\gamma(x)\alpha(y,dx) = h(y)$ for $y \in S_0$ and $h \in B(S_0)$,
    that is, $\alpha(y,\gamma^{-1}(y)) = 1$.
    Assume that $\tilde \psi(y) \equiv \int_S \psi(z) \alpha(y,dz) < \infty$ for each $y \in S_0$
    and define
    $$  C = \{
        \int_S f(z) \alpha(\cdot,dz),
        \int_S Af(z) \alpha(\cdot,dz)
        \; : \; f \in \mathcal{D}(A)
    \} . $$
    Let $\mu_0 \in \mathcal{P}(S_0)$ and define $\nu_0 = \int \alpha(y,\cdot) \mu_0(dy)$.
    \begin{itemize}
        \item[(a)]
            If $\tilde Y$ satisfies $\int_0^t \IE[\tilde \psi(\tilde Y(s))]ds < \infty$
            for all $t \ge 0$ and $\tilde Y$ is a solution of the martingale problem for $(C,\mu_0)$,
            then there exists a solution $X$ of the martingale problem for $(A,\nu_0)$
            such that $\tilde Y$ has the same distribution on $M_{S_0}[0,\infty)$
            as $Y = \gamma \circ X$.
            If $Y$ and $\tilde Y$ are c\'adl\'ag, then $Y$ and $\tilde Y$ have the same distribution
            on $D_{S_0}[0,\infty)$.
        \item[(b)]
            For $t \ge 0$,
            $$ \IP\{ X(t) \in \Gamma \;|\; \hat{\mathcal{F}}^Y_t \}
            = \alpha(Y(t),\Gamma), \qquad \text{for } \Gamma \in \mathcal{B}(S).
            $$
        \item[(c)]
            If, in addition, uniqueness holds for the martingale problem for $(A,\nu_0)$,
            then uniqueness holds for the $M_{S_0}[0,\infty)$-martingale problem for $(C,\mu_0)$.
            If $\tilde Y$ has sample paths in $D_{S_0}[0,\infty)$ then uniqueness holds for the
            $D_{S_0}[0,\infty)$-martingale problem for $(C,\mu_0)$.
        \item[(d)]
            If uniqueness holds for the martingale problem for $(A,\nu_0)$
            then $Y$ is a Markov process.
    \end{itemize}
\end{theorem}

In our application, we have taken $S$ to be the space of locally finite counting measures
on $\IR^d \times [0,N)$ or on $\IR^d \times [0,\infty)$,
and $S_0$ the space of finite measures on $\bar \IR^d$.
Then, $A$ corresponds to the generator for the lookdown process (i.e., either $A^N$ or $A$),
and $C$ corresponds to the generator for the spatial population process
(i.e., either $\Pgen^N$ or $\Pgen$).
The ``$\gamma$'' of the theorem is our spatial projection operator
that we have called $\kappa^N$ or $\kappa$,
and the ``$\alpha$'' of the theorem will be named $\Gamma^N$ or $\Gamma$ below.
Finally, ``$X$'' of the theorem is our lookdown process, $\lp$,
and ``$Y$'' is our spatial process, $\eta$.

% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %% %%
\subsection{Lookdown Generators}

In this section we verify one of the conditions of the Markov Mapping Theorem,
namely, that ``integrating out levels'' in the generator of the lookdown process
we obtain the generator of the projected process.
We will work with test functions of the form
$$  f(\lp) = \prod_{(x, u) \in \lp} g(x, u) = \exp\left(\langle \log g, \lp \rangle \right) . $$
Furthermore, recall that $\kappa^N(\lp)(\cdot) = \lp(\cdot \times [0, N)) / N$
is the ``spatial projection operator'',
and define the probability kernel $\Gamma^N : \measures \to \mathcal{M}(\IR^d \times [0,N))$
so that for $\eta \in \measures$, if $\hat g(x) = \int_0^N g(x, u) du / N$, then
\begin{align*}
    F^N_g(\eta)
    &:=
    \int f(\lp) \Gamma^N(\eta, d\lp) \\
    &=
    \exp\left(
        N \left\langle
            \log \frac{1}{N} \int_0^N g(x, u) du, \eta(dx)
        \right\rangle
    \right) \\
    &=
    \exp\left(
        N \left\langle \log \hat g(x), \eta(dx) \right\rangle
    \right) ,
\end{align*}
i.e., $\Gamma^N$ assigns independent labels on $[0, N]$ to each of the points in $\eta$.
It follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
\comment{(beware the different definition of $f$ to that lemma)}
the generator of $\eta^N_t$ is
\begin{align} \label{eqn:pgen_defn_again}
    \begin{split}
    \Pgen^N F^N_g(\eta)
    =
    F^N_g(\eta)
        N \theta & \bigg\langle
        \gamma(x, \eta) \int r(z, \eta) \left( \hat g(z) - 1 \right) q_\theta(x, dz)
    \\ &\qquad \qquad {}
        +
        \mu_\theta(x, \eta) \left( \frac{1}{\hat g(x)} - 1 \right)
        ,
        \eta(dx)
    \bigg\rangle
    \end{split}
\end{align}
The generator of $\lp^N_t$ is $A^N$, defined in equation~\eqref{eqn:lookdown_generator}.

\begin{lemma}
    \label{lem:averaged_generators}
    For all finite counting measures $\eta$ on $\IR^d$,
    \begin{align}
        \int A^N f(\lp) \Gamma^N(\eta, d\lp)
        &=
        \Pgen^N F^N_g(\eta) .
    \end{align}
\end{lemma}

\begin{lemma}
    \label{lem:averaged_generators_limit}
    For all $\eta \in \measures$,
    \begin{align}
        \int A f(\lp) \Gamma(\eta, d\lp)
        &=
        \Pgen^\infty F_g(\eta) .
    \end{align}
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:averaged_generators}]
First, break the generator $A^N$ into three parts, 
\begin{align*}
    A^N_1f(\lp)
    &=
    f(\lp) \sum_{(x,u)\in\lp} 2 c_\theta(x,\eta)
    \int_u^{N}\Bigg(
        \frac{1}{2}\frac{g(x,v_1)}{g(x,u)}
        \int_{\IR^d} (g(y,u)-g(x,u))q^m_{\theta}(x,dy,\eta)
    \Bigg)dv_1 ,
    \\
    A^N_2f(\lp)
    &=
    f(\lp) \sum_{(x,u)\in\lp} 2 c_\theta(x,\eta)
    \int_u^{N}\Bigg(
        \frac{1}{2}\int_{\IR^d}
            \left(\frac{g(y,v_1) + g(x,v_1)}{2} - 1\right)
        q^m_{\theta}(x,dy,\eta)
    \Bigg)dv_1 ,
    \\
    A^N_3f(\lp)
    &=
    f(\lp) \sum_{(x,u)\in\lp}\,
    \left( c_\theta(x,\eta) u^2 - b_{\theta}(x,\eta)u \right) \frac{\partial_u g(x,u)}{g(x,u)},
\end{align*}
where $q^m$ was defined in equation~\eqref{eqn:qm_defn},
so that 
$$ A^Nf(\lp) = A^N_1f(\lp) + A^N_2f(\lp) + A^N_3f(\lp). $$

We now integrate each piece against $\Gamma^N$.
First note that by the product form of $f$,
$$
    \int f(\lp) \sum_{(x,u)\in\lp} \frac{\ell(x,u)}{g(x,u)} \Gamma^N(\eta,d\lp)
    =
    F_g^N(\eta) \bigg\langle
        \frac{1}{N \hat g(x)} \int_0^N \ell(x,u) du,
        N \eta(dx)
    \bigg\rangle .
$$
Therefore,
\footnotesize
\begin{align*}
&
    \int A^N_1f(\lp) \Gamma^N(\eta,d\lp) 
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \int_0^N \Bigg\{
        \int_u^N \Bigg(
            \frac{1}{2} g(x,v_1) \int_{\IR^d}(g(y,u)-g(x,u))q^m_{\theta}(x,dy,\eta)
        \Bigg) dv_1
    \Bigg\} du, \eta(dx) \bigg \rangle
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \Bigg\{
    \int_{\IR^d} \left\{
        \int_{0}^{N} \int_u^N g(x,v_1) (g(y,u)-g(x,u)) dv_1 du
    \right\} q^m_{\theta}(x,dy,\eta)
    , \eta(dx) \bigg \rangle .
\end{align*}
\normalsize
For the second generator, we have
\footnotesize
\begin{align*}
&
    \int A^N_2 f(\lp) \Gamma^N(\eta, d\lp)
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)} \int_0^N g(x, u) \Bigg\{
        \int_u^N\Bigg(\int_{\IR^d}
        \left(\frac{g(y,v_1)+g(x,v_1)}{2}-1\right) q^m_{\theta}(x,dy,\eta)
        \Bigg)dv_1 \Bigg\}
    , \eta(dx) \bigg\rangle
\\ &\qquad =
    F_g^N(\eta) \bigg\langle \frac{c_\theta(x,\eta)}{\hat g(x)}
    \int_{\IR^d}\Bigg\{ \int_{0}^{N}
    \int_u^N g(x,u) \left(
        g(y,v_1)+g(x,v_1)-2
    \right) dv_1 du\Bigg\} q^m_{\theta}(x,dy,\eta)
    , \eta(dx) \bigg\rangle .
\end{align*}  
\normalsize
For the third generator we have that, since $g(x,N) = 1$,
\footnotesize
\begin{align*}
&
    \int A^N_3f(\xi)\Gamma^N(\eta,d\lp)
\\ & \qquad =
    F_g^N(\eta) \bigg\langle \frac{1}{\hat g(x)}
    \int_0^N
        \left( c_\theta(x,\eta) u^2 - b_\theta(x,\eta) u \right) \partial_u g(x,u)
    du
    , \eta(dx) \bigg\rangle 
\\ & \qquad =
    F_g^N(\eta) \bigg\langle \frac{1}{\hat g(x)}
    \int_0^N
        \left( b_\theta(x,\eta) - 2 c_\theta(x,\eta) u \right) (g(x,u) - 1)
    du
    , \eta(dx) \bigg\rangle .
\end{align*}
\normalsize

Note that $2 \int_0^N \int_u^N g(x,v_1) g(y,u) dv_1 du = N^2 \hat g(x) \hat g(y)$,
and so
\footnotesize
\begin{align*}
&
    \int_{0}^{N} \int_u^N g(x,v_1) (g(y,u)-g(x,u)) dv_1 du
    + \int_{0}^{N} \int_u^N g(x,u) \left(g(y,v_1)+g(x,v_1)-2\right)dv_1 du
\\ &\qquad = 
    N^2 \hat{g}(x)(\hat{g}(y) - 2) + 2\int_{0}^{N} u g(x,u) du.
\end{align*}
\normalsize

Combining the last equations,
and using the fact that
$N c_\theta(x,\eta) - b_\theta(x,\eta) = \theta \mu_\theta(x,\eta)$,
we have 
\begin{align*}
&
    \int \Bigg( A^N_1 f(\xi) + A^N_2 f(\xi) + A^N_3 f(\xi) \Bigg) \Gamma^N(\eta,d\lp)
\\ &\qquad =
    F_g^N(\eta) \bigg\langle
        c_\theta(x,\eta) N^2 \int_{\IR^d} (\hat g(y) - 2) q^m(x,dy,\eta)
        +
        \frac{1}{\hat g(x)} c_\theta(x,\eta) \int_0^N 2u du
\\ &\qquad \qquad \qquad {}
        +
        \frac{1}{\hat g(x)} b_\theta(x,\eta) \int_0^N (g(x,u) - 1) du
    , \eta(dx) \bigg\rangle 
\\ &\qquad =
    F_g^N(\eta) \bigg\langle
        c_\theta(x,\eta) N^2 \int_{\IR^d} (\hat g(y) - 1) q^m(x,dy,\eta)
        +
        N^2 c_\theta(x,\eta) \left(\frac{1}{\hat g(x)} - 1\right)
\\ &\qquad \qquad \qquad {}
        +
        N b_\theta(x,\eta) \left(1 - \frac{1}{\hat g(x)}\right)
    , \eta(dx) \bigg\rangle 
\\ &\qquad =
    F_g^N(\eta) N \bigg\langle
        N c_\theta(x,\eta) \int_{\IR^d} (\hat g(y) - 1) q^m(x,dy,\eta)
        +
        \theta \mu_\theta(x,\eta) \left(\frac{1}{\hat g(x)} - 1\right)
    , \eta(dx) \bigg\rangle  .
\end{align*}
This matches equation~\eqref{eqn:pgen_defn_again}, as desired,
because $N c_\theta(x,\eta) q_m(x,dy,\eta) = \theta \gamma(x,\eta) q_\theta(x,dy)$.
\end{proof}

For the limiting process,
we now require test functions such that $g(x,u) = 1$ for all $u$ larger than some $u_0$;
As before, recall that $\kappa(\lp)(\cdot) = \lim_{u \to \infty} \lp(\cdot \times [0, u)) / u$
is the ``spatial projection operator'',
and define the probability kernel $\Gamma : \measures \to \flpmeasures$
so that for $\eta \in \measures$,
defining $h(x) = \int_0^\infty (g(x, u) - 1) du$,
\begin{align*}
    F_g(\eta)
    &:=
    \int f(\lp) \Gamma(\eta, d\lp) \\
    &=
    \exp\left(
        \big \langle
            \int_0^\infty (g(x, u) - 1) du,
            \eta(dx)
        \big \rangle
    \right) \\
    &=
    e^{ \langle h(x), \eta(dx) \rangle } .
\end{align*}
i.e., $\Gamma(\eta, \cdot)$ is the distribution of a conditionally Poisson process
with intensity a product of $\eta$ and Lebesgue measure.
It again follows from Lemma~\ref{def: MP definition of limit} that for test functions of this form
\comment{(beware the different definition of $f$ to that lemma)}
the generator of $\eta_t$ is
\begin{align} \label{eqn:pgen_infty}
    \Pgen^\infty F_g(\eta)
    &=
    F_g(\eta)
    \left\langle
        \gamma(x, \eta)
        \DG( h(\cdot) r(\cdot) )(x)
        + F(x, \eta) h(x)
        + \alpha \gamma(x, \eta) r(x, \eta) h^2(x)
        ,
        \eta(dx)
    \right\rangle .
\end{align}
The generator of $\lp_t$ is $A$, defined in equation~\eqref{eqn:limiting_lookdown_generator}.

Before proving Lemma~\ref{lem:averaged_generators_limit},
we recall an important equality for conditionally
Poisson point processes (\cite{kurtz/rodrigues:2011} Lemma A.3).

\begin{lemma} \label{lem:poisson_eqn}
If $\lp = \sum_{i}\delta_{Z_i}$ is a Poisson random measure with mean measure $\nu$, 
then for $\ell \in L^{1}(\nu)$ and $g\geq0$ with $\log g \in L^{1}(\nu)$,
\begin{equation}
\IE\left[\sum_{j} \ell(Z_j) \prod_{i}g(Z_i)\right] = \int \ell g d\nu e^{\int (g-1) d \nu}.
\end{equation}
\end{lemma}

\begin{proof}[of Lemma~\ref{lem:averaged_generators_limit}]
By Lemma~\ref{lem:poisson_eqn},
$$
    \int f(\lp) \sum_{(x,u)\in\lp} \frac{\ell(x,u)}{g(x,u)} \Gamma(\eta, \lp)
    =
    F_g(\eta) \bigg\langle \int_0^\infty \ell(x,u) du , \eta(dx) \bigg\rangle .
$$
Comparing this to the definition of $A$
(equation~\eqref{eqn:limiting_lookdown_generator}),
we see that
\begin{align*}
    \int Af(\lp) \Gamma(\eta,d\lp)
    &=
    F_g(\eta) \bigg\langle
        \int_0^\infty
        (\ell_1(x,u) + \ell_2(x,u) + \ell_3(x,u))
        du, \eta(dx)
    \bigg\rangle ,
\end{align*}
where
\begin{align*}
    \ell_1(x,u)
    &=
    \gamma(x,\eta)\left(
        \DG(g(\cdot,u) r(\cdot,\eta))(x) - g(x,u) \DG r(x,\eta)
    \right) \\
    &=
    \gamma(x,\eta)\left(
        \DG((g(\cdot,u)-1) r(\cdot,\eta))(x) - (g(x,u)-1) \DG r(x,\eta)
    \right)
\end{align*}
and
\begin{align*}
    \ell_2(x,u)
    =
    2 g(x,u) \alpha \gamma(x,\eta) r(x,\eta) \int_u^\infty(g(x,v) - 1)dv
\end{align*}
and
\begin{align*}
    \ell_3(x,u)
    =
    \left( \alpha \gamma(x,\eta)r(x,\eta)u^2 - \{\gamma(x,\eta)\DG r(x,\eta) + F(x,\eta)\} u \right)
    \partial_u g(x,u) .
\end{align*}

First note that since $\DG$ acts on space, it commutes with the integral over levels, and so
\begin{align*}
    \int_0^\infty \ell_1(x,u) du
    &=
    \gamma(x,\eta)\left(
        \DG(h(\cdot) r(\cdot,\eta))(x) - h(x) \DG r(x,\eta)
    \right) ,
\end{align*}
since $h(x) = \int_0^\infty (g(x,u) - 1) du$.
Next,
\begin{align*}
    \int_0^\infty \ell_2(x,u) du
    &=
    \alpha \gamma(x,\eta) r(x,\eta) 2\int_0^\infty g(x,u) \int_u^\infty (g(x,v)-1) dv du .
\end{align*}
Finally, integrating by parts,
\begin{align*}
    \int_0^\infty \ell_3(x,u) du
    &=
    - \alpha \gamma(x,\eta) r(x,\eta) \int_0^\infty 2u(g(x,u) - 1) du
    \\ & \qquad {}
    + \{ \gamma(x,\eta) \DG r(x,\eta) + F(x,\eta) \} h(x)
\end{align*}
Now, note that
\begin{align*}
&
    \int_0^\infty g(x,u) \int_u^\infty (g(x,v)-1) dv du - \int_0^\infty u(g(x,u)-1)du
\\ & \qquad =
    \int_0^\infty g(x,u) \int_u^\infty (g(x,v)-1) dv du - \int_0^\infty \int_v^\infty (g(x,u) -1) du dv
\\ & \qquad =
    \int_0^\infty (g(x,u)-1) \int_u^\infty (g(x,v)-1) dv du
\\ & \qquad =
    h(x)^2/2 .
\end{align*}

Adding these together, we get that
\begin{align*}
    &
    \int_0^\infty (\ell_1(x,u) + \ell_2(x,u) + \ell_3(x,u)) du
    \\ &\qquad=
    \gamma(x,\eta) \DG(h(\cdot)r(\cdot,\eta))(x)
    + F(x,\eta) h(x) 
    + \alpha \gamma(x,\eta) r(x,\eta) h(x)^2 ,
\end{align*}
which agrees with \eqref{eqn:pgen_infty}, as desired.
\end{proof}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Technical Lemmas}

\subsection{Gaussian Kernel Estimates}

\begin{lemma}\label{eq: Gaussin density estimates}
\begin{equation}
 \theta H_{\theta}(x,w)\rho_r(x-w)\leq \rho_{\gamma}(x-w),
\end{equation}
where
$$H_{\theta}(x,w)=\int_{\IR^d}
    \left(
        \frac{\rho_r(y-w)}{\rho_r(x-w)}
        -
        1
    \right)^2
q_{\theta}(x,dy).$$
\end{lemma}
\begin{proof}

Writing out $H$,
\begin{align}
H_\theta(x, w)
=
\int
    \left(
        \exp\left(
            - \frac{1}{2\epsilon_r^2}
            \left\{
                ||y - w||^2 - ||x - w||^2
            \right\}
        \right)
        -
        1
    \right)^2
    \frac{
        e^{-\theta (y - x-\meanq)^{T}\covq^{-1}(y - x-\meanq)  / 2}
    }{
        (2 \pi/ \theta)^{d/2} |\covq|^{1/2}
    }
dy.
\end{align}
Writing $z=y-x-\meanq$, we have that 
\begin{equation}
\begin{aligned}
||y - w||^2 - ||x - w||^2 
=& ||y||^2-||x||^2 -2\langle y-x, w\rangle \\
=&||x+z+ \meanq||^2-||x||^2-2 \langle z+ \meanq, w\rangle\\
=&2\langle x, z+ \meanq \rangle + ||z+ \meanq||^2-2 \langle z+ \meanq, w\rangle \\
=& 2 \langle x-w, z+ \meanq \rangle + ||z+ \meanq||^2\\
=& 2\langle x-w+ \meanq,\meanq \rangle + 2\langle x-w+ \meanq, z \rangle +||z||^2
\end{aligned}
\end{equation}
As a result, we have 
\begin{align}
H_\theta(x, w)
=&
\int
    \left(
        \exp\left(
            - \frac{1}{2\epsilon_r^2}
            \left\{
                2 \langle x-w, z+ \meanq \rangle + ||z+ \meanq||^2
            \right\}
        \right)
        -
        1
    \right)^2
    \frac{
        e^{-\theta z^{T}\covq^{-1}z  / 2}
    }{
        (2 \pi/ \theta)^{d/2} |\covq|^{1/2}
    }
dz\\
=&\exp\left(-\frac{2}{\epsilon_r^2}\langle x-w+ \meanq,\meanq \rangle \right)\\
& \times \int
        \frac{
        \exp\left(
            - \frac{2}{\epsilon_r^2} \langle x-w+ \meanq, z \rangle
            - \frac{1}{\epsilon_r^2} ||z||^2
            \right)
            e^{-\theta z^{T}\covq^{-1}z  / 2}
    }{
        (2 \pi/ \theta)^{d/2} |\covq|^{1/2}
    }
dz\\ \label{eq: Second moment term matrix calculation}
&-2\exp\left(-\frac{1}{\epsilon_r^2}\langle x-w+ \meanq,\meanq \rangle \right)\\
& \qquad \qquad \times     \int
    \frac{ \exp\left(
             -\frac{1}{\epsilon_r^2}\langle x-w+ \meanq, z \rangle
              -\frac{1}{2\epsilon_r^2} ||z||^2
            \right)e^{-\theta z^{T}\covq^{-1}z  / 2}
        }{
        (2 \pi/ \theta)^{d/2} |\covq|^{1/2} \label{eq: first moment term matrix calculation}
    }
dz\\
&+1.
\end{align}
We now consider the integral of the form 
\begin{equation}
\int \exp(v^{T}z-C||z||^2) \frac{
        e^{-\theta z^{T}\covq^{-1}z  / 2}
    }{
        (2 \pi/ \theta)^{d/2} |\covq|^{1/2}
    }
dz   
\end{equation}
for $v \in \IR^d$ and $C>0$. 
We have 
\begin{equation}
\begin{aligned}
&\int \exp(v^{T}z-C||z||^2) \frac{
        e^{-\theta z^{T}\covq^{-1}z  / 2}
    }{
        (2 \pi/ \theta)^{d/2} |\covq|^{1/2}
    }
dz\\
=&\frac{1}{|\covq|^{1/2}|  |\covq^{-1}(I+2C\covq/\theta )|^{1/2}  }\int \exp(v^{T}z) \frac{
        e^{-\theta z^{T}\covq^{-1}(I+2C\covq/\theta )z  / 2}
    }{
        (2 \pi/ \theta)^{d/2}|\covq^{-1}(I+2C\covq/\theta )|^{-1/2} 
    }
dz\\
=& \frac{1}{|I+2C\covq/\theta |^{1/2}} \mathbb{E}_{Z}[\exp (v^{T}Z)]\\
=&\frac{1}{|I+2C\covq/\theta |^{1/2}}
\exp 
\left(
\frac{1}{2}v^{T}\covq/\theta(I+2C\covq/\theta )^{-1}v
\right),
\end{aligned}
\end{equation}
where $Z \sim N(0, \covq/\theta \times (I+2C\covq/\theta )^{-1})$.

Now writing $v=\frac{2}{\epsilon_r^2}( w-x-\meanq)$
and $C=\frac{1}{\epsilon_r^2}$,
we have the expression in Equation
\eqref{eq: Second moment term matrix calculation} is 
$$\frac{1}{|I+2\covq/(\epsilon_r^2\theta) |^{1/2}}
\exp 
\left(
\frac{2}{\epsilon_r^4}( w-x-\meanq)^{T}\covq/\theta\left(I+2\covq/(\epsilon_r^2\theta)\right)^{-1}( w-x-\meanq)
\right).$$
Similarly, writing $v=\frac{1}{\epsilon_r^2}( w-x-\meanq)$
and $C=\frac{1}{2\epsilon_r^2}$,
we have the expression in Equation
\eqref{eq: first moment term matrix calculation} is 
$$\frac{1}{|I+\covq/(\epsilon_r^2\theta) |^{1/2}}
\exp 
\left(
\frac{1}{2\epsilon_r^4}( w-x-\meanq)^{T}\covq\left(I+\covq/\theta/(\epsilon_r^2\theta)\right)^{-1}( w-x-\meanq)
\right).$$
As a result, we have 
\begin{align*}
&H_{\theta}(x,w)\\
=& \frac{1}{|I+2\covq/(\epsilon_r^2\theta) |^{1/2}}
\exp 
\left(
\frac{2}{\epsilon_r^4}( w-x-\meanq)^{T}\covq/\theta\left(I+2\covq/(\epsilon_r^2\theta)\right)^{-1}( w-x-\meanq)
\right)\\
&-2\frac{1}{|I+\covq/(\epsilon_r^2\theta) |^{1/2}}
\exp 
\left(
\frac{1}{2\epsilon_r^4}( w-x-\meanq)^{T}\covq/\theta\left(I+\covq/(\epsilon_r^2\theta)\right)^{-1}( w-x-\meanq)
\right)\\
&+1.
\end{align*}
If we write $F(1/\theta)=H_{\theta}(x,w)$, i.e.
\begin{align*}
&F(s)\\
=& \frac{1}{|I+2/\epsilon_r^2 \covq s |^{1/2}}
\exp 
\left(
\frac{2}{\epsilon_r^4}( w-x-\meanq)^{T}\covq s \left(I+2/\epsilon_r^2 \covq s\right)^{-1}( w-x-\meanq)
\right)\\
&-2\frac{1}{|I+2/\epsilon_r^2 \covq s |^{1/2}}
\exp 
\left(
\frac{1}{2\epsilon_r^4}( w-x-\meanq)^{T}\covq s \left(I+2/\epsilon_r^2 \covq s\right)^{-1}( w-x-\meanq)
\right)\\
&+1\\
&=G_1(s)+2G_2(s)+1
\end{align*}
we can check that $F(0)=0$.
As a result, 
$$\theta H_{\theta}(x,w)\rho_r(x-w)= \theta F(1/\theta)\rho_r(x-w)=F'(s)\rho_r(x-w),$$
for some $s \in [0, 1/\theta]$.

In Equation \eqref{eq: gamma-r convolution comparison}, we have established that 
$\rho_r(x-w) / \rho_\gamma(x-w) = C' \exp(-K (x-w)^2 / 2)$
with $K = \frac{1}{\epsilon_r^2} - \frac{1}{\epsilon_\gamma^2}$.
Therefore it suffices to show that for all $x,w \in \IR^d$ and $s \in [0, 1/\theta]$, 
\begin{align}\label{eqn:goal3}
    F'(s) \exp\left(-A (x-w)^2 \right) \le C .
\end{align}

We consider how to differentiate function $G$ of the form 
$$G(s):=\frac{1}{|I+As|^{1/2}}\exp(v^{T}Bs(I+Cs)^{-1}v),$$
where $A,B,C$ are matrices with non-negative entries and $I$ is the identity matrix.

Since 
\begin{equation}
\begin{aligned}
G'(s)=& - \frac{|A|}{2|I+As|^{3/2}}\exp(v^{T}Bs(I+Cs)^{-1}v)\\
      & + \frac{1}{|I+As|^{1/2}}\exp(v^{T}Bs(I+Cs)^{-1}v)\\
      & \qquad \times \left\{ v^{T}B(I+Cs)^{-1}v
                            - sv^{T}BC(I+Cs)^{-2}v
                    \right\}\\
      &=\left(- \frac{|A|}{2|I+As|}+v^{T}B(I+Cs)^{-2}v
        \right)\\
      & \qquad \times
        \frac{1}{|I+As|^{1/2}}\exp(v^{T}Bs(I+Cs)^{-1}v),
\end{aligned}
\end{equation}
using non-negativity of $A,B,C$ we have
\begin{equation}
\label{eq: bounds on G'}
|G'(s)| \leq (|A|+||B||_{2}||v||_2^2)G(s).    
\end{equation}
As a result, we have that 
\begin{equation}
\label{eq: bounds on G}
G(s) \leq K e^{|A|+||B||_{2}||v||_2^2)s}, ~~   G'(s)(|A|+||B||_{2}||v||_2^2)K e^{|A|+||B||_{2}||v||_2^2)s}.  
\end{equation}

Finally note that, 
$$F'(s) \leq G_1'(s)+2G_2'(s),$$
where 
$$A_1= 2\epsilon_r^{-2} \covq, ~ v_1=\epsilon_r^{-2}(w-x-\meanq), ~B_1 = 2\covq, ~C_1=2\epsilon_r^{-2} \covq,$$
and 
$$A_2= 2\epsilon_r^{-2} \covq, ~ v_2=\epsilon_r^{-2}(w-x-\meanq), ~B_2 = \covq/2, ~C_2=2\epsilon_r^{-2} \covq,$$
Substituting these quantities on Equation \eqref{eq: bounds on G} shows that Equation \eqref{eqn:goal3} is satisfied as long as $\epsilon_r < \epsilon_\gamma$.

\end{proof}

\subsection{Constraints on kernel widths}

\begin{lemma}
    \label{lem:gamma_bound}
\comment{TODO; REWRITE TO SAY THAT STHG IMPLIES CONDITION 2 OF LEMMA \ref{lem:conditions on r} (see reference on p.13)}
    Under the assumptions of Definition~\ref{def:model_setup},
    for all $f \in \mathcal{C}^{2}_{0}(\IR^d \times [0,\infty))$,
    there exists a $K_f < \infty$ such that
    for all $\theta > 0$ and all $\eta \in \measures$,
    \begin{equation} \label{eqn:gamma_bound}
    \gamma\big(x,\smooth{\gamma} \eta(x)\big)
    \left| \theta 
        \int_{\IR^n} 
            \left\{
                f(y, \smooth{r} \eta(y)) - f(x, \smooth{r} \eta(x))
            \right\}
        q_{\theta}(x,dy)
    \right|
    \leq C_f     .
    \end{equation}
\end{lemma}

The key assumption here from Definition~\ref{def:model_setup}
will be that the spatial scale over which local density affects birth rate
is larger than the scale over which it affects establishment:
concretely, that $\epsilon_{r}^2 + \frac{2\lambda}{\theta} < \epsilon_{\gamma}^2$,
where $\lambda = \sup_x \sup_{y : \|y\| = 1} y^T \covq(x) y$
is the largest eigenvalue of $\covq(x)$ across all $x$.
In the simple case of $\meanq = 0$ and $\covq = \sigma^2 I$,
the condition is simply that $\epsilon_r^2 + 2\sigma^2 / \theta < \epsilon_\gamma^2$.

\begin{proof}[of Lemma \ref{lem:gamma_bound}]

    Since we have assumed that there is a $K$ for which $m^2 \gamma(x, m) < K$
    for all $x \in \IR^d$ and $m \ge 0$,
    equation~\eqref{eqn:gamma_bound} will follow if we can show that
    \begin{equation*}
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
        \leq
        C_f \left( 1 + \smooth{\gamma} \eta(x)^2 \right) .
    \end{equation*}
    To do this, first note that 
    \begin{align} \label{eqn:q_bound_split}
    &
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad {} \leq
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(y, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(y)\right)
                \right\}
            q_{\theta}(x,dy)
        \right|
    \\ &\qquad \qquad {}  +
        \left| \theta 
            \int_{\IR^n} 
                \left\{
                    f\left(x, \smooth{r}\eta(y)\right)
                    -
                    f\left(x, \smooth{r}\eta(x)\right)
                \right\}
            q_{\theta}(x,dy) 
        \right| .
    \end{align}

    Since 
    \begin{align*}
        f\left(y, \smooth{r}\eta(y)\right) - f\left(x, \smooth{r}\eta(y)\right)
        &=
        \sum_i (y_i - x_i) \partial_{x_i} f \left(x, \smooth{r}\eta(y)\right) 
        \\ &\qquad {}
        +
        \frac{1}{2} \sum_{ij} (y_i - x_i)(y_j - x_j)
            \partial_{x_i x_j} f \left(z, \smooth{r}\eta(y)\right) 
    \end{align*}
    where $z = t x + (1-t) y$ for some $0 \le t \le 1$,
    writing
    $K_1 = \sup_{x,m} \max_i |\partial_{x_i} f(x, m)|$
    and 
    $K_2 = \sup_{x,m} \max_{i,j} |\partial_{x_i x_j} f(x, m)|$
    for the bounds on the spatial derivatives of $f$,
    the first term is bounded by
    \begin{align*}
        K_1 \left|
            \theta \int_{\IR^d} \sum_i (y_i - x_i) q_\theta(x, dy)
        \right|
        +
        K_2 \left|
            \theta \int_{\IR^d} \sum_{ij} (y_i - x_i) (y_j - x_j)  q_\theta(x, dy)
        \right| .
    \end{align*}
    This is bounded uniformly because
    $q_\theta(x, dy)$ is the density of a Gaussian with mean $\meanq(x)/\theta$
    and covariance $\covq(x)/\theta$, and both $\meanq(x)$ and $\covq(x)$
    are bounded uniformly in $x$.

    Similarly, 
    \begin{align*}
        f(x, \smooth{r}\eta(y)) - f(x, \smooth{r}\eta(x))
        &=
        (\smooth{r}\eta(y) - \smooth{r}\eta(x)) f'(x, \smooth{r}\eta(x))
        + \frac{1}{2} (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2 f''(x, m)
    \end{align*}
    for some $\smooth{r}\eta(x) \le m \le \smooth{r}\eta(y)$,
    where $f'$ and $f''$ denote the first and second derivatives with respect to the second argument.
    So,
    the second term in~\eqref{eqn:q_bound_split} is bounded by
    \begin{align*}
        % note that the f' term does not depend on x, just y,
        % so can come out of the integral
        \|f'\|_\infty
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))
            q_\theta(x, dy)
        \right|
        +
        \|f''\|_\infty
        \left|
            \theta \int_{\IR^d}
                (\smooth{r}\eta(y) - \smooth{r}\eta(x))^2
            q_\theta(x, dy)
        \right| .
    \end{align*}
    Therefore, it suffices to show that there is a $K < \infty$ such that
    for all $\eta \in \measures$ and $x \in \IR^d$,
    \begin{equation}
        \label{eqn:first_moment_rho}
        \left|\theta \int_{\IR^d}
            ( \smooth{r} \eta(y)-\smooth{r} \eta(x) )
        q_\theta(x,dy)\right|
        \leq K \smooth{\gamma} \eta(x)
    \end{equation}
    and
    \begin{equation}
        \label{eqn:second_moment_rho}
        \theta \int_{\IR^d}
                \left( \smooth{r} \eta(y) - \smooth{r} \eta(x) \right)^2
        q_\theta(x,dy)
        \leq K \left(\smooth{\gamma}\eta(x) \right)^2 .
    \end{equation}
    Note that the right hand side of each is the average density
    over a wider region (since $\epsilon_\gamma > \epsilon_r$).

    First we prove \eqref{eqn:first_moment_rho}.
    Recall that $\smooth{r} \eta(x) = \int p_{\epsilon^2_r}(x - w) \eta(dw)$,
    where $p_t$ is the density of a Gaussian with mean 0 and variance $t$,
    so that applying Fubini, \eqref{eqn:first_moment_rho} is
    \begin{align*}
        \left|
        \int_{\IR^d}
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            q_\theta(x, dy)
        \eta(dw)
        \right| .
    \end{align*}
    Write $p_{s, x}(\cdot)$ for the density of a Gaussian
    with mean $s \meanq(x)$ and covariance $\epsilon_r^2 I + s \covq(x)$,
    so that $\int p_{\epsilon_r^2}(y - w) q_\theta(x, dy) = p_{1/\theta, x}(w-x)$.
    It therefore suffices to show that for all $x$ and $w \in \IR^d$,
    there exists $K$ such that
    \begin{align} \label{eqn:goal1}
        \left|
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y-w) - p_{\epsilon_r^2}(x-w) )
            q_\theta(x, dy)
        \right|
        &=
        \theta \left|
            p_{1/\theta, x}(w-x) - p_{0, x}(w-x)
        \right|
        \le
        K p_{\epsilon^2_\gamma}(w-x) .
    \end{align}
    However,
    $\theta( p_{1/\theta, x}(z) - p_{0, x}(z) )
    = \partial_s p_{s, x}(z)$ for some $0 \le s \le 1/\theta$.
    Write $\Gamma(s,x) = \epsilon_r^2 I + s \covq(x)$,
    so that
    \begin{align*}
        p_{s, x}(z)
        &=
        \frac{1}{\left(2 \pi |\Gamma(s,x)|\right)^{d/2}}
        \exp\left(
            -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x))
        \right) ,
    \end{align*}
    and note that if $\lambda_i$ are the eigenvalues of $\covq(x)$
    then $|\Gamma(s, x)| = \prod_i (\epsilon_r^2 + s \lambda_i)$,
    and
    $\partial_s |\Gamma(s, x)| = \sum_i \lambda_i |\Gamma(s, x)| / (\epsilon_r^2 + s \lambda_i)$.
    Therefore,
    \begin{align*}
        \partial_s p_{s, x}(z)
        &=
        \left(
            \meanq(x)^T \Gamma(s,x)^{-1} (z - s \meanq)
            +
            (z - s \meanq(x))^T
            \Gamma(s,x)^{-1} \covq(x) \Gamma(s,x)^{-1}
            (z - s \meanq(x))
        \right. \\ & \qquad \qquad \qquad \left. {}
            -
            \sum_i \frac{\lambda_i}{\epsilon_r^2 + s \lambda_i}
        \right)
        p_{s, x}(z) ,
    \end{align*}
    where $z^T$ is the transpose of $z$.
    This implies that
    \begin{align*}
        \frac{
        \theta
            \int_{\IR^d}
                ( p_{\epsilon_r^2}(y-w) - p_{\epsilon_r^2}(x-w) )
            q_\theta(x, dy)
        }{
            p_{\epsilon^2_\gamma}(x-w) .
        }
        =
        h(x-w) e^{k(x-w)},
    \end{align*}
    where $h(z)$ and $k(z)$ are quadratic polynomials in $z$
    whose coefficients depend on $s$ and $x$ but are uniformly bounded for $s \ge 1$
    and $x \in \IR^d$.
    Therefore, \eqref{eqn:goal1} holds for some finite $K$
    if and only if $k(z) < 0$ for all $z$ outside some bounded region.
    This is
    \begin{align*}
        k(z)
        =
        \frac{1}{2\epsilon_\gamma^2} \|z\|^2
        -\frac{1}{2} (z - s\meanq(x))^T \Gamma(s,x)^{-1} (z - s\meanq(x)) ,
    \end{align*}
    which is negative for all $z$ outside a bounded region.
    Since $\inf_z z^T \Gamma(s,x)^{-1} z / \|z\|^2 = 1 / (s \lambda(x) + \epsilon_r^2)$,
    where $\lambda(x) = \sup z^T \covq(x) z / \|z\|^2$ is the largest eigenvalue of $\covq(x)$,
    equation~\eqref{eqn:first_moment_rho} follows from the assumption
    that $\epsilon_r^2 + 2 \sup_x \lambda(x)/\theta < \epsilon_\gamma^2$.
    (Note that we do not yet need the factor of 2.)

    Next we prove equation~\eqref{eqn:second_moment_rho}, in a similar way.
    Again applying Fubini,
    \begin{align*}
        &
        \theta \int_{\IR^d}
                \left( \smooth{r} \eta(y) - \smooth{r} \eta(x) \right)^2
        q_\theta(x,dy)
        \\ &\qquad
        =
        \int_{\IR^d} \int_{\IR^d}
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \eta(dv) \eta(dw) ,
    \end{align*}
    and so as before, equation~\eqref{eqn:second_moment_rho} will follow if
    the integrand is bounded by $K p_\gamma(x-w) p_\gamma(x - v)$.
    % \begin{align*}
    %     \theta \int_{\IR^d}
    %         ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
    %         ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
    %     q_\theta(x, dy)
    %     \le K p_\gamma(x-w) p_\gamma(x - v) .
    % \end{align*}
    Now, let $Y_1$, $Y_2$, and $Z$ be independent $d$-dimensional Gaussians with mean zero,
    where $Y_1$ and $Y_2$ have covariance $\epsilon^2_r I$,
    and $Z$ has covariance $\covq(x)$.
    Write $p_{s,t,x}(\cdot, \cdot)$ for the joint density of
    $Y_1 + \sqrt{s} Z + s \meanq(x)$ and $Y_2 + \sqrt{t} Z + t \meanq(x)$.
    Then, observe that
    \begin{align*}
        &
        \theta \int_{\IR^d}
            ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
            ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
        q_\theta(x, dy)
        \\ &\qquad
        =
        \theta \left(
            p_{1/\theta, 1/\theta, x}(x-w, x-v)
            - p_{0, 1/\theta, x}(x-w, x-v)
        \right. \\ &\qquad \qquad \left. {}
            - p_{1/\theta, 0, x}(x-w, x-v)
            + p_{0, 0, x}(x-w, x-v)
        \right) 
        \\ &\qquad 
        =
            \partial_s p_{s, 1/\theta, x}(x-w, x-v)
            - \partial_t p_{0, t, x}(x-w, x-v) ,
    \end{align*}
    for some $0 \le s, t \le 1/\theta$.
    As before,
    \begin{align*}
        \frac{
            \theta \int_{\IR^d}
                ( p_{\epsilon_r^2}(y - w) - p_{\epsilon_r^2}(x - w) )
                ( p_{\epsilon_r^2}(y - v) - p_{\epsilon_r^2}(x - v) )
            q_\theta(x, dy)
        }{
            p_{\epsilon_\gamma^2}(x - w)
            p_{\epsilon_\gamma^2}(x - v)
        }
        =
        h(x-w, x-v) e^{ k(x-w, x-v) },
    \end{align*}
    where $h(z_1, z_2)$ is a polynomial with uniformly bounded coefficients and
    \begin{align*}
        k(z_1, z_2)
        =
        (\|z_1\|^2 + \|z_2\|^2)/(2 \epsilon_\gamma^2)
            - \frac{1}{2} [z_1, z_2[^T \Gamma(s, t, x)^{-1} [z_1, z_2] ,
    \end{align*}
    where $[z_1, z_2]$ is the $\IR^{2d}$ vector formed by concatenating $z_1$ and $z_2$,
    and $\Gamma(s, t, x)$ is the block matrix
    \begin{align*}
        \Gamma(s, t, x)
        =
        \left[
        \begin{array}{cc}
            \epsilon_r^2 I + s \covq(x) & \sqrt{st} \covq(x) \\
            \sqrt{st} \covq(x) & \epsilon_r^2 I + t \covq(x) \\
        \end{array}
        \right] .
    \end{align*}
    If $\covq(x) u = a u$ for some $a \in \IR$,
    then $[u \sqrt{s}, u \sqrt{t}]$ is an eigenvector of $\Gamma(s, t, x)$
    with eigenvalue $\epsilon^2_r + (s+t) a$,
    and $[u \sqrt{t}, - u \sqrt{s}]$ is an eigenvector of $\Gamma(s, t, x)$
    with eigenvalue 0.
    This implies the largest eigenvalue of $\Gamma(s, t, x)$
    is equal to $\epsilon^2_r + (s+t) \lambda(x)$,
    where $\lambda(x)$ is again the largest eigenvalue of $\covq(x)$.
    Therefore, if $s + t \le 2 / \theta$,
    \begin{align*}
        &
        (\|z_1\|^2 + \|z_2\|^2) / \epsilon^2_\gamma
            - [z_1, z_2[^T \Gamma(s, t, x)^{-1} [z_1, z_2]
        \\ &\qquad \le
        (\|z_1\|^2 + \|z_2\|^2) \left(
            \frac{1}{\epsilon^2_\gamma}
            - \frac{1}{\epsilon^2_r +  2\lambda(x)/\theta}
        \right) ,
    \end{align*}
    which is negative by assumption.
    Therefore, there is a $K$ such that
    \begin{align*}
        \frac{ \left|
            \partial_s p_{s, 1/\theta, x}(x-w, x-v)
            - \partial_t p_{0, t, x}(x-w, x-v)
        \right| }{
            p_{\epsilon^2_\gamma}(x - w)
            p_{\epsilon^2_\gamma}(x - v)
        }
        \le K
    \end{align*}
    for all $\theta > 1$ and all $x$, $v$, and $w \in \IR^d$,
    proving equation~\eqref{eqn:second_moment_rho} and hence the lemma.
\end{proof}

\subsection{Tightness of processes}
\label{apx:aldous rebolledo}

We record for reference above
one version of the Aldous-Rebolledo criteria for tightness of a sequence real-valued processes
(as it appears in \citet{etheridge2000introduction}):
\comment{TODO: reference this elsewhere, if we keep it?}

\begin{theorem}[\citet{rebolledo:1980}]
    \label{thm:aldous_rebolledo}
    Let $\{Y^(n)\}_{n \ge 1}$ be a sequence of real-valued processes
    with c\`adl\`ag paths.
    Suppose that the following conditions are satisfied.
    \begin{enumerate}
        \item For each fixed $t$, $\{Y_t^{(n)}\}_{n \ge 1}$  is tight.
        \item Given a sequence of stopping times $\tau_n$, bounded by $T$,
            for each $\epsilon > 0$ there exists $\delta > 0$ and $n_0$ such that
            \begin{align*}
                \sup_{n \ge n_0}
                \sup_{\theta \in [0, \delta]}
                \IP\left\{ \left|
                    Y^{(n)}_{\tau_n + \theta} - Y^{(n)}_{\tau_n}
                \right| > \epsilon
                \right\}
                \le \epsilon .
            \end{align*}
    \end{enumerate}
    Then the sequence $\{Y^{(n)}\}_{n \ge 1}$ is tight.
\end{theorem}

\begin{lemma}
    \label{lem:product_tightness}
    Let $\{X^{(n)}\}_{n \ge 1}$ and $\{Y^{(n)}\}_{n \ge 1}$ sequences of jointly defined
    real-valued processes with c\`adl\`ag paths
    satisfying the conditions of Theorem~\ref{thm:aldous_rebolledo}.
    Then $\{X^{(n)}Y^{(n)}\}_{n \ge 1}$ and $\{X^{(n)} + Y^{(n)}\}_{n \ge 1}$ also
    satisfy the conditions of Theorem~\ref{thm:aldous_rebolledo}.
\end{lemma}

By ``jointly defined'' we mean that $X^{(n)}$ and $Y^{(n)}$ are defined on the same
probability space, so that the products and sums make sense.

\begin{proof}[of Lemma~\ref{lem:product_tightness}]
    The proof for $X^{(n)} + Y^{(n)}$ is similar to but more straightforward than
    for $X^{(n)} Y^{(n)}$, so on only prove the Lemma for the latter.
    First, note that for any $\epsilon > 0$, by tightness of $X^{(n)}_t$ and $Y^{(n)}_t$
    there is a $K$ such that
    $\IP\{ X^{(n)}_t > \sqrt{K} \}$ and $\IP\{ Y^{(n)}_t > \sqrt{K} \}$
    are both less than $\epsilon/2$,
    and hence
    \begin{align*}
        \IP\{
            X^{(n)}_t Y^{(n)}_t > K
        \}
        \le
        \IP\{
            X^{(n)}_t > \sqrt{K}
        \}
        +
        \IP\{
            Y^{(n)}_t > \sqrt{K}
        \} 
        \le \epsilon.
    \end{align*}
    Therefore, $X^{(n)}_t Y^{(n)}_t$ is tight.

    Next, note that for $0 \le \tau_n \le T$,
    \begin{align*}
        &
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        \\ &\qquad 
        \le
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left| X^{(n)}_{\tau_n + \theta} \right|
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \left| Y^{(n)}_{\tau_n} \right|
        \\ &\qquad 
        \le
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
        +
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta}
            -
            X^{(n)}_{\tau_n}
        \right|
        \sup_{0 \le t \le T}
        \left| Y^{(n)}_t \right| ,
    \end{align*}
    so that for any $C$,
    \begin{align} \label{eqn:xy_bound}
        \begin{split}
        &
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} Y^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} Y^{(n)}_{\tau_n} 
        \right|
        > \epsilon
        \right\}
    \\ &\qquad \le
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| X^{(n)}_t \right|
            > C
        \right\}
    +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            Y^{(n)}_{\tau_n + \theta} 
            -
            Y^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    \\ &\qquad {} +
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
    +
        \IP\left\{
            \sup_{0 \le t \le T}
                \left| Y^{(n)}_t \right|
            > C
        \right\}
        \end{split}
    \end{align}
    Now, since $\max_{0 \le t \le T} X^{(n)}_t$ is tight (and likewise for $Y$)
    (see, e.g., Remark 3.7.3 in \citet{EK}),
    we may choose a $C \ge 4$ for which
    $$
    \IP\left\{
        \sup_{0 \le t \le T}
            \left| X^{(n)}_t \right|
        > C
    \right\} \le \frac{\epsilon}{4} .
    $$
    Simliarly, by assumption we can choose a $\delta$ for which
    $$
        \IP\left\{
        \sup_{0 \le \theta \le \min(\delta, T-\tau_n)} 
        \left|
            X^{(n)}_{\tau_n + \theta} 
            -
            X^{(n)}_{\tau_n} 
        \right|
            > \epsilon/C
        \right\}
        \le \frac{\epsilon}{C} .
    $$
    If we choose $C$ and $\delta$ that do this for both $X^{(n)}$ and $Y^{(n)}$,
    then each of the terms in equation~\eqref{eqn:xy_bound} are bounded by $\epsilon/4$,
    and condition (2) is satisfied for the product process.
\end{proof}

\bibliographystyle{plainnat}
\bibliography{refs,plr_refs,MendeleyIan}

\end{document}